---
title: "table one sedated"
output: html_document
---

## Sup Table 2: Not Sedated

All variables.  Stratification (ie column groupings) should be by sodium categories.

```{r,echo = F}
dataframe_for_table1<-df_sub_notsedated


# we are re-difining the 'Others' category
dataframe_for_table1$unittype<-plyr::revalue(
  dataframe_for_table1$unittype,c( 'Neuro ICU'='Others' , 'SICU'= 'Others', 'Cardiac ICU'='Others' , 'CCU-CTICU'='Others'  ,'CSICU'='Others' , 'CTICU'='Others')
)

vars_in_table1<-c("age", "gender", "AP_grp", "ethnicity", "apachescore", "mechvent_day01", "unittype", "charlson_chf1", "charlson_liver1", "creatinine", "creatinine_great2", "gcs_baseline", "normal_gcs", "hosp_mortality","sodium_cat","adm_source","sedation_status")
# lists the variables we are going to be including in table1 (excepts for the stratification variable)

dataframe_for_table1<-dataframe_for_table1[,vars_in_table1]
cat_variables<-rep(NA, length(vars_in_table1))
# creates a dataset only with select variable using our original dataset and generated an empty list that is going to be used for storing categorical variables 

stratifyby<-"sodium_cat"
label(dataframe_for_table1$age) <- 'Age'
label(dataframe_for_table1$gender) <- 'Gender'
label(dataframe_for_table1$AP_grp) <- 'APACHE Diagnosis Group'
label(dataframe_for_table1$ethnicity) <- 'Ethnicity'
label(dataframe_for_table1$apachescore) <- 'APACHE Score'
label(dataframe_for_table1$mechvent_day01) <- 'Mechanical Ventilation on Day 1'
label(dataframe_for_table1$unittype) <- 'ICU Type'
label(dataframe_for_table1$charlson_chf1) <- 'Charlson Comorbidity - Congestive Heart Failure'
label(dataframe_for_table1$charlson_liver1) <- 'Charlson Comorbidity - Liver Disease'
label(dataframe_for_table1$creatinine) <- 'Creatinine'
label(dataframe_for_table1$creatinine_great2) <- 'Creatinine > 2'
label(dataframe_for_table1$gcs_baseline) <- 'Baseline Glasgow Coma Scale'
label(dataframe_for_table1$normal_gcs) <- 'Normal Glasgow Coma Scale'
label(dataframe_for_table1$hosp_mortality) <- 'Hospital Mortality'
label(dataframe_for_table1$sodium_cat) <- 'Sodium Category'
label(dataframe_for_table1$adm_source) <- 'Admission Source'
label(dataframe_for_table1$sedation_status) <- 'Sedation Status'
# selects variable to stratify by and label is useful for prettyfying variable names we want to appear on table 1

# detects whether a variable is categorical or not based on the number of distinct variables per category (10 is the threshold)
cont<-1
for (i in 1:length(vars_in_table1) ) {
  if ( n_distinct(dataframe_for_table1[vars_in_table1[i] ])<=10 ) {
    print(i)
    print(vars_in_table1[i])
    print(names(dataframe_for_table1[vars_in_table1[i]]))
    cat_variables[cont]<-names(dataframe_for_table1[vars_in_table1[i]])
    cont<-cont+1
  }
}  
cat_variables<-cat_variables[!is.na(cat_variables)]
sup_table2<-print(CreateTableOne(vars = vars_in_table1
                                   ,strata = stratifyby
                                  , factorVars = cat_variables
    ,data = dataframe_for_table1, addOverall=T),varLabels = T ,nonnormal  = c('creatinine','gcs_baseline')
)

sup_table2<-as.data.frame(sup_table2)

# Assume your data frame is named df
# Convert row names to a column
sup_table2 <- sup_table2 %>% 
  rownames_to_column(var = "RowNames")

# # Create the flextable with the modified dataframe
# table1_flextable <- flextable(sup_table2) %>%
#   fontsize(size = 6, part = "all")

sup_table2_flextable <- flextable(sup_table2) %>%
  fontsize(size = 6, part = "all") %>%
  set_table_properties(width = .95, layout = "autofit") %>% # Adjust the table width
  border_outer(border = fp_border(color = "black", width = 1)) %>%
  border_inner(border = fp_border(color = "black", width = 1)) %>%
  border(border = fp_border(color = "black", width = 1))

# Create a new Word document
doc <- read_docx()

# Add the flextable to the document
doc <- body_add_flextable(doc, value = sup_table2_flextable)

# Save the document
print(doc, target = "tables/sup_table2.docx")
```
## Sup Table 3: Sedated

All variables.  Stratification (ie column groupings) should be by sodium categories.

```{r,echo = F}
dataframe_for_table1<-df_sub_sedated


# we are re-difining the 'Others' category
dataframe_for_table1$unittype<-plyr::revalue(
  dataframe_for_table1$unittype,c( 'Neuro ICU'='Others' , 'SICU'= 'Others', 'Cardiac ICU'='Others' , 'CCU-CTICU'='Others'  ,'CSICU'='Others' , 'CTICU'='Others')
)

vars_in_table1<-c("age", "gender", "AP_grp", "ethnicity", "apachescore", "mechvent_day01", "unittype", "charlson_chf1", "charlson_liver1", "creatinine", "creatinine_great2", "gcs_baseline", "normal_gcs", "hosp_mortality","sodium_cat","adm_source","sedation_status")
# lists the variables we are going to be including in table1 (excepts for the stratification variable)

dataframe_for_table1<-dataframe_for_table1[,vars_in_table1]
cat_variables<-rep(NA, length(vars_in_table1))
# creates a dataset only with select variable using our original dataset and generated an empty list that is going to be used for storing categorical variables 

stratifyby<-"sodium_cat"
label(dataframe_for_table1$age) <- 'Age'
label(dataframe_for_table1$gender) <- 'Gender'
label(dataframe_for_table1$AP_grp) <- 'APACHE Diagnosis Group'
label(dataframe_for_table1$ethnicity) <- 'Ethnicity'
label(dataframe_for_table1$apachescore) <- 'APACHE Score'
label(dataframe_for_table1$mechvent_day01) <- 'Mechanical Ventilation on Day 1'
label(dataframe_for_table1$unittype) <- 'ICU Type'
label(dataframe_for_table1$charlson_chf1) <- 'Charlson Comorbidity - Congestive Heart Failure'
label(dataframe_for_table1$charlson_liver1) <- 'Charlson Comorbidity - Liver Disease'
label(dataframe_for_table1$creatinine) <- 'Creatinine'
label(dataframe_for_table1$creatinine_great2) <- 'Creatinine > 2'
label(dataframe_for_table1$gcs_baseline) <- 'Baseline Glasgow Coma Scale'
label(dataframe_for_table1$normal_gcs) <- 'Normal Glasgow Coma Scale'
label(dataframe_for_table1$hosp_mortality) <- 'Hospital Mortality'
label(dataframe_for_table1$sodium_cat) <- 'Sodium Category'
label(dataframe_for_table1$adm_source) <- 'Admission Source'
label(dataframe_for_table1$sedation_status) <- 'Sedation Status'
# selects variable to stratify by and label is useful for prettyfying variable names we want to appear on table 1

# detects whether a variable is categorical or not based on the number of distinct variables per category (10 is the threshold)
cont<-1
for (i in 1:length(vars_in_table1) ) {
  if ( n_distinct(dataframe_for_table1[vars_in_table1[i] ])<=10 ) {
    print(i)
    print(vars_in_table1[i])
    print(names(dataframe_for_table1[vars_in_table1[i]]))
    cat_variables[cont]<-names(dataframe_for_table1[vars_in_table1[i]])
    cont<-cont+1
  }
}  
cat_variables <- cat_variables[!is.na(cat_variables)]
sup_table3 <- print(CreateTableOne(vars = vars_in_table1,
                                    strata = stratifyby,
                                    factorVars = cat_variables,
                                    data = dataframe_for_table1, addOverall = TRUE), varLabels = TRUE, nonnormal = c('creatinine', 'gcs_baseline')
)

sup_table3<-as.data.frame(sup_table3)

# Assume your data frame is named df
# Convert row names to a column
sup_table3 <- sup_table3 %>% 
  rownames_to_column(var = "RowNames")

# # Create the flextable with the modified dataframe
# table1_flextable <- flextable(sup_table2) %>%
#   fontsize(size = 6, part = "all")

sup_table3_flextable <- flextable(sup_table3) %>%
  fontsize(size = 6, part = "all") %>%
  set_table_properties(width = .95, layout = "autofit") %>% # Adjust the table width
  border_outer(border = fp_border(color = "black", width = 1)) %>%
  border_inner(border = fp_border(color = "black", width = 1)) %>%
  border(border = fp_border(color = "black", width = 1))

# Create a new Word document
doc <- read_docx()

# Add the flextable to the document
doc <- body_add_flextable(doc, value = sup_table3_flextable)

# Save the document
print(doc, target = "tables/sup_table3.docx")
```

## Exploratory Table Suspected error

All variables.  Stratification (ie column groupings) should be by sodium categories.

```{r,echo = F}
dataframe_for_table1<-df_sub_susp_error


# we are re-difining the 'Others' category
dataframe_for_table1$unittype<-plyr::revalue(
  dataframe_for_table1$unittype,c( 'Neuro ICU'='Others' , 'SICU'= 'Others', 'Cardiac ICU'='Others' , 'CCU-CTICU'='Others'  ,'CSICU'='Others' , 'CTICU'='Others')
)

vars_in_table1<-c("age", "gender", "AP_grp", "ethnicity", "apachescore", "mechvent_day01", "unittype", "charlson_chf1", "charlson_liver1", "creatinine", "creatinine_great2", "gcs_baseline", "normal_gcs", "hosp_mortality","sodium_cat","adm_source","sedation_status")
# lists the variables we are going to be including in table1 (excepts for the stratification variable)

dataframe_for_table1<-dataframe_for_table1[,vars_in_table1]
cat_variables<-rep(NA, length(vars_in_table1))
# creates a dataset only with select variable using our original dataset and generated an empty list that is going to be used for storing categorical variables 

stratifyby<-"sodium_cat"
label(dataframe_for_table1$age) <- 'Age'
label(dataframe_for_table1$gender) <- 'Gender'
label(dataframe_for_table1$AP_grp) <- 'APACHE Diagnosis Group'
label(dataframe_for_table1$ethnicity) <- 'Ethnicity'
label(dataframe_for_table1$apachescore) <- 'APACHE Score'
label(dataframe_for_table1$mechvent_day01) <- 'Mechanical Ventilation on Day 1'
label(dataframe_for_table1$unittype) <- 'ICU Type'
label(dataframe_for_table1$charlson_chf1) <- 'Charlson Comorbidity - Congestive Heart Failure'
label(dataframe_for_table1$charlson_liver1) <- 'Charlson Comorbidity - Liver Disease'
label(dataframe_for_table1$creatinine) <- 'Creatinine'
label(dataframe_for_table1$creatinine_great2) <- 'Creatinine > 2'
label(dataframe_for_table1$gcs_baseline) <- 'Baseline Glasgow Coma Scale'
label(dataframe_for_table1$normal_gcs) <- 'Normal Glasgow Coma Scale'
label(dataframe_for_table1$hosp_mortality) <- 'Hospital Mortality'
label(dataframe_for_table1$sodium_cat) <- 'Sodium Category'
label(dataframe_for_table1$adm_source) <- 'Admission Source'
label(dataframe_for_table1$sedation_status) <- 'Sedation Status'
# selects variable to stratify by and label is useful for prettyfying variable names we want to appear on table 1

# detects whether a variable is categorical or not based on the number of distinct variables per category (10 is the threshold)
cont<-1
for (i in 1:length(vars_in_table1) ) {
  if ( n_distinct(dataframe_for_table1[vars_in_table1[i] ])<=10 ) {
    print(i)
    print(vars_in_table1[i])
    print(names(dataframe_for_table1[vars_in_table1[i]]))
    cat_variables[cont]<-names(dataframe_for_table1[vars_in_table1[i]])
    cont<-cont+1
  }
}  
cat_variables<-cat_variables[!is.na(cat_variables)]
table1_base<-print(CreateTableOne(vars = vars_in_table1
                                   ,strata = stratifyby
                                  , factorVars = cat_variables
    ,data = dataframe_for_table1, addOverall=T),varLabels = T ,nonnormal  = c('creatinine','gcs_baseline')
)


# run this in console for html output, the code below uses library kableExtra::
# the line add_header_above() will need to be modified depending on the number of columns of the table
starification_cats<-n_distinct(dataframe_for_table1[,stratifyby])
table1_base %>%
  kbl(caption = "Table 1 Suspected Error" , align = "c") %>%
  kable_classic_2(full_width = F, html_font = "Cambria")%>%
  add_header_above(c(" "," ", 'Sodium Band Category' = starification_cats," ", "" ))

# Function to save results to a Word document
save_to_word <- function(data, file_name, title) {
  doc <- read_docx()
  doc <- doc %>%
    body_add_par(title, style = "heading 1") %>%
    body_add_table(value = data, style = "table_template")
  
  print(doc, target = file_name)
}

table1_base<-as.data.frame(table1_base)

# Save overall elective and not elective results
save_to_word(table1_base, "tables/Exploratory Table Suspected error.docx","Exploratory Table Suspected error")

```

If suspected error looks a lot like “sedated” then it may be reasonable to group them together.
After reviewing table one we conclude we wouldn’t integrate. GCS suggests they’re not sedated.
Nate 16/4/24 via whatsapp

### Results export

