---
title: "Sodium GCS Mediation Mortality"
author: Miguel Ángel Armengol de la Hoz
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output:
  # html_notebook:
    code_folding: hide
    number_sections: yes
    theme: flatly
    toc: yes
    toc_float: yes

knit: (function(inputFile, encoding) { rmarkdown::render(inputFile, encoding = encoding, output_file = paste0(substr(inputFile,1,nchar(inputFile)-4)," ",Sys.Date(),'.html')) })
---

# Environment

```{r message=FALSE, warning=FALSE}
library(tidyverse)
```
# Subanalysis 1, Sedation Score

## Creating subanalises dataframes

```{r}
df_sub_sedated<-selected_df_imputed%>%filter(sedation_status=='Sedated by Med')
df_sub_notsedated<-selected_df_imputed%>%filter(sedation_status=='Not sedated by Med')
df_sub_susp_error<-selected_df_imputed%>%filter(sedation_status=='Suspected error')
```

## Creating categorical dataframes for subanalyses

### Sedated by Med Group

```{r}
df_sub_sedated$sodium_cat<-as.numeric(as.character(df_sub_sedated$sodium_cat))
sodium_cat_1_sedated_df<-df_sub_sedated%>%filter(sodium_cat%in%c(1,5))
sodium_cat_2_sedated_df<-df_sub_sedated%>%filter(sodium_cat%in%c(2,5))
sodium_cat_3_sedated_df<-df_sub_sedated%>%filter(sodium_cat%in%c(3,5))
sodium_cat_4_sedated_df<-df_sub_sedated%>%filter(sodium_cat%in%c(4,5))
sodium_cat_6_sedated_df<-df_sub_sedated%>%filter(sodium_cat%in%c(6,5))
sodium_cat_7_sedated_df<-df_sub_sedated%>%filter(sodium_cat%in%c(7,5))
sodium_cat_8_sedated_df<-df_sub_sedated%>%filter(sodium_cat%in%c(8,5))
```

### Non Sedated by Med Group

```{r}
df_sub_notsedated$sodium_cat<-as.numeric(as.character(df_sub_notsedated$sodium_cat))
sodium_cat_1_notsedated_df<-df_sub_notsedated%>%filter(sodium_cat%in%c(1,5))
sodium_cat_2_notsedated_df<-df_sub_notsedated%>%filter(sodium_cat%in%c(2,5))
sodium_cat_3_notsedated_df<-df_sub_notsedated%>%filter(sodium_cat%in%c(3,5))
sodium_cat_4_notsedated_df<-df_sub_notsedated%>%filter(sodium_cat%in%c(4,5))
sodium_cat_6_notsedated_df<-df_sub_notsedated%>%filter(sodium_cat%in%c(6,5))
sodium_cat_7_notsedated_df<-df_sub_notsedated%>%filter(sodium_cat%in%c(7,5))
sodium_cat_8_notsedated_df<-df_sub_notsedated%>%filter(sodium_cat%in%c(8,5))
```

# Subanalysis 2, Surgery Type

## Creating subanalises dataframes

```{r}
df_sub_electivesurgery<-selected_df_imputed%>%filter(electivesurgery==1)
df_sub_not_electivesurgery<-selected_df_imputed%>%filter(electivesurgery==0)
```


Getting ready the table I use for generating the plots but without generating the plots

The mediation estimate.
The total effect OR.
The total effect OR confidence interval.
Maybe also P value?
Columns: or    or.lb    or.u , p, Prop. Mediated (average)

## Creating categorical dataframes for subanalyses

### Elective Surgery Group

```{r}
df_sub_electivesurgery$sodium_cat <- as.numeric(as.character(df_sub_electivesurgery$sodium_cat))
sodium_cat_1_electivesurgery_df <- df_sub_electivesurgery %>% filter(sodium_cat %in% c(1, 5))
sodium_cat_2_electivesurgery_df <- df_sub_electivesurgery %>% filter(sodium_cat %in% c(2, 5))
sodium_cat_3_electivesurgery_df <- df_sub_electivesurgery %>% filter(sodium_cat %in% c(3, 5))
sodium_cat_4_electivesurgery_df <- df_sub_electivesurgery %>% filter(sodium_cat %in% c(4, 5))
sodium_cat_6_electivesurgery_df <- df_sub_electivesurgery %>% filter(sodium_cat %in% c(6, 5))
sodium_cat_7_electivesurgery_df <- df_sub_electivesurgery %>% filter(sodium_cat %in% c(7, 5))
sodium_cat_8_electivesurgery_df <- df_sub_electivesurgery %>% filter(sodium_cat %in% c(8, 5))
```

### Non-Elective Surgery Group

```{r}
df_sub_not_electivesurgery$sodium_cat <- as.numeric(as.character(df_sub_not_electivesurgery$sodium_cat))
sodium_cat_1_not_electivesurgery_df <- df_sub_not_electivesurgery %>% filter(sodium_cat %in% c(1, 5))
sodium_cat_2_not_electivesurgery_df <- df_sub_not_electivesurgery %>% filter(sodium_cat %in% c(2, 5))
sodium_cat_3_not_electivesurgery_df <- df_sub_not_electivesurgery %>% filter(sodium_cat %in% c(3, 5))
sodium_cat_4_not_electivesurgery_df <- df_sub_not_electivesurgery %>% filter(sodium_cat %in% c(4, 5))
sodium_cat_6_not_electivesurgery_df <- df_sub_not_electivesurgery %>% filter(sodium_cat %in% c(6, 5))
sodium_cat_7_not_electivesurgery_df <- df_sub_not_electivesurgery %>% filter(sodium_cat %in% c(7, 5))
sodium_cat_8_not_electivesurgery_df <- df_sub_not_electivesurgery %>% filter(sodium_cat %in% c(8, 5))
```

# Subanalysis 3, ICU Type

## Creating subanalises dataframes

```{r}
# Subset for each ICU type
icu_types <- c("Cardiac ICU", "CCU-CTICU", "CSICU+CTICU", "Med-Surg ICU", "MICU", "Neuro ICU", "SICU")
icu_names <- c("icu_a", "icu_b", "icu_c", "icu_d", "icu_e", "icu_f", "icu_g")

df_list <- list()
for (i in 1:length(icu_types)) {
  df_list[[icu_names[i]]] <- subset(selected_df_imputed, unittype == icu_types[i])
}
```

## Creating categorical dataframes for subanalyses

```{r}
# Convert sodium_cat to numeric
for (icu in icu_names) {
  df_list[[icu]]$sodium_cat <- as.numeric(as.character(df_list[[icu]]$sodium_cat))
}

# Define function to create sodium_cat categories
create_sodium_cat_df <- function(df, sodium_cat_value) {
  return(df %>% filter(sodium_cat %in% c(sodium_cat_value, 5)))
}

# Apply the function to create categories for each ICU
sodium_cat_values <- c(1, 2, 3, 4, 6, 7, 8)
for (icu in icu_names) {
  for (value in sodium_cat_values) {
    assign(paste0("sodium_cat_", value, "_", icu, "_df"), create_sodium_cat_df(df_list[[icu]], value))
  }
}
```

# Subanalysis 4, Mechanically ventilated

## Creating subanalises dataframes

```{r}
df_sub_mechvent <- selected_df_imputed %>% filter(mechvent_day01 == 1)
df_sub_not_mechvent <- selected_df_imputed %>% filter(mechvent_day01 == 0)
```

## Creating categorical dataframes for subanalyses

### Mechanically Ventilated Group

```{r}
df_sub_mechvent$sodium_cat <- as.numeric(as.character(df_sub_mechvent$sodium_cat))

sodium_cat_1_mechvent_df <- df_sub_mechvent %>% filter(sodium_cat %in% c(1, 5))
sodium_cat_2_mechvent_df <- df_sub_mechvent %>% filter(sodium_cat %in% c(2, 5))
sodium_cat_3_mechvent_df <- df_sub_mechvent %>% filter(sodium_cat %in% c(3, 5))
sodium_cat_4_mechvent_df <- df_sub_mechvent %>% filter(sodium_cat %in% c(4, 5))
sodium_cat_6_mechvent_df <- df_sub_mechvent %>% filter(sodium_cat %in% c(6, 5))
sodium_cat_7_mechvent_df <- df_sub_mechvent %>% filter(sodium_cat %in% c(7, 5))
sodium_cat_8_mechvent_df <- df_sub_mechvent %>% filter(sodium_cat %in% c(8, 5))
```

### Non-Mechanically Ventilated Group

```{r}
df_sub_not_mechvent$sodium_cat <- as.numeric(as.character(df_sub_not_mechvent$sodium_cat))

sodium_cat_1_not_mechvent_df <- df_sub_not_mechvent %>% filter(sodium_cat %in% c(1, 5))
sodium_cat_2_not_mechvent_df <- df_sub_not_mechvent %>% filter(sodium_cat %in% c(2, 5))
sodium_cat_3_not_mechvent_df <- df_sub_not_mechvent %>% filter(sodium_cat %in% c(3, 5))
sodium_cat_4_not_mechvent_df <- df_sub_not_mechvent %>% filter(sodium_cat %in% c(4, 5))
sodium_cat_6_not_mechvent_df <- df_sub_not_mechvent %>% filter(sodium_cat %in% c(6, 5))
sodium_cat_7_not_mechvent_df <- df_sub_not_mechvent %>% filter(sodium_cat %in% c(7, 5))
sodium_cat_8_not_mechvent_df <- df_sub_not_mechvent %>% filter(sodium_cat %in% c(8, 5))
```


