---
title: "Sodium GCS Mediation Mortality"
author: Miguel Ángel Armengol de la Hoz
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output:
  html_notebook:
    code_folding: hide
    number_sections: yes
    theme: flatly
    toc: yes
    toc_float: yes

knit: (function(inputFile, encoding) { rmarkdown::render(inputFile, encoding = encoding, output_file = paste0(substr(inputFile,1,nchar(inputFile)-4)," ",Sys.Date(),'.html')) })
---

# Environment

```{r message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)
library(mediation)
library(mgcv)
library(plotly)
library(voxel)
library(htmlwidgets)
library(readr)
library(bigrquery)
winsorize_x = function(x, cut = 0.05){
  cut_point_top <- quantile(x, 1 - cut, na.rm = T)
  cut_point_bottom <- quantile(x, cut, na.rm = T)
  i = which(x >= cut_point_top) 
  x[i] = cut_point_top
  j = which(x <= cut_point_bottom) 
  x[j] = cut_point_bottom
  return(x)
}

'%!in%' <- function(x,y)!('%in%'(x,y))

```

# Data Load

## Set up BigQuery related functions

This chunks also creates the run_query and get_sql function.

```{r setup, include=FALSE}
# Updated for our year
project_id <- "hst-953-2018"
options(httr_oauth_cache=FALSE)
# Function that takes in a sql command and runs it on bigquery
run_query <- function(query){
  data <- query_exec(query, project=project_id, use_legacy_sql=FALSE,max_pages = Inf)
  return(data)
}

# function for reading sql files
getSQL <- function(filepath){
  con = file(filepath, "r")
  sql.string <- ""

  while (TRUE){
    line <- readLines(con, n = 1)

    if ( length(line) == 0 ){
      break
    }

    line <- gsub("\\t", " ", line)

    if(grepl("--",line) == TRUE){
      line <- paste(sub("--","/*",line),"*/")
    }

    sql.string <- paste(sql.string, line)
  }

  close(con)
  return(sql.string)
}

```

## SQL Extraction

### Apache

The _apacheapsvar_ table contains the variables used to calculate the Acute Physiology Score (APS) III for patients. APS-III is an established method of summarizing patient severity of illness on admission to the ICU.

The score is part of the Acute Physiology Age Chronic Health Evaluation (APACHE) system of equations for predicting outcomes for ICU patients. See: http://eicu-crd.mit.edu/eicutables/apacheApsVar/

The _apachepredvar_ table provides variables underlying the APACHE predictions. Acute Physiology Age Chronic Health Evaluation (APACHE) consists of a groups of equations used for predicting outcomes in critically ill patients. See: http://eicu-crd.mit.edu/eicutables/apachePredVar/

We are extracting all variables needed to calculate apache except sodium.

```{r}
apacheapsvar <- run_query(getSQL("sql/apacheapsvar.sql" ))
apachepredvar <- run_query(getSQL("sql/apachepredvar.sql" ))
selected_columns<-names(apachepredvar[names(apachepredvar) %!in% names(apacheapsvar)])
selected_columns<-c('patientunitstayid',selected_columns)

allapachevars<-inner_join(apacheapsvar
                         ,apachepredvar[,selected_columns]
                        ,by='patientunitstayid')
```

### Other variables

```{r}
basic_demographics <- run_query(getSQL("sql/basic_demographics.sql" ))
gcs_baseline <- run_query(getSQL("sql/gcs_baseline.sql" ))
admission_sodium <- run_query(getSQL("sql/admission_sodium.sql" ))
charlson_score <- run_query(getSQL("sql/charlson_score.sql" ))
charlson_score<-charlson_score%>%select(patientunitstayid,charlson_liver1 ,charlson_chf1)
ventilated_binary_first_day <- run_query(getSQL("sql/ventilated_binary_first_day.sql" ))
apache_groups <- run_query(getSQL("sql/apache_groups.sql" ))

```

## Excel Import

We are not using this anymore

```{r eval=FALSE, include=FALSE}
john_dataset <- read_csv("../data/Final dataset 96,590of Sodium concentration, consciousness, and mortality in critical illness 2-4-20 for miguel.csv")

# the provided hospital mortality is not the valid one.
john_dataset$hospmortality<-NULL
john_dataset$mechvent_day01<-as.character( john_dataset$mechvent_day01)
```

# Inner Join

```{r}
initial_dataset<-Reduce(function(...) merge(..., all.x=TRUE), list(
   basic_demographics
  ,gcs_baseline
  ,admission_sodium
  ,charlson_score
  ,ventilated_binary_first_day
  ,apache_groups
))
```


# Addresing variables distribution

## Selecting variables

```{r}
selected_df<-initial_dataset%>%select(patientunitstayid,sodium1,gcs_baseline,apache_iv,age,gender,ethnicity,unittype  , charlson_liver1 ,charlson_chf1,mechvent_day01,apachedxgroup)
```

# Exclusion criteria

We are not excluding patients with a baseline sodium outside of our groups.

```{r}
selected_df<- selected_df[complete.cases(selected_df),]
```

# New variables creation

```{r}
selected_df$normal_gcs<-if_else( selected_df$gcs_baseline==15,1,0)
```

# Addressing outliers

We are applying the mean to sodium1 outliers values.

```{r}
# Median imputation for outliers
outliers <- boxplot(selected_df$sodium1, plot=FALSE)$out
selected_df$sodium1[which(selected_df$sodium1 %in% outliers)]<-median(selected_df$sodium1)
```

## Studying initial variables distribution

```{r}
par(mfrow=c(1,2))
for (i in 1:ncol( selected_df)) {
colname<-names( selected_df[i])  
ifelse( nrow(unique( selected_df[i]))<10  
        ,barplot(table( selected_df[i]),main=colname,xlab=colname)
        ,hist(as.numeric(unlist( selected_df[,i]))
        ,main = paste("Histogram of" ,colname),xlab=colname)
         )
cat('Summary of ',colname,'\n',sep = '')  
cat(summary( selected_df[i]),'\n',sep = '|')
}
```


