---
title: "Sodium GCS Mediation Mortality (SubAnalysis)"
author: "Miguel, Sicheng"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output:
  # html_notebook:
    code_folding: hide
    number_sections: yes
    theme: flatly
    toc: yes
    toc_float: yes
---

# Environment

```{r message=FALSE, warning=FALSE,echo = F}
library(dplyr)
library(ggplot2)
library(mediation)
library(mgcv)
library(readr)
library(stringi)
library(mediation)
library(glmnet)
library(data.table)
library(cowplot)
library(causalweight)
library(logger)

'%!in%' <- function(x,y)!('%in%'(x,y))
extract_mediation_summary <- function (x) { 
  clp <- 100 * x$conf.level
  isLinear.y <- ((class(x$model.y)[1] %in% c("lm", "rq")) || 
                   (inherits(x$model.y, "glm") && x$model.y$family$family == 
                      "gaussian" && x$model.y$family$link == "identity") || 
                   (inherits(x$model.y, "survreg") && x$model.y$dist == 
                      "gaussian"))
  printone <- !x$INT && isLinear.y
  if (printone) {
    smat <- c(x$d1, x$d1.ci, x$d1.p)
    smat <- rbind(smat, c(x$z0, x$z0.ci, x$z0.p))
    smat <- rbind(smat, c(x$tau.coef, x$tau.ci, x$tau.p))
    smat <- rbind(smat, c(x$n0, x$n0.ci, x$n0.p))
    rownames(smat) <- c("ACME", "ADE", "Total Effect", "Prop. Mediated")
  } else {
    smat <- c(x$d0, x$d0.ci, x$d0.p)
    smat <- rbind(smat, c(x$d1, x$d1.ci, x$d1.p))
    smat <- rbind(smat, c(x$z0, x$z0.ci, x$z0.p))
    smat <- rbind(smat, c(x$z1, x$z1.ci, x$z1.p))
    smat <- rbind(smat, c(x$tau.coef, x$tau.ci, x$tau.p))
    smat <- rbind(smat, c(x$n0, x$n0.ci, x$n0.p))
    smat <- rbind(smat, c(x$n1, x$n1.ci, x$n1.p))
    smat <- rbind(smat, c(x$d.avg, x$d.avg.ci, x$d.avg.p))
    smat <- rbind(smat, c(x$z.avg, x$z.avg.ci, x$z.avg.p))
    smat <- rbind(smat, c(x$n.avg, x$n.avg.ci, x$n.avg.p))
    rownames(smat) <- c("ACME (control)", "ACME (treated)", 
                        "ADE (control)", "ADE (treated)", "Total Effect", 
                        "Prop. Mediated (control)", "Prop. Mediated (treated)", 
                        "ACME (average)", "ADE (average)", "Prop. Mediated (average)")
  }
  colnames(smat) <- c("Estimate", paste(clp, "% CI Lower", sep = ""), 
                      paste(clp, "% CI Upper", sep = ""), "p-value")
  smat
}
```

# SubAnalysis 1, Sedation Score

## SubAnalysis 1A: Sedated by Med SubAnalysis

### Mediation package

#### Baseline sodium_cat_1

##### Mediation Analysis

```{r}
total_effect_sedated<-NULL
sodium_cat_1_sedated_df$sodium_cat<-as.factor(sodium_cat_1_sedated_df$sodium_cat)
sodium_cat_1_sedated_df$sodium_cat<-relevel(sodium_cat_1_sedated_df$sodium_cat,ref = "5")
fit_totaleffect<-glm(as.factor(hosp_mortality)~sodium_cat+ AP_grp + ethnicity + gender + age + MAP + temperature + heartrate + respiratoryrate + mechvent_day01 + day1fio2 + day1pao2 + pco2 + ph + urine + creatinine + albumin + bilirubin + wbc + cirrhosis + aids + leukemia + immunosuppression + charlson_chf1 + charlson_liver1

                   , sodium_cat_1_sedated_df
,family = binomial)
summ<-summary(fit_totaleffect)
total_effect_sedated<-rbind(total_effect_sedated,data.frame(cat = 1,estimated = summ$coefficients[2],se = summ$coefficients[2+length(summ$coefficients)/4],p = summ$coefficients[2+3*length(summ$coefficients)/4]))



fit_mediator<-lm(gcs_baseline ~ sodium_cat+AP_grp + ethnicity + gender + age + MAP + temperature + heartrate + respiratoryrate + mechvent_day01 + day1fio2 + day1pao2 + pco2 + ph + urine + creatinine + albumin + bilirubin + wbc + cirrhosis + aids + leukemia + immunosuppression + charlson_chf1 + charlson_liver1,
data =  sodium_cat_1_sedated_df)

fit_dv<-glm(as.factor(hosp_mortality)~gcs_baseline+sodium_cat+ AP_grp + ethnicity + gender + age + MAP + temperature + heartrate + respiratoryrate + mechvent_day01 + day1fio2 + day1pao2 + pco2 + ph + urine + creatinine + albumin + bilirubin + wbc + cirrhosis + aids + leukemia + immunosuppression + charlson_chf1 + charlson_liver1
                   , sodium_cat_1_sedated_df
           ,family = binomial("probit"))

summary(fit_dv)

#treat is the independend variable

mediation_results_cat1_sedated <- mediate(fit_mediator, fit_dv, 
                                  treat="sodium_cat", mediator="gcs_baseline",control.value = 5,treat.value = 1,
                                  sims = 1000)
mediation_results_cat1_sedated_sum<-as.data.frame(extract_mediation_summary(mediation_results_cat1_sedated))
mediation_results_cat1_sedated_sum<-mediation_results_cat1_sedated_sum[c(5,8,9,10),]

#replace effect of treated group with average effect
#mediation_results_cat1_sedated_sum<-mediation_results_cat1_sedated_sum[grepl("treated|Total Effect", rownames(mediation_results_cat1_sedated_sum)), ]#

mediation_results_cat1_sedated_sum
```

##### Sensitivity Analysis

```{r}
sens.out_sedated1 <- medsens(mediation_results_cat1_sedated, rho.by = 0.1, effect.type = "indirect", sims = 200)
log_info()
```

#### Baseline sodium_cat_2

##### Mediation Analysis

```{r}
sodium_cat_2_sedated_df$sodium_cat<-as.factor(sodium_cat_2_sedated_df$sodium_cat)
sodium_cat_2_sedated_df$sodium_cat<-relevel(sodium_cat_2_sedated_df$sodium_cat,ref = "5")
fit_totaleffect<-glm(as.factor(hosp_mortality)~sodium_cat + AP_grp + ethnicity + gender + age + MAP + temperature + heartrate + respiratoryrate + mechvent_day01 + day1fio2 + day1pao2 + pco2 + ph + urine + creatinine + albumin + bilirubin + wbc + cirrhosis + aids + leukemia + immunosuppression + charlson_chf1 + charlson_liver1

                   , sodium_cat_2_sedated_df
,family = binomial)
summ<-summary(fit_totaleffect)
total_effect_sedated<-rbind(total_effect_sedated,data.frame(cat = 2,estimated = summ$coefficients[2],se = summ$coefficients[2+length(summ$coefficients)/4],p = summ$coefficients[2+3*length(summ$coefficients)/4]))



fit_mediator<-lm(gcs_baseline ~ sodium_cat + AP_grp + ethnicity + gender + age + MAP + temperature + heartrate + respiratoryrate + mechvent_day01 + day1fio2 + day1pao2 + pco2 + ph + urine + creatinine + albumin + bilirubin + wbc + cirrhosis + aids + leukemia + immunosuppression + charlson_chf1 + charlson_liver1,
data =  sodium_cat_2_sedated_df)
fit_dv<-glm(as.factor(hosp_mortality) ~gcs_baseline+sodium_cat+AP_grp + ethnicity + gender + age + MAP + temperature + heartrate + respiratoryrate + mechvent_day01 + day1fio2 + day1pao2 + pco2 + ph + urine + creatinine + albumin + bilirubin + wbc + cirrhosis + aids + leukemia + immunosuppression + charlson_chf1 + charlson_liver1
                   , sodium_cat_2_sedated_df
           ,family = binomial("probit"))
summary(fit_dv)
#treat is the independend variable
mediation_results_cat2_sedated <- mediate(fit_mediator, fit_dv, 
                                  treat="sodium_cat", mediator="gcs_baseline",control.value = 5,treat.value = 2,
                                  sims = 1000)
mediation_results_cat2_sedated_sum<-as.data.frame(extract_mediation_summary(mediation_results_cat2_sedated))
mediation_results_cat2_sedated_sum<-mediation_results_cat2_sedated_sum[c(5,8,9,10),]

mediation_results_cat2_sedated_sum
```

##### Sensitivity Analysis

```{r}
sens.out_sedated2 <- medsens(mediation_results_cat2_sedated, rho.by = 0.1, effect.type = "indirect", sims = 200)
log_info()
```

#### Baseline sodium_cat_3

##### Mediation Analysis

```{r}
sodium_cat_3_sedated_df$sodium_cat<-as.factor(sodium_cat_3_sedated_df$sodium_cat)
sodium_cat_3_sedated_df$sodium_cat<-relevel(sodium_cat_3_sedated_df$sodium_cat,ref = "5")
fit_totaleffect<-glm(as.factor(hosp_mortality)~sodium_cat+age+AP_grp+gender+ethnicity+unittype+charlson_liver1 +charlson_chf1+mechvent_day01+respiratoryrate+MAP

                   , sodium_cat_3_sedated_df
,family = binomial)
summ<-summary(fit_totaleffect)
total_effect_sedated<-rbind(total_effect_sedated,data.frame(cat = 3,estimated = summ$coefficients[2],se = summ$coefficients[2+length(summ$coefficients)/4],p = summ$coefficients[2+3*length(summ$coefficients)/4]))

fit_mediator<-lm(gcs_baseline ~ sodium_cat+AP_grp + ethnicity + gender + age + MAP + temperature + heartrate + respiratoryrate + mechvent_day01 + day1fio2 + day1pao2 + pco2 + ph + urine + creatinine + albumin + bilirubin + wbc + cirrhosis + aids + leukemia + immunosuppression + charlson_chf1 + charlson_liver1,
data =  sodium_cat_3_sedated_df)
fit_dv<-glm(as.factor(hosp_mortality) ~gcs_baseline+sodium_cat+AP_grp + ethnicity + gender + age + MAP + temperature + heartrate + respiratoryrate + mechvent_day01 + day1fio2 + day1pao2 + pco2 + ph + urine + creatinine + albumin + bilirubin + wbc + cirrhosis + aids + leukemia + immunosuppression + charlson_chf1 + charlson_liver1
                   , sodium_cat_3_sedated_df
           ,family = binomial("probit"))
summary(fit_dv)
#treat is the independend variable
mediation_results_cat3_sedated <- mediate(fit_mediator, fit_dv, 
                                  treat="sodium_cat", mediator="gcs_baseline",control.value = 5,treat.value = 3,
                                  sims = 1000)
mediation_results_cat3_sedated_sum<-as.data.frame(extract_mediation_summary(mediation_results_cat3_sedated))
mediation_results_cat3_sedated_sum<-mediation_results_cat3_sedated_sum[c(5,8,9,10),]

mediation_results_cat3_sedated_sum
```

##### Sensitivity Analysis

```{r}
sens.out_sedated3 <- medsens(mediation_results_cat3_sedated, rho.by = 0.1, effect.type = "indirect", sims = 200)
```

#### Baseline sodium_cat_4

##### Mediation Analysis

```{r}
sodium_cat_4_sedated_df$sodium_cat<-as.factor(sodium_cat_4_sedated_df$sodium_cat)
sodium_cat_4_sedated_df$sodium_cat<-relevel(sodium_cat_4_sedated_df$sodium_cat,ref = "5")
fit_totaleffect<-glm(as.factor(hosp_mortality)~sodium_cat+AP_grp + ethnicity + gender + age + MAP + temperature + heartrate + respiratoryrate + mechvent_day01 + day1fio2 + day1pao2 + pco2 + ph + urine + creatinine + albumin + bilirubin + wbc + cirrhosis + aids + leukemia + immunosuppression + charlson_chf1 + charlson_liver1
                   , sodium_cat_4_sedated_df
,family = binomial)
summ<-summary(fit_totaleffect)
total_effect_sedated<-rbind(total_effect_sedated,data.frame(cat = 4,estimated = summ$coefficients[2],se = summ$coefficients[2+length(summ$coefficients)/4],p = summ$coefficients[2+3*length(summ$coefficients)/4]))


fit_mediator<-lm(gcs_baseline ~ sodium_cat+AP_grp + ethnicity + gender + age + MAP + temperature + heartrate + respiratoryrate + mechvent_day01 + day1fio2 + day1pao2 + pco2 + ph + urine + creatinine + albumin + bilirubin + wbc + cirrhosis + aids + leukemia + immunosuppression + charlson_chf1 + charlson_liver1,
data =  sodium_cat_4_sedated_df)
fit_dv<-glm(as.factor(hosp_mortality) ~gcs_baseline+sodium_cat+AP_grp + ethnicity + gender + age + MAP + temperature + heartrate + respiratoryrate + mechvent_day01 + day1fio2 + day1pao2 + pco2 + ph + urine + creatinine + albumin + bilirubin + wbc + cirrhosis + aids + leukemia + immunosuppression + charlson_chf1 + charlson_liver1
                   , sodium_cat_4_sedated_df
           ,family = binomial("probit"))
summary(fit_dv)
#treat is the independend variable
mediation_results_cat4_sedated <- mediate(fit_mediator, fit_dv, 
                                  treat="sodium_cat", mediator="gcs_baseline",control.value = 5,treat.value = 4,
                                  sims = 1000)
mediation_results_cat4_sedated_sum<-as.data.frame(extract_mediation_summary(mediation_results_cat4_sedated))
mediation_results_cat4_sedated_sum<-mediation_results_cat4_sedated_sum[c(5,8,9,10),]
mediation_results_cat4_sedated_sum

```

##### Sensitivity Analysis

```{r}
sens.out_sedated4 <- medsens(mediation_results_cat4_sedated, rho.by = 0.1, effect.type = "indirect", sims = 200)
```
#### Baseline sodium_cat_6

##### Mediation Analysis

```{r}
sodium_cat_6_sedated_df$sodium_cat<-as.factor(sodium_cat_6_sedated_df$sodium_cat)
sodium_cat_6_sedated_df$sodium_cat<-relevel(sodium_cat_6_sedated_df$sodium_cat,ref = "5")
fit_totaleffect<-glm(as.factor(hosp_mortality)~sodium_cat+AP_grp + ethnicity + gender + age + MAP + temperature + heartrate + respiratoryrate + mechvent_day01 + day1fio2 + day1pao2 + pco2 + ph + urine + creatinine + albumin + bilirubin + wbc + cirrhosis + aids + leukemia + immunosuppression + charlson_chf1 + charlson_liver1

                   , sodium_cat_6_sedated_df
,family = binomial)
summ<-summary(fit_totaleffect)
total_effect_sedated<-rbind(total_effect_sedated,data.frame(cat = 6,estimated = summ$coefficients[2],se = summ$coefficients[2+length(summ$coefficients)/4],p = summ$coefficients[2+3*length(summ$coefficients)/4]))


fit_mediator<-lm(gcs_baseline ~ sodium_cat+AP_grp + ethnicity + gender + age + MAP + temperature + heartrate + respiratoryrate + mechvent_day01 + day1fio2 + day1pao2 + pco2 + ph + urine + creatinine + albumin + bilirubin + wbc + cirrhosis + aids + leukemia + immunosuppression + charlson_chf1 + charlson_liver1,
data =  sodium_cat_6_sedated_df)
fit_dv<-glm(as.factor(hosp_mortality) ~gcs_baseline+sodium_cat+AP_grp + ethnicity + gender + age + MAP + temperature + heartrate + respiratoryrate + mechvent_day01 + day1fio2 + day1pao2 + pco2 + ph + urine + creatinine + albumin + bilirubin + wbc + cirrhosis + aids + leukemia + immunosuppression + charlson_chf1 + charlson_liver1
                   , sodium_cat_6_sedated_df
           ,family = binomial("probit"))
summary(fit_dv)
#treat is the independend variable
mediation_results_cat6_sedated <- mediate(fit_mediator, fit_dv, 
                                  treat="sodium_cat", mediator="gcs_baseline",control.value = 5,treat.value = 6,
                                  sims = 1000)
mediation_results_cat6_sedated_sum<-as.data.frame(extract_mediation_summary(mediation_results_cat6_sedated))
mediation_results_cat6_sedated_sum<-mediation_results_cat6_sedated_sum[c(5,8,9,10),]
mediation_results_cat6_sedated_sum
```

##### Sensitivity Analysis

```{r}
sens.out_sedated6 <- medsens(mediation_results_cat6_sedated, rho.by = 0.1, effect.type = "indirect", sims = 200)
```

#### Baseline sodium_cat_7

##### Mediation Analysis

```{r}
sodium_cat_7_sedated_df$sodium_cat<-as.factor(sodium_cat_7_sedated_df$sodium_cat)
sodium_cat_7_sedated_df$sodium_cat<-relevel(sodium_cat_7_sedated_df$sodium_cat,ref = "5")
fit_totaleffect<-glm(as.factor(hosp_mortality)~sodium_cat+AP_grp + ethnicity + gender + age + MAP + temperature + heartrate + respiratoryrate + mechvent_day01 + day1fio2 + day1pao2 + pco2 + ph + urine + creatinine + albumin + bilirubin + wbc + cirrhosis + aids + leukemia + immunosuppression + charlson_chf1 + charlson_liver1

                   , sodium_cat_7_sedated_df
,family = binomial)
summ<-summary(fit_totaleffect)
total_effect_sedated<-rbind(total_effect_sedated,data.frame(cat = 7,estimated = summ$coefficients[2],se = summ$coefficients[2+length(summ$coefficients)/4],p = summ$coefficients[2+3*length(summ$coefficients)/4]))


fit_mediator<-lm(gcs_baseline ~ sodium_cat + AP_grp + ethnicity + gender + age + MAP + temperature + heartrate + respiratoryrate + mechvent_day01 + day1fio2 + day1pao2 + pco2 + ph + urine + creatinine + albumin + bilirubin + wbc + cirrhosis + aids + leukemia + immunosuppression + charlson_chf1 + charlson_liver1,
data =  sodium_cat_7_sedated_df)
fit_dv<-glm(as.factor(hosp_mortality) ~gcs_baseline+sodium_cat + AP_grp + ethnicity + gender + age + MAP + temperature + heartrate + respiratoryrate + mechvent_day01 + day1fio2 + day1pao2 + pco2 + ph + urine + creatinine + albumin + bilirubin + wbc + cirrhosis + aids + leukemia + immunosuppression + charlson_chf1 + charlson_liver1
                   , sodium_cat_7_sedated_df
           ,family = binomial("probit"))
summary(fit_dv)
#treat is the independend variable
mediation_results_cat7_sedated <- mediate(fit_mediator, fit_dv, 
                                  treat="sodium_cat", mediator="gcs_baseline",control.value = 5,treat.value = 7,
                                  sims = 1000)
mediation_results_cat7_sedated_sum<-as.data.frame(extract_mediation_summary(mediation_results_cat7_sedated))
mediation_results_cat7_sedated_sum<-mediation_results_cat7_sedated_sum[c(5,8,9,10),]
mediation_results_cat7_sedated_sum
```

##### Sensitivity Analysis

```{r}
sens.out_sedated7 <- medsens(mediation_results_cat7_sedated, rho.by = 0.1, effect.type = "indirect", sims = 200)
```

#### Baseline sodium_cat_8

##### Mediation Analysis

```{r}
sodium_cat_8_sedated_df$sodium_cat<-as.factor(sodium_cat_8_sedated_df$sodium_cat)
sodium_cat_8_sedated_df$sodium_cat<-relevel(sodium_cat_8_sedated_df$sodium_cat,ref = "5")
fit_totaleffect<-glm(as.factor(hosp_mortality)~sodium_cat + AP_grp + ethnicity + gender + age + MAP + temperature + heartrate + respiratoryrate + mechvent_day01 + day1fio2 + day1pao2 + pco2 + ph + urine + creatinine + albumin + bilirubin + wbc + cirrhosis + aids + leukemia + immunosuppression + charlson_chf1 + charlson_liver1
                   , sodium_cat_8_sedated_df
,family = binomial)
summ<-summary(fit_totaleffect)
total_effect_sedated<-rbind(total_effect_sedated,data.frame(cat = 8,estimated = summ$coefficients[2],se = summ$coefficients[2+length(summ$coefficients)/4],p = summ$coefficients[2+3*length(summ$coefficients)/4]))


fit_mediator<-lm(gcs_baseline ~ sodium_cat + AP_grp + ethnicity + gender + age + MAP + temperature + heartrate + respiratoryrate + mechvent_day01 + day1fio2 + day1pao2 + pco2 + ph + urine + creatinine + albumin + bilirubin + wbc + cirrhosis + aids + leukemia + immunosuppression + charlson_chf1 + charlson_liver1,
data =  sodium_cat_8_sedated_df)
fit_dv<-glm(as.factor(hosp_mortality) ~gcs_baseline+sodium_cat + AP_grp + ethnicity + gender + age + MAP + temperature + heartrate + respiratoryrate + mechvent_day01 + day1fio2 + day1pao2 + pco2 + ph + urine + creatinine + albumin + bilirubin + wbc + cirrhosis + aids + leukemia + immunosuppression + charlson_chf1 + charlson_liver1
                   , sodium_cat_8_sedated_df
           ,family = binomial("probit"))
summary(fit_dv)
#treat is the independend variable
mediation_results_cat8_sedated <- mediate(fit_mediator, fit_dv, 
                                  treat="sodium_cat", mediator="gcs_baseline",control.value = 5,treat.value = 8,
                                  sims = 1000)
mediation_results_cat8_sedated_sum<-as.data.frame(extract_mediation_summary(mediation_results_cat8_sedated))
mediation_results_cat8_sedated_sum<-mediation_results_cat8_sedated_sum[c(5,8,9,10),]
mediation_results_cat8_sedated_sum
```

##### Sensitivity Analysis

```{r}
sens.out_sedated8 <- medsens(mediation_results_cat8_sedated, rho.by = 0.1, effect.type = "indirect", sims = 200)
```

## SubAnalysis 1B: Not Sedated by Med SubAnalysis

### Mediation package

#### Baseline sodium_cat_1

##### Mediation Analysis

```{r}
total_effect_notsedated<-NULL
sodium_cat_1_notsedated_df$sodium_cat<-as.factor(sodium_cat_1_notsedated_df$sodium_cat)
sodium_cat_1_notsedated_df$sodium_cat<-relevel(sodium_cat_1_notsedated_df$sodium_cat,ref = "5")
fit_totaleffect<-glm(as.factor(hosp_mortality)~sodium_cat+ AP_grp + ethnicity + gender + age + MAP + temperature + heartrate + respiratoryrate + mechvent_day01 + day1fio2 + day1pao2 + pco2 + ph + urine + creatinine + albumin + bilirubin + wbc + cirrhosis + aids + leukemia + immunosuppression + charlson_chf1 + charlson_liver1

                   , sodium_cat_1_notsedated_df
,family = binomial)
summ<-summary(fit_totaleffect)
total_effect_notsedated<-rbind(total_effect_notsedated,data.frame(cat = 1,estimated = summ$coefficients[2],se = summ$coefficients[2+length(summ$coefficients)/4],p = summ$coefficients[2+3*length(summ$coefficients)/4]))



fit_mediator<-lm(gcs_baseline ~ sodium_cat+AP_grp + ethnicity + gender + age + MAP + temperature + heartrate + respiratoryrate + mechvent_day01 + day1fio2 + day1pao2 + pco2 + ph + urine + creatinine + albumin + bilirubin + wbc + cirrhosis + aids + leukemia + immunosuppression + charlson_chf1 + charlson_liver1,
data =  sodium_cat_1_notsedated_df)

fit_dv<-glm(as.factor(hosp_mortality)~gcs_baseline+sodium_cat+ AP_grp + ethnicity + gender + age + MAP + temperature + heartrate + respiratoryrate + mechvent_day01 + day1fio2 + day1pao2 + pco2 + ph + urine + creatinine + albumin + bilirubin + wbc + cirrhosis + aids + leukemia + immunosuppression + charlson_chf1 + charlson_liver1
                   , sodium_cat_1_notsedated_df
           ,family = binomial("probit"))

summary(fit_dv)

#treat is the independend variable

mediation_results_cat1_notsedated <- mediate(fit_mediator, fit_dv, 
                                  treat="sodium_cat", mediator="gcs_baseline",control.value = 5,treat.value = 1,
                                  sims = 1000)
mediation_results_cat1_notsedated_sum<-as.data.frame(extract_mediation_summary(mediation_results_cat1_notsedated))
mediation_results_cat1_notsedated_sum<-mediation_results_cat1_notsedated_sum[c(5,8,9,10),]

#replace effect of treated group with average effect
#mediation_results_cat1_notsedated_sum<-mediation_results_cat1_notsedated_sum[grepl("treated|Total Effect", rownames(mediation_results_cat1_notsedated_sum)), ]#

mediation_results_cat1_notsedated_sum
```

##### Sensitivity Analysis

```{r}
sens.out_notsedated1 <- medsens(mediation_results_cat1_notsedated, rho.by = 0.1, effect.type = "indirect", sims = 200)
```

#### Baseline sodium_cat_2

##### Mediation Analysis

```{r}
sodium_cat_2_notsedated_df$sodium_cat<-as.factor(sodium_cat_2_notsedated_df$sodium_cat)
sodium_cat_2_notsedated_df$sodium_cat<-relevel(sodium_cat_2_notsedated_df$sodium_cat,ref = "5")
fit_totaleffect<-glm(as.factor(hosp_mortality)~sodium_cat + AP_grp + ethnicity + gender + age + MAP + temperature + heartrate + respiratoryrate + mechvent_day01 + day1fio2 + day1pao2 + pco2 + ph + urine + creatinine + albumin + bilirubin + wbc + cirrhosis + aids + leukemia + immunosuppression + charlson_chf1 + charlson_liver1

                   , sodium_cat_2_notsedated_df
,family = binomial)
summ<-summary(fit_totaleffect)
total_effect_notsedated<-rbind(total_effect_notsedated,data.frame(cat = 2,estimated = summ$coefficients[2],se = summ$coefficients[2+length(summ$coefficients)/4],p = summ$coefficients[2+3*length(summ$coefficients)/4]))



fit_mediator<-lm(gcs_baseline ~ sodium_cat + AP_grp + ethnicity + gender + age + MAP + temperature + heartrate + respiratoryrate + mechvent_day01 + day1fio2 + day1pao2 + pco2 + ph + urine + creatinine + albumin + bilirubin + wbc + cirrhosis + aids + leukemia + immunosuppression + charlson_chf1 + charlson_liver1,
data =  sodium_cat_2_notsedated_df)
fit_dv<-glm(as.factor(hosp_mortality) ~gcs_baseline+sodium_cat+AP_grp + ethnicity + gender + age + MAP + temperature + heartrate + respiratoryrate + mechvent_day01 + day1fio2 + day1pao2 + pco2 + ph + urine + creatinine + albumin + bilirubin + wbc + cirrhosis + aids + leukemia + immunosuppression + charlson_chf1 + charlson_liver1
                   , sodium_cat_2_notsedated_df
           ,family = binomial("probit"))
summary(fit_dv)
#treat is the independend variable
mediation_results_cat2_notsedated <- mediate(fit_mediator, fit_dv, 
                                  treat="sodium_cat", mediator="gcs_baseline",control.value = 5,treat.value = 2,
                                  sims = 1000)
mediation_results_cat2_notsedated_sum<-as.data.frame(extract_mediation_summary(mediation_results_cat2_notsedated))
mediation_results_cat2_notsedated_sum<-mediation_results_cat2_notsedated_sum[c(5,8,9,10),]

mediation_results_cat2_notsedated_sum
```

##### Sensitivity Analysis

```{r}
sens.out_notsedated2 <- medsens(mediation_results_cat2_notsedated, rho.by = 0.1, effect.type = "indirect", sims = 200)
```

#### Baseline sodium_cat_3

##### Mediation Analysis

```{r}
sodium_cat_3_notsedated_df$sodium_cat<-as.factor(sodium_cat_3_notsedated_df$sodium_cat)
sodium_cat_3_notsedated_df$sodium_cat<-relevel(sodium_cat_3_notsedated_df$sodium_cat,ref = "5")
fit_totaleffect<-glm(as.factor(hosp_mortality)~sodium_cat+age+AP_grp+gender+ethnicity+unittype+charlson_liver1 +charlson_chf1+mechvent_day01+respiratoryrate+MAP

                   , sodium_cat_3_notsedated_df
,family = binomial)
summ<-summary(fit_totaleffect)
total_effect_notsedated<-rbind(total_effect_notsedated,data.frame(cat = 3,estimated = summ$coefficients[2],se = summ$coefficients[2+length(summ$coefficients)/4],p = summ$coefficients[2+3*length(summ$coefficients)/4]))

fit_mediator<-lm(gcs_baseline ~ sodium_cat+AP_grp + ethnicity + gender + age + MAP + temperature + heartrate + respiratoryrate + mechvent_day01 + day1fio2 + day1pao2 + pco2 + ph + urine + creatinine + albumin + bilirubin + wbc + cirrhosis + aids + leukemia + immunosuppression + charlson_chf1 + charlson_liver1,
data =  sodium_cat_3_notsedated_df)
fit_dv<-glm(as.factor(hosp_mortality) ~gcs_baseline+sodium_cat+AP_grp + ethnicity + gender + age + MAP + temperature + heartrate + respiratoryrate + mechvent_day01 + day1fio2 + day1pao2 + pco2 + ph + urine + creatinine + albumin + bilirubin + wbc + cirrhosis + aids + leukemia + immunosuppression + charlson_chf1 + charlson_liver1
                   , sodium_cat_3_notsedated_df
           ,family = binomial("probit"))
summary(fit_dv)
#treat is the independend variable
mediation_results_cat3_notsedated <- mediate(fit_mediator, fit_dv, 
                                  treat="sodium_cat", mediator="gcs_baseline",control.value = 5,treat.value = 3,
                                  sims = 1000)
mediation_results_cat3_notsedated_sum<-as.data.frame(extract_mediation_summary(mediation_results_cat3_notsedated))
mediation_results_cat3_notsedated_sum<-mediation_results_cat3_notsedated_sum[c(5,8,9,10),]

mediation_results_cat3_notsedated_sum
```

##### Sensitivity Analysis

```{r}
sens.out_notsedated3 <- medsens(mediation_results_cat3_notsedated, rho.by = 0.1, effect.type = "indirect", sims = 200)
```

#### Baseline sodium_cat_4

##### Mediation Analysis

```{r}
sodium_cat_4_notsedated_df$sodium_cat<-as.factor(sodium_cat_4_notsedated_df$sodium_cat)
sodium_cat_4_notsedated_df$sodium_cat<-relevel(sodium_cat_4_notsedated_df$sodium_cat,ref = "5")
fit_totaleffect<-glm(as.factor(hosp_mortality)~sodium_cat+AP_grp + ethnicity + gender + age + MAP + temperature + heartrate + respiratoryrate + mechvent_day01 + day1fio2 + day1pao2 + pco2 + ph + urine + creatinine + albumin + bilirubin + wbc + cirrhosis + aids + leukemia + immunosuppression + charlson_chf1 + charlson_liver1
                   , sodium_cat_4_notsedated_df
,family = binomial)
summ<-summary(fit_totaleffect)
total_effect_notsedated<-rbind(total_effect_notsedated,data.frame(cat = 4,estimated = summ$coefficients[2],se = summ$coefficients[2+length(summ$coefficients)/4],p = summ$coefficients[2+3*length(summ$coefficients)/4]))


fit_mediator<-lm(gcs_baseline ~ sodium_cat+AP_grp + ethnicity + gender + age + MAP + temperature + heartrate + respiratoryrate + mechvent_day01 + day1fio2 + day1pao2 + pco2 + ph + urine + creatinine + albumin + bilirubin + wbc + cirrhosis + aids + leukemia + immunosuppression + charlson_chf1 + charlson_liver1,
data =  sodium_cat_4_notsedated_df)
fit_dv<-glm(as.factor(hosp_mortality) ~gcs_baseline+sodium_cat+AP_grp + ethnicity + gender + age + MAP + temperature + heartrate + respiratoryrate + mechvent_day01 + day1fio2 + day1pao2 + pco2 + ph + urine + creatinine + albumin + bilirubin + wbc + cirrhosis + aids + leukemia + immunosuppression + charlson_chf1 + charlson_liver1
                   , sodium_cat_4_notsedated_df
           ,family = binomial("probit"))
summary(fit_dv)
#treat is the independend variable
mediation_results_cat4_notsedated <- mediate(fit_mediator, fit_dv, 
                                  treat="sodium_cat", mediator="gcs_baseline",control.value = 5,treat.value = 4,
                                  sims = 1000)
mediation_results_cat4_notsedated_sum<-as.data.frame(extract_mediation_summary(mediation_results_cat4_notsedated))
mediation_results_cat4_notsedated_sum<-mediation_results_cat4_notsedated_sum[c(5,8,9,10),]
mediation_results_cat4_notsedated_sum

```

##### Sensitivity Analysis

```{r}
sens.out_notsedated4 <- medsens(mediation_results_cat4_notsedated, rho.by = 0.1, effect.type = "indirect", sims = 200)
```
#### Baseline sodium_cat_6

##### Mediation Analysis

```{r}
sodium_cat_6_notsedated_df$sodium_cat<-as.factor(sodium_cat_6_notsedated_df$sodium_cat)
sodium_cat_6_notsedated_df$sodium_cat<-relevel(sodium_cat_6_notsedated_df$sodium_cat,ref = "5")
fit_totaleffect<-glm(as.factor(hosp_mortality)~sodium_cat+AP_grp + ethnicity + gender + age + MAP + temperature + heartrate + respiratoryrate + mechvent_day01 + day1fio2 + day1pao2 + pco2 + ph + urine + creatinine + albumin + bilirubin + wbc + cirrhosis + aids + leukemia + immunosuppression + charlson_chf1 + charlson_liver1

                   , sodium_cat_6_notsedated_df
,family = binomial)
summ<-summary(fit_totaleffect)
total_effect_notsedated<-rbind(total_effect_notsedated,data.frame(cat = 6,estimated = summ$coefficients[2],se = summ$coefficients[2+length(summ$coefficients)/4],p = summ$coefficients[2+3*length(summ$coefficients)/4]))


fit_mediator<-lm(gcs_baseline ~ sodium_cat+AP_grp + ethnicity + gender + age + MAP + temperature + heartrate + respiratoryrate + mechvent_day01 + day1fio2 + day1pao2 + pco2 + ph + urine + creatinine + albumin + bilirubin + wbc + cirrhosis + aids + leukemia + immunosuppression + charlson_chf1 + charlson_liver1,
data =  sodium_cat_6_notsedated_df)
fit_dv<-glm(as.factor(hosp_mortality) ~gcs_baseline+sodium_cat+AP_grp + ethnicity + gender + age + MAP + temperature + heartrate + respiratoryrate + mechvent_day01 + day1fio2 + day1pao2 + pco2 + ph + urine + creatinine + albumin + bilirubin + wbc + cirrhosis + aids + leukemia + immunosuppression + charlson_chf1 + charlson_liver1
                   , sodium_cat_6_notsedated_df
           ,family = binomial("probit"))
summary(fit_dv)
#treat is the independend variable
mediation_results_cat6_notsedated <- mediate(fit_mediator, fit_dv, 
                                  treat="sodium_cat", mediator="gcs_baseline",control.value = 5,treat.value = 6,
                                  sims = 1000)
mediation_results_cat6_notsedated_sum<-as.data.frame(extract_mediation_summary(mediation_results_cat6_notsedated))
mediation_results_cat6_notsedated_sum<-mediation_results_cat6_notsedated_sum[c(5,8,9,10),]
mediation_results_cat6_notsedated_sum
```

##### Sensitivity Analysis

```{r}
sens.out_notsedated6 <- medsens(mediation_results_cat6_notsedated, rho.by = 0.1, effect.type = "indirect", sims = 200)
```

#### Baseline sodium_cat_7

##### Mediation Analysis

```{r}
sodium_cat_7_notsedated_df$sodium_cat<-as.factor(sodium_cat_7_notsedated_df$sodium_cat)
sodium_cat_7_notsedated_df$sodium_cat<-relevel(sodium_cat_7_notsedated_df$sodium_cat,ref = "5")
fit_totaleffect<-glm(as.factor(hosp_mortality)~sodium_cat+AP_grp + ethnicity + gender + age + MAP + temperature + heartrate + respiratoryrate + mechvent_day01 + day1fio2 + day1pao2 + pco2 + ph + urine + creatinine + albumin + bilirubin + wbc + cirrhosis + aids + leukemia + immunosuppression + charlson_chf1 + charlson_liver1

                   , sodium_cat_7_notsedated_df
,family = binomial)
summ<-summary(fit_totaleffect)
total_effect_notsedated<-rbind(total_effect_notsedated,data.frame(cat = 7,estimated = summ$coefficients[2],se = summ$coefficients[2+length(summ$coefficients)/4],p = summ$coefficients[2+3*length(summ$coefficients)/4]))


fit_mediator<-lm(gcs_baseline ~ sodium_cat + AP_grp + ethnicity + gender + age + MAP + temperature + heartrate + respiratoryrate + mechvent_day01 + day1fio2 + day1pao2 + pco2 + ph + urine + creatinine + albumin + bilirubin + wbc + cirrhosis + aids + leukemia + immunosuppression + charlson_chf1 + charlson_liver1,
data =  sodium_cat_7_notsedated_df)
fit_dv<-glm(as.factor(hosp_mortality) ~gcs_baseline+sodium_cat + AP_grp + ethnicity + gender + age + MAP + temperature + heartrate + respiratoryrate + mechvent_day01 + day1fio2 + day1pao2 + pco2 + ph + urine + creatinine + albumin + bilirubin + wbc + cirrhosis + aids + leukemia + immunosuppression + charlson_chf1 + charlson_liver1
                   , sodium_cat_7_notsedated_df
           ,family = binomial("probit"))
summary(fit_dv)
#treat is the independend variable
mediation_results_cat7_notsedated <- mediate(fit_mediator, fit_dv, 
                                  treat="sodium_cat", mediator="gcs_baseline",control.value = 5,treat.value = 7,
                                  sims = 1000)
mediation_results_cat7_notsedated_sum<-as.data.frame(extract_mediation_summary(mediation_results_cat7_notsedated))
mediation_results_cat7_notsedated_sum<-mediation_results_cat7_notsedated_sum[c(5,8,9,10),]
mediation_results_cat7_notsedated_sum
```

##### Sensitivity Analysis

```{r}
sens.out_notsedated7 <- medsens(mediation_results_cat7_notsedated, rho.by = 0.1, effect.type = "indirect", sims = 200)
```

#### Baseline sodium_cat_8

##### Mediation Analysis

```{r}
sodium_cat_8_notsedated_df$sodium_cat<-as.factor(sodium_cat_8_notsedated_df$sodium_cat)
sodium_cat_8_notsedated_df$sodium_cat<-relevel(sodium_cat_8_notsedated_df$sodium_cat,ref = "5")
fit_totaleffect<-glm(as.factor(hosp_mortality)~sodium_cat + AP_grp + ethnicity + gender + age + MAP + temperature + heartrate + respiratoryrate + mechvent_day01 + day1fio2 + day1pao2 + pco2 + ph + urine + creatinine + albumin + bilirubin + wbc + cirrhosis + aids + leukemia + immunosuppression + charlson_chf1 + charlson_liver1
                   , sodium_cat_8_notsedated_df
,family = binomial)
summ<-summary(fit_totaleffect)
total_effect_notsedated<-rbind(total_effect_notsedated,data.frame(cat = 8,estimated = summ$coefficients[2],se = summ$coefficients[2+length(summ$coefficients)/4],p = summ$coefficients[2+3*length(summ$coefficients)/4]))


fit_mediator<-lm(gcs_baseline ~ sodium_cat + AP_grp + ethnicity + gender + age + MAP + temperature + heartrate + respiratoryrate + mechvent_day01 + day1fio2 + day1pao2 + pco2 + ph + urine + creatinine + albumin + bilirubin + wbc + cirrhosis + aids + leukemia + immunosuppression + charlson_chf1 + charlson_liver1,
data =  sodium_cat_8_notsedated_df)
fit_dv<-glm(as.factor(hosp_mortality) ~gcs_baseline+sodium_cat + AP_grp + ethnicity + gender + age + MAP + temperature + heartrate + respiratoryrate + mechvent_day01 + day1fio2 + day1pao2 + pco2 + ph + urine + creatinine + albumin + bilirubin + wbc + cirrhosis + aids + leukemia + immunosuppression + charlson_chf1 + charlson_liver1
                   , sodium_cat_8_notsedated_df
           ,family = binomial("probit"))
summary(fit_dv)
#treat is the independend variable
mediation_results_cat8_notsedated <- mediate(fit_mediator, fit_dv, 
                                  treat="sodium_cat", mediator="gcs_baseline",control.value = 5,treat.value = 8,
                                  sims = 1000)
mediation_results_cat8_notsedated_sum<-as.data.frame(extract_mediation_summary(mediation_results_cat8_notsedated))
mediation_results_cat8_notsedated_sum<-mediation_results_cat8_notsedated_sum[c(5,8,9,10),]
mediation_results_cat8_notsedated_sum
```

##### Sensitivity Analysis

```{r}
sens.out_notsedated8 <- medsens(mediation_results_cat8_notsedated, rho.by = 0.1, effect.type = "indirect", sims = 200)
```

# SubAnalysis 2, Surgery Type

## SubAnalysis 2A: Elective Surgery SubAnalysis

### Mediation package

#### Baseline sodium_cat_1

##### Mediation Analysis

```{r}
total_effect_elective<-NULL
sodium_cat_1_electivesurgery_df$sodium_cat<-as.factor(sodium_cat_1_electivesurgery_df$sodium_cat)
sodium_cat_1_electivesurgery_df$sodium_cat<-relevel(sodium_cat_1_electivesurgery_df$sodium_cat,ref = "5")
fit_totaleffect<-glm(as.factor(hosp_mortality)~sodium_cat+ AP_grp + ethnicity + gender + age + MAP + temperature + heartrate + respiratoryrate + mechvent_day01 + day1fio2 + day1pao2 + pco2 + ph + urine + creatinine + albumin + bilirubin + wbc + cirrhosis + aids + leukemia + immunosuppression + charlson_chf1 + charlson_liver1

                   , sodium_cat_1_electivesurgery_df
,family = binomial)
summ<-summary(fit_totaleffect)
total_effect_elective<-rbind(total_effect_elective,data.frame(cat = 1,estimated = summ$coefficients[2],se = summ$coefficients[2+length(summ$coefficients)/4],p = summ$coefficients[2+3*length(summ$coefficients)/4]))



fit_mediator<-lm(gcs_baseline ~ sodium_cat+AP_grp + ethnicity + gender + age + MAP + temperature + heartrate + respiratoryrate + mechvent_day01 + day1fio2 + day1pao2 + pco2 + ph + urine + creatinine + albumin + bilirubin + wbc + cirrhosis + aids + leukemia + immunosuppression + charlson_chf1 + charlson_liver1,
data =  sodium_cat_1_electivesurgery_df)

fit_dv<-glm(as.factor(hosp_mortality)~gcs_baseline+sodium_cat+ AP_grp + ethnicity + gender + age + MAP + temperature + heartrate + respiratoryrate + mechvent_day01 + day1fio2 + day1pao2 + pco2 + ph + urine + creatinine + albumin + bilirubin + wbc + cirrhosis + aids + leukemia + immunosuppression + charlson_chf1 + charlson_liver1
                   , sodium_cat_1_electivesurgery_df
           ,family = binomial("probit"))

summary(fit_dv)

#treat is the independend variable

mediation_results_cat1_elective <- mediate(fit_mediator, fit_dv, 
                                  treat="sodium_cat", mediator="gcs_baseline",control.value = 5,treat.value = 1,
                                  sims = 1000)
mediation_results_cat1_elective_sum<-as.data.frame(extract_mediation_summary(mediation_results_cat1_elective))
mediation_results_cat1_elective_sum<-mediation_results_cat1_elective_sum[c(5,8,9,10),]

#replace effect of treated group with average effect
#mediation_results_cat1_elective_sum<-mediation_results_cat1_elective_sum[grepl("treated|Total Effect", rownames(mediation_results_cat1_elective_sum)), ]#

mediation_results_cat1_elective_sum
```

##### Sensitivity Analysis

```{r}
sens.out_elective1 <- medsens(mediation_results_cat1_elective, rho.by = 0.1, effect.type = "indirect", sims = 200)
log_info()
```

#### Baseline sodium_cat_2

##### Mediation Analysis

```{r}
sodium_cat_2_electivesurgery_df$sodium_cat<-as.factor(sodium_cat_2_electivesurgery_df$sodium_cat)
sodium_cat_2_electivesurgery_df$sodium_cat<-relevel(sodium_cat_2_electivesurgery_df$sodium_cat,ref = "5")
fit_totaleffect<-glm(as.factor(hosp_mortality)~sodium_cat + AP_grp + ethnicity + gender + age + MAP + temperature + heartrate + respiratoryrate + mechvent_day01 + day1fio2 + day1pao2 + pco2 + ph + urine + creatinine + albumin + bilirubin + wbc + cirrhosis + aids + leukemia + immunosuppression + charlson_chf1 + charlson_liver1

                   , sodium_cat_2_electivesurgery_df
,family = binomial)
summ<-summary(fit_totaleffect)
total_effect_elective<-rbind(total_effect_elective,data.frame(cat = 2,estimated = summ$coefficients[2],se = summ$coefficients[2+length(summ$coefficients)/4],p = summ$coefficients[2+3*length(summ$coefficients)/4]))



fit_mediator<-lm(gcs_baseline ~ sodium_cat + AP_grp + ethnicity + gender + age + MAP + temperature + heartrate + respiratoryrate + mechvent_day01 + day1fio2 + day1pao2 + pco2 + ph + urine + creatinine + albumin + bilirubin + wbc + cirrhosis + aids + leukemia + immunosuppression + charlson_chf1 + charlson_liver1,
data =  sodium_cat_2_electivesurgery_df)
fit_dv<-glm(as.factor(hosp_mortality) ~gcs_baseline+sodium_cat+AP_grp + ethnicity + gender + age + MAP + temperature + heartrate + respiratoryrate + mechvent_day01 + day1fio2 + day1pao2 + pco2 + ph + urine + creatinine + albumin + bilirubin + wbc + cirrhosis + aids + leukemia + immunosuppression + charlson_chf1 + charlson_liver1
                   , sodium_cat_2_electivesurgery_df
           ,family = binomial("probit"))
summary(fit_dv)
#treat is the independend variable
mediation_results_cat2_elective <- mediate(fit_mediator, fit_dv, 
                                  treat="sodium_cat", mediator="gcs_baseline",control.value = 5,treat.value = 2,
                                  sims = 1000)
mediation_results_cat2_elective_sum<-as.data.frame(extract_mediation_summary(mediation_results_cat2_elective))
mediation_results_cat2_elective_sum<-mediation_results_cat2_elective_sum[c(5,8,9,10),]

mediation_results_cat2_elective_sum
```

##### Sensitivity Analysis

```{r}
sens.out_elective2 <- medsens(mediation_results_cat2_elective, rho.by = 0.1, effect.type = "indirect", sims = 200)
log_info()
```

#### Baseline sodium_cat_3

##### Mediation Analysis

```{r}
sodium_cat_3_electivesurgery_df$sodium_cat<-as.factor(sodium_cat_3_electivesurgery_df$sodium_cat)
sodium_cat_3_electivesurgery_df$sodium_cat<-relevel(sodium_cat_3_electivesurgery_df$sodium_cat,ref = "5")
fit_totaleffect<-glm(as.factor(hosp_mortality)~sodium_cat+age+AP_grp+gender+ethnicity+unittype+charlson_liver1 +charlson_chf1+mechvent_day01+respiratoryrate+MAP

                   , sodium_cat_3_electivesurgery_df
,family = binomial)
summ<-summary(fit_totaleffect)
total_effect_elective<-rbind(total_effect_elective,data.frame(cat = 3,estimated = summ$coefficients[2],se = summ$coefficients[2+length(summ$coefficients)/4],p = summ$coefficients[2+3*length(summ$coefficients)/4]))

fit_mediator<-lm(gcs_baseline ~ sodium_cat+AP_grp + ethnicity + gender + age + MAP + temperature + heartrate + respiratoryrate + mechvent_day01 + day1fio2 + day1pao2 + pco2 + ph + urine + creatinine + albumin + bilirubin + wbc + cirrhosis + aids + leukemia + immunosuppression + charlson_chf1 + charlson_liver1,
data =  sodium_cat_3_electivesurgery_df)
fit_dv<-glm(as.factor(hosp_mortality) ~gcs_baseline+sodium_cat+AP_grp + ethnicity + gender + age + MAP + temperature + heartrate + respiratoryrate + mechvent_day01 + day1fio2 + day1pao2 + pco2 + ph + urine + creatinine + albumin + bilirubin + wbc + cirrhosis + aids + leukemia + immunosuppression + charlson_chf1 + charlson_liver1
                   , sodium_cat_3_electivesurgery_df
           ,family = binomial("probit"))
summary(fit_dv)
#treat is the independend variable
mediation_results_cat3_elective <- mediate(fit_mediator, fit_dv, 
                                  treat="sodium_cat", mediator="gcs_baseline",control.value = 5,treat.value = 3,
                                  sims = 1000)
mediation_results_cat3_elective_sum<-as.data.frame(extract_mediation_summary(mediation_results_cat3_elective))
mediation_results_cat3_elective_sum<-mediation_results_cat3_elective_sum[c(5,8,9,10),]

mediation_results_cat3_elective_sum
```

##### Sensitivity Analysis

```{r}
sens.out_elective3 <- medsens(mediation_results_cat3_elective, rho.by = 0.1, effect.type = "indirect", sims = 200)
```

#### Baseline sodium_cat_4

##### Mediation Analysis

```{r}
sodium_cat_4_electivesurgery_df$sodium_cat<-as.factor(sodium_cat_4_electivesurgery_df$sodium_cat)
sodium_cat_4_electivesurgery_df$sodium_cat<-relevel(sodium_cat_4_electivesurgery_df$sodium_cat,ref = "5")
fit_totaleffect<-glm(as.factor(hosp_mortality)~sodium_cat+AP_grp + ethnicity + gender + age + MAP + temperature + heartrate + respiratoryrate + mechvent_day01 + day1fio2 + day1pao2 + pco2 + ph + urine + creatinine + albumin + bilirubin + wbc + cirrhosis + aids + leukemia + immunosuppression + charlson_chf1 + charlson_liver1
                   , sodium_cat_4_electivesurgery_df
,family = binomial)
summ<-summary(fit_totaleffect)
total_effect_elective<-rbind(total_effect_elective,data.frame(cat = 4,estimated = summ$coefficients[2],se = summ$coefficients[2+length(summ$coefficients)/4],p = summ$coefficients[2+3*length(summ$coefficients)/4]))


fit_mediator<-lm(gcs_baseline ~ sodium_cat+AP_grp + ethnicity + gender + age + MAP + temperature + heartrate + respiratoryrate + mechvent_day01 + day1fio2 + day1pao2 + pco2 + ph + urine + creatinine + albumin + bilirubin + wbc + cirrhosis + aids + leukemia + immunosuppression + charlson_chf1 + charlson_liver1,
data =  sodium_cat_4_electivesurgery_df)
fit_dv<-glm(as.factor(hosp_mortality) ~gcs_baseline+sodium_cat+AP_grp + ethnicity + gender + age + MAP + temperature + heartrate + respiratoryrate + mechvent_day01 + day1fio2 + day1pao2 + pco2 + ph + urine + creatinine + albumin + bilirubin + wbc + cirrhosis + aids + leukemia + immunosuppression + charlson_chf1 + charlson_liver1
                   , sodium_cat_4_electivesurgery_df
           ,family = binomial("probit"))
summary(fit_dv)
#treat is the independend variable
mediation_results_cat4_elective <- mediate(fit_mediator, fit_dv, 
                                  treat="sodium_cat", mediator="gcs_baseline",control.value = 5,treat.value = 4,
                                  sims = 1000)
mediation_results_cat4_elective_sum<-as.data.frame(extract_mediation_summary(mediation_results_cat4_elective))
mediation_results_cat4_elective_sum<-mediation_results_cat4_elective_sum[c(5,8,9,10),]
mediation_results_cat4_elective_sum

```

##### Sensitivity Analysis

```{r}
sens.out_elective4 <- medsens(mediation_results_cat4_elective, rho.by = 0.1, effect.type = "indirect", sims = 200)
```

#### Baseline sodium_cat_6

##### Mediation Analysis

```{r}
sodium_cat_6_electivesurgery_df$sodium_cat<-as.factor(sodium_cat_6_electivesurgery_df$sodium_cat)
sodium_cat_6_electivesurgery_df$sodium_cat<-relevel(sodium_cat_6_electivesurgery_df$sodium_cat,ref = "5")
fit_totaleffect<-glm(as.factor(hosp_mortality)~sodium_cat+AP_grp + ethnicity + gender + age + MAP + temperature + heartrate + respiratoryrate + mechvent_day01 + day1fio2 + day1pao2 + pco2 + ph + urine + creatinine + albumin + bilirubin + wbc + cirrhosis + aids + leukemia + immunosuppression + charlson_chf1 + charlson_liver1

                   , sodium_cat_6_electivesurgery_df
,family = binomial)
summ<-summary(fit_totaleffect)
total_effect_elective<-rbind(total_effect_elective,data.frame(cat = 6,estimated = summ$coefficients[2],se = summ$coefficients[2+length(summ$coefficients)/4],p = summ$coefficients[2+3*length(summ$coefficients)/4]))


fit_mediator<-lm(gcs_baseline ~ sodium_cat+AP_grp + ethnicity + gender + age + MAP + temperature + heartrate + respiratoryrate + mechvent_day01 + day1fio2 + day1pao2 + pco2 + ph + urine + creatinine + albumin + bilirubin + wbc + cirrhosis + aids + leukemia + immunosuppression + charlson_chf1 + charlson_liver1,
data =  sodium_cat_6_electivesurgery_df)
fit_dv<-glm(as.factor(hosp_mortality) ~gcs_baseline+sodium_cat+AP_grp + ethnicity + gender + age + MAP + temperature + heartrate + respiratoryrate + mechvent_day01 + day1fio2 + day1pao2 + pco2 + ph + urine + creatinine + albumin + bilirubin + wbc + cirrhosis + aids + leukemia + immunosuppression + charlson_chf1 + charlson_liver1
                   , sodium_cat_6_electivesurgery_df
           ,family = binomial("probit"))
summary(fit_dv)
#treat is the independend variable
mediation_results_cat6_elective <- mediate(fit_mediator, fit_dv, 
                                  treat="sodium_cat", mediator="gcs_baseline",control.value = 5,treat.value = 6,
                                  sims = 1000)
mediation_results_cat6_elective_sum<-as.data.frame(extract_mediation_summary(mediation_results_cat6_elective))
mediation_results_cat6_elective_sum<-mediation_results_cat6_elective_sum[c(5,8,9,10),]
mediation_results_cat6_elective_sum
```

##### Sensitivity Analysis

```{r}
sens.out_elective6 <- medsens(mediation_results_cat6_elective, rho.by = 0.1, effect.type = "indirect", sims = 200)
```

#### Baseline sodium_cat_7

##### Mediation Analysis

```{r}
sodium_cat_7_electivesurgery_df$sodium_cat<-as.factor(sodium_cat_7_electivesurgery_df$sodium_cat)
sodium_cat_7_electivesurgery_df$sodium_cat<-relevel(sodium_cat_7_electivesurgery_df$sodium_cat,ref = "5")
fit_totaleffect<-glm(as.factor(hosp_mortality)~sodium_cat+AP_grp + ethnicity + gender + age + MAP + temperature + heartrate + respiratoryrate + mechvent_day01 + day1fio2 + day1pao2 + pco2 + ph + urine + creatinine + albumin + bilirubin + wbc + cirrhosis + aids + leukemia + immunosuppression + charlson_chf1 + charlson_liver1

                   , sodium_cat_7_electivesurgery_df
,family = binomial)
summ<-summary(fit_totaleffect)
total_effect_elective<-rbind(total_effect_elective,data.frame(cat = 7,estimated = summ$coefficients[2],se = summ$coefficients[2+length(summ$coefficients)/4],p = summ$coefficients[2+3*length(summ$coefficients)/4]))


fit_mediator<-lm(gcs_baseline ~ sodium_cat + AP_grp + ethnicity + gender + age + MAP + temperature + heartrate + respiratoryrate + mechvent_day01 + day1fio2 + day1pao2 + pco2 + ph + urine + creatinine + albumin + bilirubin + wbc + cirrhosis + aids + leukemia + immunosuppression + charlson_chf1 + charlson_liver1,
data =  sodium_cat_7_electivesurgery_df)
fit_dv<-glm(as.factor(hosp_mortality) ~gcs_baseline+sodium_cat + AP_grp + ethnicity + gender + age + MAP + temperature + heartrate + respiratoryrate + mechvent_day01 + day1fio2 + day1pao2 + pco2 + ph + urine + creatinine + albumin + bilirubin + wbc + cirrhosis + aids + leukemia + immunosuppression + charlson_chf1 + charlson_liver1
                   , sodium_cat_7_electivesurgery_df
           ,family = binomial("probit"))
summary(fit_dv)
#treat is the independend variable
mediation_results_cat7_elective <- mediate(fit_mediator, fit_dv, 
                                  treat="sodium_cat", mediator="gcs_baseline",control.value = 5,treat.value = 7,
                                  sims = 1000)
mediation_results_cat7_elective_sum<-as.data.frame(extract_mediation_summary(mediation_results_cat7_elective))
mediation_results_cat7_elective_sum<-mediation_results_cat7_elective_sum[c(5,8,9,10),]
mediation_results_cat7_elective_sum
```

##### Sensitivity Analysis

```{r}
sens.out_elective7 <- medsens(mediation_results_cat7_elective, rho.by = 0.1, effect.type = "indirect", sims = 200)
```

#### Baseline sodium_cat_8

##### Mediation Analysis

```{r}
sodium_cat_8_electivesurgery_df$sodium_cat<-as.factor(sodium_cat_8_electivesurgery_df$sodium_cat)
sodium_cat_8_electivesurgery_df$sodium_cat<-relevel(sodium_cat_8_electivesurgery_df$sodium_cat,ref = "5")
fit_totaleffect<-glm(as.factor(hosp_mortality)~sodium_cat + AP_grp + ethnicity + gender + age + MAP + temperature + heartrate + respiratoryrate + mechvent_day01 + day1fio2 + day1pao2 + pco2 + ph + urine + creatinine + albumin + bilirubin + wbc + cirrhosis + aids + leukemia + immunosuppression + charlson_chf1 + charlson_liver1
                   , sodium_cat_8_electivesurgery_df
,family = binomial)
summ<-summary(fit_totaleffect)
total_effect_elective<-rbind(total_effect_elective,data.frame(cat = 8,estimated = summ$coefficients[2],se = summ$coefficients[2+length(summ$coefficients)/4],p = summ$coefficients[2+3*length(summ$coefficients)/4]))


fit_mediator<-lm(gcs_baseline ~ sodium_cat + AP_grp + ethnicity + gender + age + MAP + temperature + heartrate + respiratoryrate + mechvent_day01 + day1fio2 + day1pao2 + pco2 + ph + urine + creatinine + albumin + bilirubin + wbc + cirrhosis + aids + leukemia + immunosuppression + charlson_chf1 + charlson_liver1,
data =  sodium_cat_8_electivesurgery_df)
fit_dv<-glm(as.factor(hosp_mortality) ~gcs_baseline+sodium_cat + AP_grp + ethnicity + gender + age + MAP + temperature + heartrate + respiratoryrate + mechvent_day01 + day1fio2 + day1pao2 + pco2 + ph + urine + creatinine + albumin + bilirubin + wbc + cirrhosis + aids + leukemia + immunosuppression + charlson_chf1 + charlson_liver1
                   , sodium_cat_8_electivesurgery_df
           ,family = binomial("probit"))
summary(fit_dv)
#treat is the independend variable
mediation_results_cat8_elective <- mediate(fit_mediator, fit_dv, 
                                  treat="sodium_cat", mediator="gcs_baseline",control.value = 5,treat.value = 8,
                                  sims = 1000)
mediation_results_cat8_elective_sum<-as.data.frame(extract_mediation_summary(mediation_results_cat8_elective))
mediation_results_cat8_elective_sum<-mediation_results_cat8_elective_sum[c(5,8,9,10),]
mediation_results_cat8_elective_sum
```

##### Sensitivity Analysis

```{r}
sens.out_elective8 <- medsens(mediation_results_cat8_elective, rho.by = 0.1, effect.type = "indirect", sims = 200)
```

## SubAnalysis 2B: Not Elective Surgery SubAnalysis

### Mediation package

#### Baseline sodium_cat_1

##### Mediation Analysis

```{r}
total_effect_not_elective<-NULL
sodium_cat_1_not_electivesurgery_df$sodium_cat<-as.factor(sodium_cat_1_not_electivesurgery_df$sodium_cat)
sodium_cat_1_not_electivesurgery_df$sodium_cat<-relevel(sodium_cat_1_not_electivesurgery_df$sodium_cat,ref = "5")
fit_totaleffect<-glm(as.factor(hosp_mortality)~sodium_cat+ AP_grp + ethnicity + gender + age + MAP + temperature + heartrate + respiratoryrate + mechvent_day01 + day1fio2 + day1pao2 + pco2 + ph + urine + creatinine + albumin + bilirubin + wbc + cirrhosis + aids + leukemia + immunosuppression + charlson_chf1 + charlson_liver1

                   , sodium_cat_1_not_electivesurgery_df
,family = binomial)
summ<-summary(fit_totaleffect)
total_effect_not_elective<-rbind(total_effect_not_elective,data.frame(cat = 1,estimated = summ$coefficients[2],se = summ$coefficients[2+length(summ$coefficients)/4],p = summ$coefficients[2+3*length(summ$coefficients)/4]))



fit_mediator<-lm(gcs_baseline ~ sodium_cat+AP_grp + ethnicity + gender + age + MAP + temperature + heartrate + respiratoryrate + mechvent_day01 + day1fio2 + day1pao2 + pco2 + ph + urine + creatinine + albumin + bilirubin + wbc + cirrhosis + aids + leukemia + immunosuppression + charlson_chf1 + charlson_liver1,
data =  sodium_cat_1_not_electivesurgery_df)

fit_dv<-glm(as.factor(hosp_mortality)~gcs_baseline+sodium_cat+ AP_grp + ethnicity + gender + age + MAP + temperature + heartrate + respiratoryrate + mechvent_day01 + day1fio2 + day1pao2 + pco2 + ph + urine + creatinine + albumin + bilirubin + wbc + cirrhosis + aids + leukemia + immunosuppression + charlson_chf1 + charlson_liver1
                   , sodium_cat_1_not_electivesurgery_df
           ,family = binomial("probit"))

summary(fit_dv)

#treat is the independend variable

mediation_results_cat1_not_elective <- mediate(fit_mediator, fit_dv, 
                                  treat="sodium_cat", mediator="gcs_baseline",control.value = 5,treat.value = 1,
                                  sims = 1000)
mediation_results_cat1_not_elective_sum<-as.data.frame(extract_mediation_summary(mediation_results_cat1_not_elective))
mediation_results_cat1_not_elective_sum<-mediation_results_cat1_not_elective_sum[c(5,8,9,10),]

#replace effect of treated group with average effect
#mediation_results_cat1_not_elective_sum<-mediation_results_cat1_not_elective_sum[grepl("treated|Total Effect", rownames(mediation_results_cat1_not_elective_sum)), ]#

mediation_results_cat1_not_elective_sum
```

##### Sensitivity Analysis

```{r}
sens.out_not_elective1 <- medsens(mediation_results_cat1_not_elective, rho.by = 0.1, effect.type = "indirect", sims = 200)
log_info()
```

#### Baseline sodium_cat_2

##### Mediation Analysis

```{r}
sodium_cat_2_not_electivesurgery_df$sodium_cat<-as.factor(sodium_cat_2_not_electivesurgery_df$sodium_cat)
sodium_cat_2_not_electivesurgery_df$sodium_cat<-relevel(sodium_cat_2_not_electivesurgery_df$sodium_cat,ref = "5")
fit_totaleffect<-glm(as.factor(hosp_mortality)~sodium_cat + AP_grp + ethnicity + gender + age + MAP + temperature + heartrate + respiratoryrate + mechvent_day01 + day1fio2 + day1pao2 + pco2 + ph + urine + creatinine + albumin + bilirubin + wbc + cirrhosis + aids + leukemia + immunosuppression + charlson_chf1 + charlson_liver1

                   , sodium_cat_2_not_electivesurgery_df
,family = binomial)
summ<-summary(fit_totaleffect)
total_effect_not_elective<-rbind(total_effect_not_elective,data.frame(cat = 2,estimated = summ$coefficients[2],se = summ$coefficients[2+length(summ$coefficients)/4],p = summ$coefficients[2+3*length(summ$coefficients)/4]))



fit_mediator<-lm(gcs_baseline ~ sodium_cat + AP_grp + ethnicity + gender + age + MAP + temperature + heartrate + respiratoryrate + mechvent_day01 + day1fio2 + day1pao2 + pco2 + ph + urine + creatinine + albumin + bilirubin + wbc + cirrhosis + aids + leukemia + immunosuppression + charlson_chf1 + charlson_liver1,
data =  sodium_cat_2_not_electivesurgery_df)
fit_dv<-glm(as.factor(hosp_mortality) ~gcs_baseline+sodium_cat+AP_grp + ethnicity + gender + age + MAP + temperature + heartrate + respiratoryrate + mechvent_day01 + day1fio2 + day1pao2 + pco2 + ph + urine + creatinine + albumin + bilirubin + wbc + cirrhosis + aids + leukemia + immunosuppression + charlson_chf1 + charlson_liver1
                   , sodium_cat_2_not_electivesurgery_df
           ,family = binomial("probit"))
summary(fit_dv)
#treat is the independend variable
mediation_results_cat2_not_elective <- mediate(fit_mediator, fit_dv, 
                                  treat="sodium_cat", mediator="gcs_baseline",control.value = 5,treat.value = 2,
                                  sims = 1000)
mediation_results_cat2_not_elective_sum<-as.data.frame(extract_mediation_summary(mediation_results_cat2_not_elective))
mediation_results_cat2_not_elective_sum<-mediation_results_cat2_not_elective_sum[c(5,8,9,10),]

mediation_results_cat2_not_elective_sum
```

##### Sensitivity Analysis

```{r}
sens.out_not_elective2 <- medsens(mediation_results_cat2_not_elective, rho.by = 0.1, effect.type = "indirect", sims = 200)
log_info()
```

#### Baseline sodium_cat_3

##### Mediation Analysis

```{r}
sodium_cat_3_not_electivesurgery_df$sodium_cat<-as.factor(sodium_cat_3_not_electivesurgery_df$sodium_cat)
sodium_cat_3_not_electivesurgery_df$sodium_cat<-relevel(sodium_cat_3_not_electivesurgery_df$sodium_cat,ref = "5")
fit_totaleffect<-glm(as.factor(hosp_mortality)~sodium_cat+age+AP_grp+gender+ethnicity+unittype+charlson_liver1 +charlson_chf1+mechvent_day01+respiratoryrate+MAP

                   , sodium_cat_3_not_electivesurgery_df
,family = binomial)
summ<-summary(fit_totaleffect)
total_effect_not_elective<-rbind(total_effect_not_elective,data.frame(cat = 3,estimated = summ$coefficients[2],se = summ$coefficients[2+length(summ$coefficients)/4],p = summ$coefficients[2+3*length(summ$coefficients)/4]))

fit_mediator<-lm(gcs_baseline ~ sodium_cat+AP_grp + ethnicity + gender + age + MAP + temperature + heartrate + respiratoryrate + mechvent_day01 + day1fio2 + day1pao2 + pco2 + ph + urine + creatinine + albumin + bilirubin + wbc + cirrhosis + aids + leukemia + immunosuppression + charlson_chf1 + charlson_liver1,
data =  sodium_cat_3_not_electivesurgery_df)
fit_dv<-glm(as.factor(hosp_mortality) ~gcs_baseline+sodium_cat+AP_grp + ethnicity + gender + age + MAP + temperature + heartrate + respiratoryrate + mechvent_day01 + day1fio2 + day1pao2 + pco2 + ph + urine + creatinine + albumin + bilirubin + wbc + cirrhosis + aids + leukemia + immunosuppression + charlson_chf1 + charlson_liver1
                   , sodium_cat_3_not_electivesurgery_df
           ,family = binomial("probit"))
summary(fit_dv)
#treat is the independend variable
mediation_results_cat3_not_elective <- mediate(fit_mediator, fit_dv, 
                                  treat="sodium_cat", mediator="gcs_baseline",control.value = 5,treat.value = 3,
                                  sims = 1000)
mediation_results_cat3_not_elective_sum<-as.data.frame(extract_mediation_summary(mediation_results_cat3_not_elective))
mediation_results_cat3_not_elective_sum<-mediation_results_cat3_not_elective_sum[c(5,8,9,10),]

mediation_results_cat3_not_elective_sum
```

##### Sensitivity Analysis

```{r}
sens.out_not_elective3 <- medsens(mediation_results_cat3_not_elective, rho.by = 0.1, effect.type = "indirect", sims = 200)
```

#### Baseline sodium_cat_4

##### Mediation Analysis

```{r}
sodium_cat_4_not_electivesurgery_df$sodium_cat<-as.factor(sodium_cat_4_not_electivesurgery_df$sodium_cat)
sodium_cat_4_not_electivesurgery_df$sodium_cat<-relevel(sodium_cat_4_not_electivesurgery_df$sodium_cat,ref = "5")
fit_totaleffect<-glm(as.factor(hosp_mortality)~sodium_cat+AP_grp + ethnicity + gender + age + MAP + temperature + heartrate + respiratoryrate + mechvent_day01 + day1fio2 + day1pao2 + pco2 + ph + urine + creatinine + albumin + bilirubin + wbc + cirrhosis + aids + leukemia + immunosuppression + charlson_chf1 + charlson_liver1
                   , sodium_cat_4_not_electivesurgery_df
,family = binomial)
summ<-summary(fit_totaleffect)
total_effect_not_elective<-rbind(total_effect_not_elective,data.frame(cat = 4,estimated = summ$coefficients[2],se = summ$coefficients[2+length(summ$coefficients)/4],p = summ$coefficients[2+3*length(summ$coefficients)/4]))


fit_mediator<-lm(gcs_baseline ~ sodium_cat+AP_grp + ethnicity + gender + age + MAP + temperature + heartrate + respiratoryrate + mechvent_day01 + day1fio2 + day1pao2 + pco2 + ph + urine + creatinine + albumin + bilirubin + wbc + cirrhosis + aids + leukemia + immunosuppression + charlson_chf1 + charlson_liver1,
data =  sodium_cat_4_not_electivesurgery_df)
fit_dv<-glm(as.factor(hosp_mortality) ~gcs_baseline+sodium_cat+AP_grp + ethnicity + gender + age + MAP + temperature + heartrate + respiratoryrate + mechvent_day01 + day1fio2 + day1pao2 + pco2 + ph + urine + creatinine + albumin + bilirubin + wbc + cirrhosis + aids + leukemia + immunosuppression + charlson_chf1 + charlson_liver1
                   , sodium_cat_4_not_electivesurgery_df
           ,family = binomial("probit"))
summary(fit_dv)
#treat is the independend variable
mediation_results_cat4_not_elective <- mediate(fit_mediator, fit_dv, 
                                  treat="sodium_cat", mediator="gcs_baseline",control.value = 5,treat.value = 4,
                                  sims = 1000)
mediation_results_cat4_not_elective_sum<-as.data.frame(extract_mediation_summary(mediation_results_cat4_not_elective))
mediation_results_cat4_not_elective_sum<-mediation_results_cat4_not_elective_sum[c(5,8,9,10),]
mediation_results_cat4_not_elective_sum

```

##### Sensitivity Analysis

```{r}
sens.out_not_elective4 <- medsens(mediation_results_cat4_not_elective, rho.by = 0.1, effect.type = "indirect", sims = 200)
```

#### Baseline sodium_cat_6

##### Mediation Analysis

```{r}
sodium_cat_6_not_electivesurgery_df$sodium_cat<-as.factor(sodium_cat_6_not_electivesurgery_df$sodium_cat)
sodium_cat_6_not_electivesurgery_df$sodium_cat<-relevel(sodium_cat_6_not_electivesurgery_df$sodium_cat,ref = "5")
fit_totaleffect<-glm(as.factor(hosp_mortality)~sodium_cat+AP_grp + ethnicity + gender + age + MAP + temperature + heartrate + respiratoryrate + mechvent_day01 + day1fio2 + day1pao2 + pco2 + ph + urine + creatinine + albumin + bilirubin + wbc + cirrhosis + aids + leukemia + immunosuppression + charlson_chf1 + charlson_liver1

                   , sodium_cat_6_not_electivesurgery_df
,family = binomial)
summ<-summary(fit_totaleffect)
total_effect_not_elective<-rbind(total_effect_not_elective,data.frame(cat = 6,estimated = summ$coefficients[2],se = summ$coefficients[2+length(summ$coefficients)/4],p = summ$coefficients[2+3*length(summ$coefficients)/4]))


fit_mediator<-lm(gcs_baseline ~ sodium_cat+AP_grp + ethnicity + gender + age + MAP + temperature + heartrate + respiratoryrate + mechvent_day01 + day1fio2 + day1pao2 + pco2 + ph + urine + creatinine + albumin + bilirubin + wbc + cirrhosis + aids + leukemia + immunosuppression + charlson_chf1 + charlson_liver1,
data =  sodium_cat_6_not_electivesurgery_df)
fit_dv<-glm(as.factor(hosp_mortality) ~gcs_baseline+sodium_cat+AP_grp + ethnicity + gender + age + MAP + temperature + heartrate + respiratoryrate + mechvent_day01 + day1fio2 + day1pao2 + pco2 + ph + urine + creatinine + albumin + bilirubin + wbc + cirrhosis + aids + leukemia + immunosuppression + charlson_chf1 + charlson_liver1
                   , sodium_cat_6_not_electivesurgery_df
           ,family = binomial("probit"))
summary(fit_dv)
#treat is the independend variable
mediation_results_cat6_not_elective <- mediate(fit_mediator, fit_dv, 
                                  treat="sodium_cat", mediator="gcs_baseline",control.value = 5,treat.value = 6,
                                  sims = 1000)
mediation_results_cat6_not_elective_sum<-as.data.frame(extract_mediation_summary(mediation_results_cat6_not_elective))
mediation_results_cat6_not_elective_sum<-mediation_results_cat6_not_elective_sum[c(5,8,9,10),]
mediation_results_cat6_not_elective_sum
```

##### Sensitivity Analysis

```{r}
sens.out_not_elective6 <- medsens(mediation_results_cat6_not_elective, rho.by = 0.1, effect.type = "indirect", sims = 200)
```

#### Baseline sodium_cat_7

##### Mediation Analysis

```{r}
sodium_cat_7_not_electivesurgery_df$sodium_cat<-as.factor(sodium_cat_7_not_electivesurgery_df$sodium_cat)
sodium_cat_7_not_electivesurgery_df$sodium_cat<-relevel(sodium_cat_7_not_electivesurgery_df$sodium_cat,ref = "5")
fit_totaleffect<-glm(as.factor(hosp_mortality)~sodium_cat+AP_grp + ethnicity + gender + age + MAP + temperature + heartrate + respiratoryrate + mechvent_day01 + day1fio2 + day1pao2 + pco2 + ph + urine + creatinine + albumin + bilirubin + wbc + cirrhosis + aids + leukemia + immunosuppression + charlson_chf1 + charlson_liver1

                   , sodium_cat_7_not_electivesurgery_df
,family = binomial)
summ<-summary(fit_totaleffect)
total_effect_not_elective<-rbind(total_effect_not_elective,data.frame(cat = 7,estimated = summ$coefficients[2],se = summ$coefficients[2+length(summ$coefficients)/4],p = summ$coefficients[2+3*length(summ$coefficients)/4]))


fit_mediator<-lm(gcs_baseline ~ sodium_cat + AP_grp + ethnicity + gender + age + MAP + temperature + heartrate + respiratoryrate + mechvent_day01 + day1fio2 + day1pao2 + pco2 + ph + urine + creatinine + albumin + bilirubin + wbc + cirrhosis + aids + leukemia + immunosuppression + charlson_chf1 + charlson_liver1,
data =  sodium_cat_7_not_electivesurgery_df)
fit_dv<-glm(as.factor(hosp_mortality) ~gcs_baseline+sodium_cat + AP_grp + ethnicity + gender + age + MAP + temperature + heartrate + respiratoryrate + mechvent_day01 + day1fio2 + day1pao2 + pco2 + ph + urine + creatinine + albumin + bilirubin + wbc + cirrhosis + aids + leukemia + immunosuppression + charlson_chf1 + charlson_liver1
                   , sodium_cat_7_not_electivesurgery_df
           ,family = binomial("probit"))
summary(fit_dv)
#treat is the independend variable
mediation_results_cat7_not_elective <- mediate(fit_mediator, fit_dv, 
                                  treat="sodium_cat", mediator="gcs_baseline",control.value = 5,treat.value = 7,
                                  sims = 1000)
mediation_results_cat7_not_elective_sum<-as.data.frame(extract_mediation_summary(mediation_results_cat7_not_elective))
mediation_results_cat7_not_elective_sum<-mediation_results_cat7_not_elective_sum[c(5,8,9,10),]
mediation_results_cat7_not_elective_sum
```

##### Sensitivity Analysis

```{r}
sens.out_not_elective7 <- medsens(mediation_results_cat7_not_elective, rho.by = 0.1, effect.type = "indirect", sims = 200)
```

#### Baseline sodium_cat_8

##### Mediation Analysis

```{r}
sodium_cat_8_not_electivesurgery_df$sodium_cat<-as.factor(sodium_cat_8_not_electivesurgery_df$sodium_cat)
sodium_cat_8_not_electivesurgery_df$sodium_cat<-relevel(sodium_cat_8_not_electivesurgery_df$sodium_cat,ref = "5")
fit_totaleffect<-glm(as.factor(hosp_mortality)~sodium_cat + AP_grp + ethnicity + gender + age + MAP + temperature + heartrate + respiratoryrate + mechvent_day01 + day1fio2 + day1pao2 + pco2 + ph + urine + creatinine + albumin + bilirubin + wbc + cirrhosis + aids + leukemia + immunosuppression + charlson_chf1 + charlson_liver1
                   , sodium_cat_8_not_electivesurgery_df
,family = binomial)
summ<-summary(fit_totaleffect)
total_effect_not_elective<-rbind(total_effect_not_elective,data.frame(cat = 8,estimated = summ$coefficients[2],se = summ$coefficients[2+length(summ$coefficients)/4],p = summ$coefficients[2+3*length(summ$coefficients)/4]))


fit_mediator<-lm(gcs_baseline ~ sodium_cat + AP_grp + ethnicity + gender + age + MAP + temperature + heartrate + respiratoryrate + mechvent_day01 + day1fio2 + day1pao2 + pco2 + ph + urine + creatinine + albumin + bilirubin + wbc + cirrhosis + aids + leukemia + immunosuppression + charlson_chf1 + charlson_liver1,
data =  sodium_cat_8_not_electivesurgery_df)
fit_dv<-glm(as.factor(hosp_mortality) ~gcs_baseline+sodium_cat + AP_grp + ethnicity + gender + age + MAP + temperature + heartrate + respiratoryrate + mechvent_day01 + day1fio2 + day1pao2 + pco2 + ph + urine + creatinine + albumin + bilirubin + wbc + cirrhosis + aids + leukemia + immunosuppression + charlson_chf1 + charlson_liver1
                   , sodium_cat_8_not_electivesurgery_df
           ,family = binomial("probit"))
summary(fit_dv)
#treat is the independend variable
mediation_results_cat8_not_elective <- mediate(fit_mediator, fit_dv, 
                                  treat="sodium_cat", mediator="gcs_baseline",control.value = 5,treat.value = 8,
                                  sims = 1000)
mediation_results_cat8_not_elective_sum<-as.data.frame(extract_mediation_summary(mediation_results_cat8_not_elective))
mediation_results_cat8_not_elective_sum<-mediation_results_cat8_not_elective_sum[c(5,8,9,10),]
mediation_results_cat8_not_elective_sum
```

##### Sensitivity Analysis

```{r}
sens.out_not_elective8 <- medsens(mediation_results_cat8_not_elective, rho.by = 0.1, effect.type = "indirect", sims = 200)
```

## Subanalysis 3, ICU Type

```{r}
# Function to perform mediation analysis
perform_mediation_analysis <- function(df, cat_value, icu_name) {
  # Convert sodium_cat to factor and set the reference
  df$sodium_cat <- as.factor(df$sodium_cat)
  df$sodium_cat <- relevel(df$sodium_cat, ref = "5")
  
  # Fit the total effect model
  fit_totaleffect <- glm(as.factor(hosp_mortality) ~ sodium_cat + AP_grp + ethnicity + gender + age + MAP + temperature + heartrate + respiratoryrate + mechvent_day01 + day1fio2 + day1pao2 + pco2 + ph + urine + creatinine + albumin + bilirubin + wbc + cirrhosis + aids + leukemia + immunosuppression + charlson_chf1 + charlson_liver1,
                         data = df, family = binomial)
  summ <- summary(fit_totaleffect)
  
  # Save the total effect results
  total_effect <- data.frame(cat = cat_value, estimated = summ$coefficients[2], se = summ$coefficients[2 + length(summ$coefficients) / 4], p = summ$coefficients[2 + 3 * length(summ$coefficients) / 4])
  
  # Fit the mediator model
  fit_mediator <- lm(gcs_baseline ~ sodium_cat + AP_grp + ethnicity + gender + age + MAP + temperature + heartrate + respiratoryrate + mechvent_day01 + day1fio2 + day1pao2 + pco2 + ph + urine + creatinine + albumin + bilirubin + wbc + cirrhosis + aids + leukemia + immunosuppression + charlson_chf1 + charlson_liver1, data = df)
  
  # Fit the dependent variable model
  fit_dv <- glm(as.factor(hosp_mortality) ~ gcs_baseline + sodium_cat + AP_grp + ethnicity + gender + age + MAP + temperature + heartrate + respiratoryrate + mechvent_day01 + day1fio2 + day1pao2 + pco2 + ph + urine + creatinine + albumin + bilirubin + wbc + cirrhosis + aids + leukemia + immunosuppression + charlson_chf1 + charlson_liver1,
                data = df, family = binomial("probit"))
  
  # Perform the mediation analysis
  mediation_results <- mediate(fit_mediator, fit_dv, treat = "sodium_cat", mediator = "gcs_baseline", control.value = 5, treat.value = cat_value, sims = 1000)
  mediation_results_sum <- as.data.frame(extract_mediation_summary(mediation_results))
  mediation_results_sum <- mediation_results_sum[c(5, 8, 9, 10),]
  
  # Perform the sensitivity analysis
  sens_out <- medsens(mediation_results, rho.by = 0.1, effect.type = "indirect", sims = 200)
  
  return(list(total_effect = total_effect, mediation_results_sum = mediation_results_sum, sens_out = sens_out))
}

# Values of sodium_cat to analyze
sodium_cat_values <- c(1, 2, 3, 4, 6, 7, 8)
total_effect_results <- list()
mediation_summary_results <- list()
sensitivity_results <- list()

# Counter to track progress
total_iterations <- length(icu_names) * length(sodium_cat_values)
current_iteration <- 0

# Apply the mediation analysis function to each ICU subset and each sodium_cat category
for (icu in icu_names) {
  for (cat_value in sodium_cat_values) {
    # Increment the progress counter
    current_iteration <- current_iteration + 1
    
    # Log progress
    message(sprintf("Starting analysis for ICU: %s, category: %d (%d/%d)", icu, cat_value, current_iteration, total_iterations))
    
    df_name <- paste0("sodium_cat_", cat_value, "_", icu, "_df")
    df <- get(df_name)
    
    # Perform the mediation analysis
    analysis_results <- perform_mediation_analysis(df, cat_value, icu)
    
    # Save the total effect results
    total_effect_results[[paste0(icu, "_cat_", cat_value)]] <- analysis_results$total_effect
    
    # Save the mediation analysis results
    mediation_summary_results[[paste0(icu, "_cat_", cat_value)]] <- analysis_results$mediation_results_sum
    
    # Save the sensitivity analysis results
    sensitivity_results[[paste0(icu, "_cat_", cat_value)]] <- analysis_results$sens_out
    
    # Log completion of the iteration
    message(sprintf("Completed analysis for ICU: %s, category: %d (%d/%d)", icu, cat_value, current_iteration, total_iterations))
  }
}

message("Mediation analysis completed for all categories and ICU types.")
```

