---
title: "Sodium GCS Mediation Mortality (Analysis)"
author: "Sicheng, Miguel"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output:
  html_notebook:
    code_folding: hide
    number_sections: yes
    theme: flatly
    toc: yes
    toc_float: yes

knit: (function(inputFile, encoding) { rmarkdown::render(inputFile, encoding = encoding, output_file = paste0(substr(inputFile,1,nchar(inputFile)-4)," ",Sys.Date(),'.html')) })
---

# Environment

```{r message=FALSE, warning=FALSE,echo = F}

library(dplyr)
library(ggplot2)
library(mediation)
library(mgcv)
library(plotly)
library(voxel)
library(htmlwidgets)
library(readr)
library(stringi)
library(mediation)
library(glmnet)
library(data.table)
library(table1)
library(cowplot)
'%!in%' <- function(x,y)!('%in%'(x,y))

# useful mediation functions

extract_mediation_summary <- function (x) { 
  clp <- 100 * x$conf.level
  isLinear.y <- ((class(x$model.y)[1] %in% c("lm", "rq")) || 
                   (inherits(x$model.y, "glm") && x$model.y$family$family == 
                      "gaussian" && x$model.y$family$link == "identity") || 
                   (inherits(x$model.y, "survreg") && x$model.y$dist == 
                      "gaussian"))
  printone <- !x$INT && isLinear.y
  if (printone) {
    smat <- c(x$d1, x$d1.ci, x$d1.p)
    smat <- rbind(smat, c(x$z0, x$z0.ci, x$z0.p))
    smat <- rbind(smat, c(x$tau.coef, x$tau.ci, x$tau.p))
    smat <- rbind(smat, c(x$n0, x$n0.ci, x$n0.p))
    rownames(smat) <- c("ACME", "ADE", "Total Effect", "Prop. Mediated")
  } else {
    smat <- c(x$d0, x$d0.ci, x$d0.p)
    smat <- rbind(smat, c(x$d1, x$d1.ci, x$d1.p))
    smat <- rbind(smat, c(x$z0, x$z0.ci, x$z0.p))
    smat <- rbind(smat, c(x$z1, x$z1.ci, x$z1.p))
    smat <- rbind(smat, c(x$tau.coef, x$tau.ci, x$tau.p))
    smat <- rbind(smat, c(x$n0, x$n0.ci, x$n0.p))
    smat <- rbind(smat, c(x$n1, x$n1.ci, x$n1.p))
    smat <- rbind(smat, c(x$d.avg, x$d.avg.ci, x$d.avg.p))
    smat <- rbind(smat, c(x$z.avg, x$z.avg.ci, x$z.avg.p))
    smat <- rbind(smat, c(x$n.avg, x$n.avg.ci, x$n.avg.p))
    rownames(smat) <- c("ACME (control)", "ACME (treated)", 
                        "ADE (control)", "ADE (treated)", "Total Effect", 
                        "Prop. Mediated (control)", "Prop. Mediated (treated)", 
                        "ACME (average)", "ADE (average)", "Prop. Mediated (average)")
  }
  colnames(smat) <- c("Estimate", paste(clp, "% CI Lower", sep = ""), 
                      paste(clp, "% CI Upper", sep = ""), "p-value")
  smat
}
```


# Mediation package

 - *ACME:*  Average Causal Mediation Effects. This is the indirect effect of the IV on the outcome that goes through the mediator.
 - *ADE:*  Average direct effects. It describes the direct effect of the IV on the Outcome.
 - *Total Effect* the Total Effect (direct + indirect) of the IV on the Outcome.
 - *Prop. Mediated* describes the proportion of the effect of the IV on the Outcome that goes through the mediator. It’s calculated by dividing the ACME through the total effect.

## Baseline sodium_cat_1

### Mediation analysis

```{r}
total_effect<-NULL
sodium_cat_1_df$sodium_cat<-as.factor(sodium_cat_1_df$sodium_cat)
sodium_cat_1_df$sodium_cat<-relevel(sodium_cat_1_df$sodium_cat,ref = "5")
fit_totaleffect<-glm(as.factor(hosp_mortality)~sodium_cat+age+AP_grp+gender+ethnicity+unittype+charlson_liver1 +charlson_chf1+mechvent_day01+respiratoryrate+MAP

                   , sodium_cat_1_df
,family = binomial)
summ<-summary(fit_totaleffect)
total_effect<-rbind(total_effect,data.frame(cat = 1,estimated = summ$coefficients[2],se = summ$coefficients[2+length(summ$coefficients)/4],p = summ$coefficients[2+3*length(summ$coefficients)/4]))



fit_mediator<-lm(gcs_baseline ~ sodium_cat+age+AP_grp+gender+ethnicity+unittype+charlson_liver1 +charlson_chf1+mechvent_day01+respiratoryrate+MAP,
data =  sodium_cat_1_df)

fit_dv<-glm(as.factor(hosp_mortality)~gcs_baseline+sodium_cat+age+AP_grp+gender+ethnicity+unittype+charlson_liver1 +charlson_chf1+mechvent_day01+respiratoryrate+MAP
                   , sodium_cat_1_df
           ,family = binomial("probit"))

summary(fit_dv)

#treat is the independend variable

mediation_results_cat1 <- mediate(fit_mediator, fit_dv, 
                                  treat="sodium_cat", mediator="gcs_baseline",control.value = 5,treat.value = 1,
                                  sims = 1000)
mediation_results_cat1_sum<-as.data.frame(extract_mediation_summary(mediation_results_cat1))
mediation_results_cat1_sum<-mediation_results_cat1_sum[c(5,8,9,10),]

#replace effect of treated group with average effect
#mediation_results_cat1_sum<-mediation_results_cat1_sum[grepl("treated|Total Effect", rownames(mediation_results_cat1_sum)), ]#

mediation_results_cat1_sum
```

### Sensitivity Analisys

```{r eval=FALSE, include=FALSE}
sens.out <- medsens(mediation_results_cat1, rho.by = 0.1, effect.type = "indirect", sims = 100)
summary(sens.out)
plot(sens.out, sens.par = "rho", main = "Sensitivity Analysis", ylim = c(-0.05, 0.05))
```

## Baseline sodium_cat_2

### Mediation analysis

```{r}
sodium_cat_2_df$sodium_cat<-as.factor(sodium_cat_2_df$sodium_cat)
sodium_cat_2_df$sodium_cat<-relevel(sodium_cat_2_df$sodium_cat,ref = "5")
fit_totaleffect<-glm(as.factor(hosp_mortality)~sodium_cat+age+AP_grp+gender+ethnicity+unittype+charlson_liver1 +charlson_chf1+mechvent_day01+respiratoryrate+MAP

                   , sodium_cat_2_df
,family = binomial)
summ<-summary(fit_totaleffect)
total_effect<-rbind(total_effect,data.frame(cat = 2,estimated = summ$coefficients[2],se = summ$coefficients[2+length(summ$coefficients)/4],p = summ$coefficients[2+3*length(summ$coefficients)/4]))



fit_mediator<-lm(gcs_baseline ~ sodium_cat+age+AP_grp+gender+ethnicity+unittype+charlson_liver1 +charlson_chf1+mechvent_day01+respiratoryrate+MAP,
data =  sodium_cat_2_df)
fit_dv<-glm(as.factor(hosp_mortality) ~gcs_baseline+sodium_cat+age+AP_grp+gender+ethnicity+unittype+charlson_liver1 +charlson_chf1+mechvent_day01+respiratoryrate+MAP
                   , sodium_cat_2_df
           ,family = binomial("probit"))
summary(fit_dv)
#treat is the independend variable
mediation_results_cat2 <- mediate(fit_mediator, fit_dv, 
                                  treat="sodium_cat", mediator="gcs_baseline",control.value = 5,treat.value = 2,
                                  sims = 1000)
mediation_results_cat2_sum<-as.data.frame(extract_mediation_summary(mediation_results_cat2))
mediation_results_cat2_sum<-mediation_results_cat2_sum[c(5,8,9,10),]

mediation_results_cat2_sum
```

### Sensitivity Analisys

```{r eval=FALSE, include=FALSE}
sens.out <- medsens(mediation_results_cat2, rho.by = 0.1, effect.type = "indirect", sims = 100)
summary(sens.out)
plot(sens.out, sens.par = "rho", main = "Sensitivity Analysis", ylim = c(-0.05, 0.05))
```

## Baseline sodium_cat_3

### Mediation analysis

```{r}
sodium_cat_3_df$sodium_cat<-as.factor(sodium_cat_3_df$sodium_cat)
sodium_cat_3_df$sodium_cat<-relevel(sodium_cat_3_df$sodium_cat,ref = "5")
fit_totaleffect<-glm(as.factor(hosp_mortality)~sodium_cat+age+AP_grp+gender+ethnicity+unittype+charlson_liver1 +charlson_chf1+mechvent_day01+respiratoryrate+MAP

                   , sodium_cat_3_df
,family = binomial)
summ<-summary(fit_totaleffect)
total_effect<-rbind(total_effect,data.frame(cat = 3,estimated = summ$coefficients[2],se = summ$coefficients[2+length(summ$coefficients)/4],p = summ$coefficients[2+3*length(summ$coefficients)/4]))

fit_mediator<-lm(gcs_baseline ~ sodium_cat+age+AP_grp+gender+ethnicity+unittype+charlson_liver1 +charlson_chf1+mechvent_day01+respiratoryrate+MAP,
data =  sodium_cat_3_df)
fit_dv<-glm(as.factor(hosp_mortality) ~gcs_baseline+sodium_cat+age+AP_grp+gender+ethnicity+unittype+charlson_liver1 +charlson_chf1+mechvent_day01+respiratoryrate+MAP
                   , sodium_cat_3_df
           ,family = binomial("probit"))
summary(fit_dv)
#treat is the independend variable
mediation_results_cat3 <- mediate(fit_mediator, fit_dv, 
                                  treat="sodium_cat", mediator="gcs_baseline",control.value = 5,treat.value = 3,
                                  sims = 1000)
mediation_results_cat3_sum<-as.data.frame(extract_mediation_summary(mediation_results_cat3))
mediation_results_cat3_sum<-mediation_results_cat3_sum[c(5,8,9,10),]

mediation_results_cat3_sum
```

### Sensitivity Analisys

```{r eval=FALSE, include=FALSE}
sens.out <- medsens(mediation_results_cat3, rho.by = 0.1, effect.type = "indirect", sims = 100)
summary(sens.out)
plot(sens.out, sens.par = "rho", main = "Sensitivity Analysis", ylim = c(-0.05, 0.05))
```

## Baseline sodium_cat_4

### Mediation analysis

```{r}
sodium_cat_4_df$sodium_cat<-as.factor(sodium_cat_4_df$sodium_cat)
sodium_cat_4_df$sodium_cat<-relevel(sodium_cat_4_df$sodium_cat,ref = "5")
fit_totaleffect<-glm(as.factor(hosp_mortality)~sodium_cat+age+AP_grp+gender+ethnicity+unittype+charlson_liver1 +charlson_chf1+mechvent_day01+respiratoryrate+MAP

                   , sodium_cat_4_df
,family = binomial)
summ<-summary(fit_totaleffect)
total_effect<-rbind(total_effect,data.frame(cat = 4,estimated = summ$coefficients[2],se = summ$coefficients[2+length(summ$coefficients)/4],p = summ$coefficients[2+3*length(summ$coefficients)/4]))


fit_mediator<-lm(gcs_baseline ~ sodium_cat+age+AP_grp+gender+ethnicity+unittype+charlson_liver1 +charlson_chf1+mechvent_day01+respiratoryrate+MAP,
data =  sodium_cat_4_df)
fit_dv<-glm(as.factor(hosp_mortality) ~gcs_baseline+sodium_cat+age+AP_grp+gender+ethnicity+unittype+charlson_liver1 +charlson_chf1+mechvent_day01+respiratoryrate+MAP
                   , sodium_cat_4_df
           ,family = binomial("probit"))
summary(fit_dv)
#treat is the independend variable
mediation_results_cat4 <- mediate(fit_mediator, fit_dv, 
                                  treat="sodium_cat", mediator="gcs_baseline",control.value = 5,treat.value = 4,
                                  sims = 1000)
mediation_results_cat4_sum<-as.data.frame(extract_mediation_summary(mediation_results_cat4))
mediation_results_cat4_sum<-mediation_results_cat4_sum[c(5,8,9,10),]
mediation_results_cat4_sum

```

### Sensitivity Analisys

```{r eval=FALSE, include=FALSE}
sens.out <- medsens(mediation_results_cat4, rho.by = 0.1, effect.type = "indirect", sims = 100)
summary(sens.out)
plot(sens.out, sens.par = "rho", main = "Sensitivity Analysis", ylim = c(-0.05, 0.05))
```
## Baseline sodium_cat_6

### Mediation analysis

```{r}
sodium_cat_6_df$sodium_cat<-as.factor(sodium_cat_6_df$sodium_cat)
sodium_cat_6_df$sodium_cat<-relevel(sodium_cat_6_df$sodium_cat,ref = "5")
fit_totaleffect<-glm(as.factor(hosp_mortality)~sodium_cat+age+AP_grp+gender+ethnicity+unittype+charlson_liver1 +charlson_chf1+mechvent_day01+respiratoryrate+MAP

                   , sodium_cat_6_df
,family = binomial)
summ<-summary(fit_totaleffect)
total_effect<-rbind(total_effect,data.frame(cat = 6,estimated = summ$coefficients[2],se = summ$coefficients[2+length(summ$coefficients)/4],p = summ$coefficients[2+3*length(summ$coefficients)/4]))


fit_mediator<-lm(gcs_baseline ~ sodium_cat+age+AP_grp+gender+ethnicity+unittype+charlson_liver1 +charlson_chf1+mechvent_day01+respiratoryrate+MAP,
data =  sodium_cat_6_df)
fit_dv<-glm(as.factor(hosp_mortality) ~gcs_baseline+sodium_cat+age+AP_grp+gender+ethnicity+unittype+charlson_liver1 +charlson_chf1+mechvent_day01+respiratoryrate+MAP
                   , sodium_cat_6_df
           ,family = binomial("probit"))
summary(fit_dv)
#treat is the independend variable
mediation_results_cat6 <- mediate(fit_mediator, fit_dv, 
                                  treat="sodium_cat", mediator="gcs_baseline",control.value = 5,treat.value = 6,
                                  sims = 1000)
mediation_results_cat6_sum<-as.data.frame(extract_mediation_summary(mediation_results_cat6))
mediation_results_cat6_sum<-mediation_results_cat6_sum[c(5,8,9,10),]
mediation_results_cat6_sum
```

### Sensitivity Analisys

```{r eval=FALSE, include=FALSE}
sens.out <- medsens(mediation_results_cat6, rho.by = 0.1, effect.type = "indirect", sims = 100)
summary(sens.out)
plot(sens.out, sens.par = "rho", main = "Sensitivity Analysis", ylim = c(-0.05, 0.05))
```

## Baseline sodium_cat_7

### Mediation analysis

```{r}
sodium_cat_7_df$sodium_cat<-as.factor(sodium_cat_7_df$sodium_cat)
sodium_cat_7_df$sodium_cat<-relevel(sodium_cat_7_df$sodium_cat,ref = "5")
fit_totaleffect<-glm(as.factor(hosp_mortality)~sodium_cat+age+AP_grp+gender+ethnicity+unittype+charlson_liver1 +charlson_chf1+mechvent_day01+respiratoryrate+MAP

                   , sodium_cat_7_df
,family = binomial)
summ<-summary(fit_totaleffect)
total_effect<-rbind(total_effect,data.frame(cat = 7,estimated = summ$coefficients[2],se = summ$coefficients[2+length(summ$coefficients)/4],p = summ$coefficients[2+3*length(summ$coefficients)/4]))


fit_mediator<-lm(gcs_baseline ~ sodium_cat+age+AP_grp+gender+ethnicity+unittype+charlson_liver1 +charlson_chf1+mechvent_day01+respiratoryrate+MAP,
data =  sodium_cat_7_df)
fit_dv<-glm(as.factor(hosp_mortality) ~gcs_baseline+sodium_cat+age+AP_grp+gender+ethnicity+unittype+charlson_liver1 +charlson_chf1+mechvent_day01+respiratoryrate+MAP
                   , sodium_cat_7_df
           ,family = binomial("probit"))
summary(fit_dv)
#treat is the independend variable
mediation_results_cat7 <- mediate(fit_mediator, fit_dv, 
                                  treat="sodium_cat", mediator="gcs_baseline",control.value = 5,treat.value = 7,
                                  sims = 1000)
mediation_results_cat7_sum<-as.data.frame(extract_mediation_summary(mediation_results_cat7))
mediation_results_cat7_sum<-mediation_results_cat7_sum[c(5,8,9,10),]
mediation_results_cat7_sum
```

### Sensitivity Analisys

```{r eval=FALSE, include=FALSE}
sens.out <- medsens(mediation_results_cat7, rho.by = 0.1, effect.type = "indirect", sims = 100)
summary(sens.out)
plot(sens.out, sens.par = "rho", main = "Sensitivity Analysis", ylim = c(-0.05, 0.05))
```

## Baseline sodium_cat_8

### Mediation analysis

```{r}
sodium_cat_8_df$sodium_cat<-as.factor(sodium_cat_8_df$sodium_cat)
sodium_cat_8_df$sodium_cat<-relevel(sodium_cat_8_df$sodium_cat,ref = "5")
fit_totaleffect<-glm(as.factor(hosp_mortality)~sodium_cat+age+AP_grp+gender+ethnicity+unittype+charlson_liver1 +charlson_chf1+mechvent_day01+respiratoryrate+MAP

                   , sodium_cat_8_df
,family = binomial)
summ<-summary(fit_totaleffect)
total_effect<-rbind(total_effect,data.frame(cat = 8,estimated = summ$coefficients[2],se = summ$coefficients[2+length(summ$coefficients)/4],p = summ$coefficients[2+3*length(summ$coefficients)/4]))


fit_mediator<-lm(gcs_baseline ~ sodium_cat+age+AP_grp+gender+ethnicity+unittype+charlson_liver1 +charlson_chf1+mechvent_day01+respiratoryrate+MAP,
data =  sodium_cat_8_df)
fit_dv<-glm(as.factor(hosp_mortality) ~gcs_baseline+sodium_cat+age+AP_grp+gender+ethnicity+unittype+charlson_liver1 +charlson_chf1+mechvent_day01+respiratoryrate+MAP
                   , sodium_cat_8_df
           ,family = binomial("probit"))
summary(fit_dv)
#treat is the independend variable
mediation_results_cat8 <- mediate(fit_mediator, fit_dv, 
                                  treat="sodium_cat", mediator="gcs_baseline",control.value = 5,treat.value = 8,
                                  sims = 1000)
mediation_results_cat8_sum<-as.data.frame(extract_mediation_summary(mediation_results_cat8))
mediation_results_cat8_sum<-mediation_results_cat8_sum[c(5,8,9,10),]
mediation_results_cat8_sum
```

### Sensitivity Analisys

```{r eval=FALSE, include=FALSE}
sens.out <- medsens(mediation_results_cat8, rho.by = 0.1, effect.type = "indirect", sims = 100)
summary(sens.out)
plot(sens.out, sens.par = "rho", main = "Sensitivity Analysis", ylim = c(-0.05, 0.05))
```

