---
title: "Sodium GCS Mediation Mortality (Analysis)"
author: "Sicheng, Miguel"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output:
  html_notebook:
    code_folding: hide
    number_sections: yes
    theme: flatly
    toc: yes
    toc_float: yes

knit: (function(inputFile, encoding) { rmarkdown::render(inputFile, encoding = encoding, output_file = paste0(substr(inputFile,1,nchar(inputFile)-4)," ",Sys.Date(),'.html')) })
---

# Environment

```{r message=FALSE, warning=FALSE,echo = F}

library(dplyr)
library(ggplot2)
library(mediation)
library(mgcv)
library(plotly)
library(voxel)
library(htmlwidgets)
library(readr)
library(stringi)
library(mediation)
library(glmnet)
library(data.table)
library(table1)
library(cowplot)
'%!in%' <- function(x,y)!('%in%'(x,y))
extract_mediation_summary <- function (x) { 
  clp <- 100 * x$conf.level
  isLinear.y <- ((class(x$model.y)[1] %in% c("lm", "rq")) || 
                   (inherits(x$model.y, "glm") && x$model.y$family$family == 
                      "gaussian" && x$model.y$family$link == "identity") || 
                   (inherits(x$model.y, "survreg") && x$model.y$dist == 
                      "gaussian"))
  printone <- !x$INT && isLinear.y
  if (printone) {
    smat <- c(x$d1, x$d1.ci, x$d1.p)
    smat <- rbind(smat, c(x$z0, x$z0.ci, x$z0.p))
    smat <- rbind(smat, c(x$tau.coef, x$tau.ci, x$tau.p))
    smat <- rbind(smat, c(x$n0, x$n0.ci, x$n0.p))
    rownames(smat) <- c("ACME", "ADE", "Total Effect", "Prop. Mediated")
  } else {
    smat <- c(x$d0, x$d0.ci, x$d0.p)
    smat <- rbind(smat, c(x$d1, x$d1.ci, x$d1.p))
    smat <- rbind(smat, c(x$z0, x$z0.ci, x$z0.p))
    smat <- rbind(smat, c(x$z1, x$z1.ci, x$z1.p))
    smat <- rbind(smat, c(x$tau.coef, x$tau.ci, x$tau.p))
    smat <- rbind(smat, c(x$n0, x$n0.ci, x$n0.p))
    smat <- rbind(smat, c(x$n1, x$n1.ci, x$n1.p))
    smat <- rbind(smat, c(x$d.avg, x$d.avg.ci, x$d.avg.p))
    smat <- rbind(smat, c(x$z.avg, x$z.avg.ci, x$z.avg.p))
    smat <- rbind(smat, c(x$n.avg, x$n.avg.ci, x$n.avg.p))
    rownames(smat) <- c("ACME (control)", "ACME (treated)", 
                        "ADE (control)", "ADE (treated)", "Total Effect", 
                        "Prop. Mediated (control)", "Prop. Mediated (treated)", 
                        "ACME (average)", "ADE (average)", "Prop. Mediated (average)")
  }
  colnames(smat) <- c("Estimate", paste(clp, "% CI Lower", sep = ""), 
                      paste(clp, "% CI Upper", sep = ""), "p-value")
  smat
}


data$sodium_cat<-as.numeric(as.character(data$sodium_cat))
sodium_cat_1_df<-data%>%filter(sodium_cat%in%c(1,5))
sodium_cat_2_df<-data%>%filter(sodium_cat%in%c(2,5))
sodium_cat_3_df<-data%>%filter(sodium_cat%in%c(3,5))
sodium_cat_4_df<-data%>%filter(sodium_cat%in%c(4,5))
sodium_cat_6_df<-data%>%filter(sodium_cat%in%c(6,5))
sodium_cat_7_df<-data%>%filter(sodium_cat%in%c(7,5))
sodium_cat_8_df<-data%>%filter(sodium_cat%in%c(8,5))



```


# Mediation package

 - *ACME:*  Average Causal Mediation Effects. This is the indirect effect of the IV on the outcome that goes through the mediator.
 - *ADE:*  Average direct effects. It describes the direct effect of the IV on the Outcome.
 - *Total Effect* the Total Effect (direct + indirect) of the IV on the Outcome.
 - *Prop. Mediated* describes the proportion of the effect of the IV on the Outcome that goes through the mediator. It’s calculated by dividing the ACME through the total effect.

## Baseline sodium_cat_1

### Mediation analysis

```{r}
total_effect<-NULL
sodium_cat_1_df$sodium_cat<-as.factor(sodium_cat_1_df$sodium_cat)
sodium_cat_1_df$sodium_cat<-relevel(sodium_cat_1_df$sodium_cat,ref = "5")
fit_totaleffect<-glm(as.factor(hosp_mortality)~sodium_cat+age+AP_grp+gender+ethnicity+unittype+charlson_liver1 +charlson_chf1+mechvent_day01+respiratoryrate+MAP

                   , sodium_cat_1_df
,family = binomial)
summ<-summary(fit_totaleffect)
total_effect<-rbind(total_effect,data.frame(cat = 1,estimated = summ$coefficients[2],se = summ$coefficients[2+length(summ$coefficients)/4],p = summ$coefficients[2+3*length(summ$coefficients)/4]))



fit_mediator<-lm(gcs_baseline ~ sodium_cat+age+AP_grp+gender+ethnicity+unittype+charlson_liver1 +charlson_chf1+mechvent_day01+respiratoryrate+MAP,
data =  sodium_cat_1_df)

fit_dv<-glm(as.factor(hosp_mortality)~gcs_baseline+sodium_cat+age+AP_grp+gender+ethnicity+unittype+charlson_liver1 +charlson_chf1+mechvent_day01+respiratoryrate+MAP
                   , sodium_cat_1_df
           ,family = binomial("probit"))

summary(fit_dv)

#treat is the independend variable

mediation_results_cat1 <- mediate(fit_mediator, fit_dv, 
                                  treat="sodium_cat", mediator="gcs_baseline",control.value = 5,treat.value = 1,
                                  sims = 1000)
mediation_results_cat1_sum<-as.data.frame(extract_mediation_summary(mediation_results_cat1))
mediation_results_cat1_sum<-mediation_results_cat1_sum[c(5,8,9,10),]

#replace effect of treated group with average effect
#mediation_results_cat1_sum<-mediation_results_cat1_sum[grepl("treated|Total Effect", rownames(mediation_results_cat1_sum)), ]#

mediation_results_cat1_sum
```

### Sensitivity Analisys

```{r eval=FALSE, include=FALSE}
sens.out <- medsens(mediation_results_cat1, rho.by = 0.1, effect.type = "indirect", sims = 100)
summary(sens.out)
plot(sens.out, sens.par = "rho", main = "Sensitivity Analysis", ylim = c(-0.05, 0.05))
```

## Baseline sodium_cat_2

### Mediation analysis

```{r}
sodium_cat_2_df$sodium_cat<-as.factor(sodium_cat_2_df$sodium_cat)
sodium_cat_2_df$sodium_cat<-relevel(sodium_cat_2_df$sodium_cat,ref = "5")
fit_totaleffect<-glm(as.factor(hosp_mortality)~sodium_cat+age+AP_grp+gender+ethnicity+unittype+charlson_liver1 +charlson_chf1+mechvent_day01+respiratoryrate+MAP

                   , sodium_cat_2_df
,family = binomial)
summ<-summary(fit_totaleffect)
total_effect<-rbind(total_effect,data.frame(cat = 2,estimated = summ$coefficients[2],se = summ$coefficients[2+length(summ$coefficients)/4],p = summ$coefficients[2+3*length(summ$coefficients)/4]))



fit_mediator<-lm(gcs_baseline ~ sodium_cat+age+AP_grp+gender+ethnicity+unittype+charlson_liver1 +charlson_chf1+mechvent_day01+respiratoryrate+MAP,
data =  sodium_cat_2_df)
fit_dv<-glm(as.factor(hosp_mortality) ~gcs_baseline+sodium_cat+age+AP_grp+gender+ethnicity+unittype+charlson_liver1 +charlson_chf1+mechvent_day01+respiratoryrate+MAP
                   , sodium_cat_2_df
           ,family = binomial("probit"))
summary(fit_dv)
#treat is the independend variable
mediation_results_cat2 <- mediate(fit_mediator, fit_dv, 
                                  treat="sodium_cat", mediator="gcs_baseline",control.value = 5,treat.value = 2,
                                  sims = 1000)
mediation_results_cat2_sum<-as.data.frame(extract_mediation_summary(mediation_results_cat2))
mediation_results_cat2_sum<-mediation_results_cat2_sum[c(5,8,9,10),]

mediation_results_cat2_sum
```

### Sensitivity Analisys

```{r eval=FALSE, include=FALSE}
sens.out <- medsens(mediation_results_cat2, rho.by = 0.1, effect.type = "indirect", sims = 100)
summary(sens.out)
plot(sens.out, sens.par = "rho", main = "Sensitivity Analysis", ylim = c(-0.05, 0.05))
```

## Baseline sodium_cat_3

### Mediation analysis

```{r}
sodium_cat_3_df$sodium_cat<-as.factor(sodium_cat_3_df$sodium_cat)
sodium_cat_3_df$sodium_cat<-relevel(sodium_cat_3_df$sodium_cat,ref = "5")
fit_totaleffect<-glm(as.factor(hosp_mortality)~sodium_cat+age+AP_grp+gender+ethnicity+unittype+charlson_liver1 +charlson_chf1+mechvent_day01+respiratoryrate+MAP

                   , sodium_cat_3_df
,family = binomial)
summ<-summary(fit_totaleffect)
total_effect<-rbind(total_effect,data.frame(cat = 3,estimated = summ$coefficients[2],se = summ$coefficients[2+length(summ$coefficients)/4],p = summ$coefficients[2+3*length(summ$coefficients)/4]))

fit_mediator<-lm(gcs_baseline ~ sodium_cat+age+AP_grp+gender+ethnicity+unittype+charlson_liver1 +charlson_chf1+mechvent_day01+respiratoryrate+MAP,
data =  sodium_cat_3_df)
fit_dv<-glm(as.factor(hosp_mortality) ~gcs_baseline+sodium_cat+age+AP_grp+gender+ethnicity+unittype+charlson_liver1 +charlson_chf1+mechvent_day01+respiratoryrate+MAP
                   , sodium_cat_3_df
           ,family = binomial("probit"))
summary(fit_dv)
#treat is the independend variable
mediation_results_cat3 <- mediate(fit_mediator, fit_dv, 
                                  treat="sodium_cat", mediator="gcs_baseline",control.value = 5,treat.value = 3,
                                  sims = 1000)
mediation_results_cat3_sum<-as.data.frame(extract_mediation_summary(mediation_results_cat3))
mediation_results_cat3_sum<-mediation_results_cat3_sum[c(5,8,9,10),]

mediation_results_cat3_sum
```

### Sensitivity Analisys

```{r eval=FALSE, include=FALSE}
sens.out <- medsens(mediation_results_cat3, rho.by = 0.1, effect.type = "indirect", sims = 100)
summary(sens.out)
plot(sens.out, sens.par = "rho", main = "Sensitivity Analysis", ylim = c(-0.05, 0.05))
```

## Baseline sodium_cat_4

### Mediation analysis

```{r}
sodium_cat_4_df$sodium_cat<-as.factor(sodium_cat_4_df$sodium_cat)
sodium_cat_4_df$sodium_cat<-relevel(sodium_cat_4_df$sodium_cat,ref = "5")
fit_totaleffect<-glm(as.factor(hosp_mortality)~sodium_cat+age+AP_grp+gender+ethnicity+unittype+charlson_liver1 +charlson_chf1+mechvent_day01+respiratoryrate+MAP

                   , sodium_cat_4_df
,family = binomial)
summ<-summary(fit_totaleffect)
total_effect<-rbind(total_effect,data.frame(cat = 4,estimated = summ$coefficients[2],se = summ$coefficients[2+length(summ$coefficients)/4],p = summ$coefficients[2+3*length(summ$coefficients)/4]))


fit_mediator<-lm(gcs_baseline ~ sodium_cat+age+AP_grp+gender+ethnicity+unittype+charlson_liver1 +charlson_chf1+mechvent_day01+respiratoryrate+MAP,
data =  sodium_cat_4_df)
fit_dv<-glm(as.factor(hosp_mortality) ~gcs_baseline+sodium_cat+age+AP_grp+gender+ethnicity+unittype+charlson_liver1 +charlson_chf1+mechvent_day01+respiratoryrate+MAP
                   , sodium_cat_4_df
           ,family = binomial("probit"))
summary(fit_dv)
#treat is the independend variable
mediation_results_cat4 <- mediate(fit_mediator, fit_dv, 
                                  treat="sodium_cat", mediator="gcs_baseline",control.value = 5,treat.value = 4,
                                  sims = 1000)
mediation_results_cat4_sum<-as.data.frame(extract_mediation_summary(mediation_results_cat4))
mediation_results_cat4_sum<-mediation_results_cat4_sum[c(5,8,9,10),]
mediation_results_cat4_sum

```

### Sensitivity Analisys

```{r eval=FALSE, include=FALSE}
sens.out <- medsens(mediation_results_cat4, rho.by = 0.1, effect.type = "indirect", sims = 100)
summary(sens.out)
plot(sens.out, sens.par = "rho", main = "Sensitivity Analysis", ylim = c(-0.05, 0.05))
```
## Baseline sodium_cat_6

### Mediation analysis

```{r}
sodium_cat_6_df$sodium_cat<-as.factor(sodium_cat_6_df$sodium_cat)
sodium_cat_6_df$sodium_cat<-relevel(sodium_cat_6_df$sodium_cat,ref = "5")
fit_totaleffect<-glm(as.factor(hosp_mortality)~sodium_cat+age+AP_grp+gender+ethnicity+unittype+charlson_liver1 +charlson_chf1+mechvent_day01+respiratoryrate+MAP

                   , sodium_cat_6_df
,family = binomial)
summ<-summary(fit_totaleffect)
total_effect<-rbind(total_effect,data.frame(cat = 6,estimated = summ$coefficients[2],se = summ$coefficients[2+length(summ$coefficients)/4],p = summ$coefficients[2+3*length(summ$coefficients)/4]))


fit_mediator<-lm(gcs_baseline ~ sodium_cat+age+AP_grp+gender+ethnicity+unittype+charlson_liver1 +charlson_chf1+mechvent_day01+respiratoryrate+MAP,
data =  sodium_cat_6_df)
fit_dv<-glm(as.factor(hosp_mortality) ~gcs_baseline+sodium_cat+age+AP_grp+gender+ethnicity+unittype+charlson_liver1 +charlson_chf1+mechvent_day01+respiratoryrate+MAP
                   , sodium_cat_6_df
           ,family = binomial("probit"))
summary(fit_dv)
#treat is the independend variable
mediation_results_cat6 <- mediate(fit_mediator, fit_dv, 
                                  treat="sodium_cat", mediator="gcs_baseline",control.value = 5,treat.value = 6,
                                  sims = 1000)
mediation_results_cat6_sum<-as.data.frame(extract_mediation_summary(mediation_results_cat6))
mediation_results_cat6_sum<-mediation_results_cat6_sum[c(5,8,9,10),]
mediation_results_cat6_sum
```

### Sensitivity Analisys

```{r eval=FALSE, include=FALSE}
sens.out <- medsens(mediation_results_cat6, rho.by = 0.1, effect.type = "indirect", sims = 100)
summary(sens.out)
plot(sens.out, sens.par = "rho", main = "Sensitivity Analysis", ylim = c(-0.05, 0.05))
```

## Baseline sodium_cat_7

### Mediation analysis

```{r}
sodium_cat_7_df$sodium_cat<-as.factor(sodium_cat_7_df$sodium_cat)
sodium_cat_7_df$sodium_cat<-relevel(sodium_cat_7_df$sodium_cat,ref = "5")
fit_totaleffect<-glm(as.factor(hosp_mortality)~sodium_cat+age+AP_grp+gender+ethnicity+unittype+charlson_liver1 +charlson_chf1+mechvent_day01+respiratoryrate+MAP

                   , sodium_cat_7_df
,family = binomial)
summ<-summary(fit_totaleffect)
total_effect<-rbind(total_effect,data.frame(cat = 7,estimated = summ$coefficients[2],se = summ$coefficients[2+length(summ$coefficients)/4],p = summ$coefficients[2+3*length(summ$coefficients)/4]))


fit_mediator<-lm(gcs_baseline ~ sodium_cat+age+AP_grp+gender+ethnicity+unittype+charlson_liver1 +charlson_chf1+mechvent_day01+respiratoryrate+MAP,
data =  sodium_cat_7_df)
fit_dv<-glm(as.factor(hosp_mortality) ~gcs_baseline+sodium_cat+age+AP_grp+gender+ethnicity+unittype+charlson_liver1 +charlson_chf1+mechvent_day01+respiratoryrate+MAP
                   , sodium_cat_7_df
           ,family = binomial("probit"))
summary(fit_dv)
#treat is the independend variable
mediation_results_cat7 <- mediate(fit_mediator, fit_dv, 
                                  treat="sodium_cat", mediator="gcs_baseline",control.value = 5,treat.value = 7,
                                  sims = 1000)
mediation_results_cat7_sum<-as.data.frame(extract_mediation_summary(mediation_results_cat7))
mediation_results_cat7_sum<-mediation_results_cat7_sum[c(5,8,9,10),]
mediation_results_cat7_sum
```

### Sensitivity Analisys

```{r eval=FALSE, include=FALSE}
sens.out <- medsens(mediation_results_cat7, rho.by = 0.1, effect.type = "indirect", sims = 100)
summary(sens.out)
plot(sens.out, sens.par = "rho", main = "Sensitivity Analysis", ylim = c(-0.05, 0.05))
```

## Baseline sodium_cat_8

### Mediation analysis

```{r}
sodium_cat_8_df$sodium_cat<-as.factor(sodium_cat_8_df$sodium_cat)
sodium_cat_8_df$sodium_cat<-relevel(sodium_cat_8_df$sodium_cat,ref = "5")
fit_totaleffect<-glm(as.factor(hosp_mortality)~sodium_cat+age+AP_grp+gender+ethnicity+unittype+charlson_liver1 +charlson_chf1+mechvent_day01+respiratoryrate+MAP

                   , sodium_cat_8_df
,family = binomial)
summ<-summary(fit_totaleffect)
total_effect<-rbind(total_effect,data.frame(cat = 8,estimated = summ$coefficients[2],se = summ$coefficients[2+length(summ$coefficients)/4],p = summ$coefficients[2+3*length(summ$coefficients)/4]))


fit_mediator<-lm(gcs_baseline ~ sodium_cat+age+AP_grp+gender+ethnicity+unittype+charlson_liver1 +charlson_chf1+mechvent_day01+respiratoryrate+MAP,
data =  sodium_cat_8_df)
fit_dv<-glm(as.factor(hosp_mortality) ~gcs_baseline+sodium_cat+age+AP_grp+gender+ethnicity+unittype+charlson_liver1 +charlson_chf1+mechvent_day01+respiratoryrate+MAP
                   , sodium_cat_8_df
           ,family = binomial("probit"))
summary(fit_dv)
#treat is the independend variable
mediation_results_cat8 <- mediate(fit_mediator, fit_dv, 
                                  treat="sodium_cat", mediator="gcs_baseline",control.value = 5,treat.value = 8,
                                  sims = 1000)
mediation_results_cat8_sum<-as.data.frame(extract_mediation_summary(mediation_results_cat8))
mediation_results_cat8_sum<-mediation_results_cat8_sum[c(5,8,9,10),]
mediation_results_cat8_sum
```

### Sensitivity Analisys

```{r eval=FALSE, include=FALSE}
sens.out <- medsens(mediation_results_cat8, rho.by = 0.1, effect.type = "indirect", sims = 100)
summary(sens.out)
plot(sens.out, sens.par = "rho", main = "Sensitivity Analysis", ylim = c(-0.05, 0.05))
```

# Ploting results

## Appending results

 - *ACME:*  Average Causal Mediation Effects. This is the indirect effect of the IV on the outcome that goes through the mediator.
 - *ADE:*  Average direct effects. It describes the direct effect of the IV on the Outcome.
 - *Total Effect* the Total Effect (direct + indirect) of the IV on the Outcome.
 - *Prop. Mediated* describes the proportion of the effect of the IV on the Outcome that goes through the mediator. It’s calculated by dividing the ACME through the total effect.

```{r}
# we want to be able to know which dataset our results are coming from
mediation_results_cat1_sum$cat<-1
mediation_results_cat1_sum$results<-stri_replace_all_fixed(rownames(mediation_results_cat1_sum),'(treated)',"" )
mediation_results_cat2_sum$cat<-2
mediation_results_cat2_sum$results<-stri_replace_all_fixed(rownames(mediation_results_cat2_sum),'(treated)',"" )
mediation_results_cat3_sum$cat<-3
mediation_results_cat3_sum$results<-stri_replace_all_fixed(rownames(mediation_results_cat3_sum),'(treated)',"" )
mediation_results_cat4_sum$cat<-4
mediation_results_cat4_sum$results<-stri_replace_all_fixed(rownames(mediation_results_cat4_sum),'(treated)',"" )
mediation_results_cat6_sum$cat<-6
mediation_results_cat6_sum$results<-stri_replace_all_fixed(rownames(mediation_results_cat6_sum),'(treated)',"" )
mediation_results_cat7_sum$cat<-7
mediation_results_cat7_sum$results<-stri_replace_all_fixed(rownames(mediation_results_cat7_sum),'(treated)',"" )
mediation_results_cat8_sum$cat<-8
mediation_results_cat8_sum$results<-stri_replace_all_fixed(rownames(mediation_results_cat8_sum),'(treated)',"" )
mediation_results_all<-rbind(
 mediation_results_cat1_sum
,mediation_results_cat2_sum
,mediation_results_cat3_sum
,mediation_results_cat4_sum
,mediation_results_cat6_sum
,mediation_results_cat7_sum
,mediation_results_cat8_sum
)
rownames(mediation_results_all)<-NULL

#add total effect (Odds ratio) to the table

total_effect<-rbind(total_effect,data.frame(cat = 5,estimated = 0,se = 0,p = NA))

total_effect<-total_effect%>%
  mutate(lb = estimated -1.96*se, ub = estimated + 1.96*se)%>%
  mutate(or = exp(estimated),or.lb = exp(lb),or.ub = exp(ub))

total_effect_copy<-total_effect%>%transmute(Estimate = or, '95% CI Upper' = or.ub,'95% CI Lower' = or.lb,cat = cat,'p-value' = p,results = "Total Effect (OR)")

results_overall<-rbind(mediation_results_all,total_effect_copy)

#fwrite(results_overall,file = "results_table.csv")
```

## Final Plot

## Prop. Mediated

```{r}
mediation_results_filtered<-mediation_results_all%>% filter(results %in% c('Prop. Mediated '))
ggplot(data=mediation_results_filtered,aes(x=cat,y=Estimate,colour=results,fill=results))+
    geom_point()+
    geom_line()+
    geom_errorbar(data=mediation_results_filtered
                  ,aes(x=cat,ymin=`95% CI Lower`,ymax=`95% CI Upper`,fill=results)
                  ,alpha=0.3
                  ,inherit.aes=FALSE
                  ,width=0.2)+
    xlab('Baseline Sodium Category')+
    ylab('Estimate')+
    #scale_color_manual(labels = c("Male", "Female"), values = c("#1abc9c","#f1c40f"))+
    #scale_fill_manual(labels = c("Male", "Female"), values = c("#1abc9c","#f1c40f"))+
    #labs(colour='sex', fill='sex') +
    theme_minimal()+
  ggtitle('Proportion Mediated')+
  scale_x_continuous(breaks = c(1,2,3,4,6,7,8))
```

### Indirect Effect

```{r}
mediation_results_filtered<-mediation_results_all%>% filter(results %in% c('ACME '))
ggplot(data=mediation_results_filtered,aes(x=cat,y=Estimate,colour=results,fill=results))+
    geom_point()+
    geom_line()+
    geom_errorbar(data=mediation_results_filtered
                  ,aes(x=cat,ymin=`95% CI Lower`,ymax=`95% CI Upper`,fill=results)
                  ,alpha=0.3
                  ,inherit.aes=FALSE
                  ,width=0.2)+
    xlab('Baseline Sodium Category')+
    ylab('Estimate')+
    #scale_color_manual(labels = c("Male", "Female"), values = c("#1abc9c","#f1c40f"))+
    #scale_fill_manual(labels = c("Male", "Female"), values = c("#1abc9c","#f1c40f"))+
    #labs(colour='sex', fill='sex') +
    theme_minimal()+
  ggtitle('ACME\nAverage Causal Mediation Effects\nIndirect effect of Sodium on mortality that goes through the mediator (GCS)')+
  scale_x_continuous(breaks = c(1,2,3,4,6,7,8))
```


# Direct Effect

```{r}
mediation_results_filtered<-mediation_results_all%>% filter(results %in% c('ADE '))
ggplot(data=mediation_results_filtered,aes(x=cat,y=Estimate,colour=results,fill=results))+
    geom_point()+
    geom_line()+
    geom_errorbar(data=mediation_results_filtered
                  ,aes(x=cat,ymin=`95% CI Lower`,ymax=`95% CI Upper`,fill=results)
                  ,alpha=0.3
                  ,inherit.aes=FALSE
                  ,width=0.2)+
    xlab('Baseline Sodium Category')+
    ylab('Estimate')+
    #scale_color_manual(labels = c("Male", "Female"), values = c("#1abc9c","#f1c40f"))+
    #scale_fill_manual(labels = c("Male", "Female"), values = c("#1abc9c","#f1c40f"))+
    #labs(colour='sex', fill='sex') +
    theme_minimal()+
  ggtitle("Average Direct Effect")+
  scale_x_continuous(breaks = c(1,2,3,4,6,7,8))
```

# Total Effect from Mediation Package

```{r}
mediation_results_filtered<-mediation_results_all%>% filter(results %in% c('Total Effect'))
ggplot(data=mediation_results_filtered,aes(x=cat,y=Estimate,colour=results,fill=results))+
    geom_point()+
    geom_line()+
    geom_errorbar(data=mediation_results_filtered
                  ,aes(x=cat,ymin=`95% CI Lower`,ymax=`95% CI Upper`,fill=results)
                  ,alpha=0.3
                  ,inherit.aes=FALSE
                  ,width=0.2)+
    xlab('Baseline Sodium Category')+
    ylab('Estimate')+
    #scale_color_manual(labels = c("Male", "Female"), values = c("#1abc9c","#f1c40f"))+
    #scale_fill_manual(labels = c("Male", "Female"), values = c("#1abc9c","#f1c40f"))+
    #labs(colour='sex', fill='sex') +
    theme_minimal()+
  ggtitle('Total Effect')+
  scale_x_continuous(breaks = c(1,2,3,4,6,7,8))
```

# Total Effect in Odds Ratio

```{r}

ggplot(data=total_effect,aes(x=cat,y=or))+
    geom_point()+
    geom_line(aes(color = "red"))+
    geom_errorbar(data=total_effect
                  ,aes(x=cat,ymin=or.lb,ymax=or.ub)
                  ,alpha=0.3
                  ,inherit.aes=FALSE
                  ,width=0.2)+
    xlab('Baseline Sodium Category')+
    ylab('Estimate')+
    #scale_color_manual(labels = c("Male", "Female"), values = c("#1abc9c","#f1c40f"))+
    #scale_fill_manual(labels = c("Male", "Female"), values = c("#1abc9c","#f1c40f"))+
    #labs(colour='sex', fill='sex') +
    theme_minimal()+
  ggtitle('Total Effect')+
  scale_x_continuous(breaks = c(1,2,3,4,5,6,7,8))
```

# Combined Plots

We are going to plot the OR of the estimates. The most interesting metrics to plot are Total Effect and Proportion Mediated

## Prop. Mediated + Total Effect

```{r fig.height=9, fig.width=7}
mediation_results_propmediated<-mediation_results_all%>%
  filter(results %in% c("Prop. Mediated ") ) %>%
  select(Estimate,cat,results)

mediation_results_propmediated<-rbind(mediation_results_propmediated,
                                      data.frame(Estimate = 0, cat = 5, results = "Prop. Mediated "))

p<-ggplot()+
    geom_area(data=mediation_results_propmediated,aes(x=cat,y=Estimate*100,fill=results))+
    xlab('Baseline Sodium Category')+
    ylab('%')+
   # scale_color_manual(labels = c("Proportion Mediated"), values = c("#f1c40f"))+
    scale_fill_manual(labels = c("Proportion Mediated"), values = c("#f1c40f"))+
    labs(colour='Mortality Risk', fill='Proportion Mediated') +
    theme(legend.position="bottom")+
    theme_minimal()+
  scale_x_continuous(breaks = c(1,2,3,4,5,6,7,8),labels = c("<120","120-125","125-130","130-135","135-140(reference)","140-145","145-150",">150")) +
  
  geom_point(data = total_effect,aes(x=cat,y=(or-1)*100,colour="red"))+
  geom_line(data = total_effect,aes(x=cat,y=(or-1)*100,colour="red"))+
  geom_errorbar(data=total_effect
                  ,aes(x=cat,ymin=(or.lb-1)*100,ymax=(or.ub-1)*100)
                  ,alpha=0.3
                  ,inherit.aes=FALSE
                  ,width=0.2)+
  scale_y_continuous(
    position ="right",
    
    # Features of the first axis
    name = "Proportion of mortality mediated by GCS score (%)",
    
    # Add a second axis and specify its features
    sec.axis = sec_axis(~(./100)+1, name="Odds ratio (95%CI) of hospital mortality")
  ) 

plot_grid(p, scale = c(1, 1),ncol = 1, align = "v")

```


```{r fig.height=9, fig.width=7}
mediation_results_total_effect<-mediation_results_all%>%filter(results %in% c("Total Effect" ) )

mediation_results_total_effect[,c(1,2,3)]<-exp(mediation_results_total_effect[,c(1,2,3)])


p1<-ggplot(data=mediation_results_total_effect,aes(x=cat,y=Estimate,colour=results,fill=results))+
    geom_point()+
    geom_line()+
    geom_ribbon(   data=mediation_results_total_effect
                  ,aes(x=cat,ymin=`95% CI Lower`,ymax=`95% CI Upper`,fill=results)
                  ,alpha=0.3
                  ,inherit.aes=F
                  ,width=0.2)+
    xlab('Baseline Sodium Category')+
    ylab('Odds Ratio')+
    scale_color_manual(labels = c("Total Effect"), values = c("#1abc9c"))+
    scale_fill_manual(labels = c("Total Effect"), values = c("#1abc9c"))+
    labs(colour='Metric', fill='Metric') +
    theme_minimal()+
  ggtitle('Total Effect')+
  scale_x_continuous(breaks = c(1,2,3,4,6,7,8)) 

mediation_results_propmediated<-mediation_results_all%>%filter(results %in% c("Prop. Mediated ") ) 

p2<-ggplot(data=mediation_results_propmediated,aes(x=cat,y=Estimate*100,fill=results))+
    geom_area()+
    xlab('Baseline Sodium Category')+
    ylab('%')+
    scale_color_manual(labels = c("Proportion Mediated"), values = c("#f1c40f"))+
    scale_fill_manual(labels = c("Proportion Mediated"), values = c("#f1c40f"))+
    labs(colour='Metric', fill='Metric') +
    theme_minimal()+
  ggtitle('Proportion Mediated')+
  scale_x_continuous(breaks = c(1,2,3,4,6,7,8)) 
p1+p2

plot_grid(p2, p1, scale = c(1, 1),ncol = 1, align = "v")
```

