---
title: "Sodium GCS Mediation Mortality"
author: Miguel Ángel Armengol de la Hoz
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output:
  html_notebook:
    code_folding: hide
    number_sections: yes
    theme: flatly
    toc: yes
    toc_float: yes

knit: (function(inputFile, encoding) { rmarkdown::render(inputFile, encoding = encoding, output_file = paste0(substr(inputFile,1,nchar(inputFile)-4)," ",Sys.Date(),'.html')) })
---

# Environment

```{r message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)
library(mediation)
library(mgcv)
library(plotly)
library(voxel)
library(htmlwidgets)
library(readr)

'%!in%' <- function(x,y)!('%in%'(x,y))

```

# Binary Baseline GCS analysis

## Mediation analysis: Baseline sodium_cat_1

### The total effect

```{r}
fit_totaleffect<-glm(as.factor(hosp_mortality)~ 
sodium_cat +                       
apache_iv+age+gender+ethnicity+unittype+ charlson_liver1 +charlson_chf1+mechvent_day01+apachedxgroup
# apache related variables
                   , sodium_cat_1_df
,family = 'binomial')
summary(fit_totaleffect)

OR_table<-as.data.frame(round(exp(cbind(OR=coef(fit_totaleffect), confint.default(fit_totaleffect))),2))

OR_table

```

### The effect of the baseline sodium onto the mediator

```{r}
fit_mediator<-lm( normal_gcs ~ sodium_cat +                       
apache_iv+age+gender+ethnicity+unittype+ charlson_liver1 +charlson_chf1+mechvent_day01+apachedxgroup
# apache related variables
 
                   ,data =  sodium_cat_1_df)


summary(fit_mediator)

```

### The effect of the mediator on the the dependent variable

```{r}
fit_dv<-glm(as.factor(hosp_mortality) ~ normal_gcs + sodium_cat +apache_iv+age+gender+ethnicity+unittype+ charlson_liver1 +charlson_chf1+mechvent_day01+apachedxgroup
                   + mechvent_day01 
                   , sodium_cat_1_df
           ,family = 'binomial')
summary(fit_dv)

OR_table<-as.data.frame(round(exp(cbind(OR=coef(fit_dv), confint.default(fit_dv))),2))

OR_table
```

### Causal Mediation Analysis

```{r}
#treat is the independend variable
results <- mediate(fit_mediator, fit_dv, treat="sodium_cat", mediator="normal_gcs",sims = 20)

summary(results)

```

## Mediation analysis: Baseline sodium_cat_2

### The total effect

```{r}
fit_totaleffect<-glm(as.factor(hosp_mortality)~ 
sodium_cat +                       
apache_iv+age+gender+ethnicity+unittype+ charlson_liver1 +charlson_chf1+mechvent_day01+apachedxgroup
# apache related variables
                   , sodium_cat_2_df
,family = 'binomial')
summary(fit_totaleffect)

OR_table<-as.data.frame(round(exp(cbind(OR=coef(fit_totaleffect), confint.default(fit_totaleffect))),2))

OR_table

```

### The effect of the baseline sodium onto the mediator

```{r}
fit_mediator<-lm( normal_gcs ~ sodium_cat +                       
apache_iv+age+gender+ethnicity+unittype+ charlson_liver1 +charlson_chf1+mechvent_day01+apachedxgroup
# apache related variables
 
                   ,data =  sodium_cat_2_df)


summary(fit_mediator)

```

### The effect of the mediator on the the dependent variable

```{r}
fit_dv<-glm(as.factor(hosp_mortality) ~ normal_gcs + sodium_cat +apache_iv+age+gender+ethnicity+unittype+ charlson_liver1 +charlson_chf1+mechvent_day01+apachedxgroup
                   + mechvent_day01 
                   , sodium_cat_2_df
           ,family = 'binomial')
summary(fit_dv)

OR_table<-as.data.frame(round(exp(cbind(OR=coef(fit_dv), confint.default(fit_dv))),2))

OR_table
```

### Causal Mediation Analysis

```{r}
#treat is the independend variable
results <- mediate(fit_mediator, fit_dv, treat="sodium_cat", mediator="normal_gcs",sims = 20)

summary(results)

```

## Mediation analysis: Baseline sodium_cat_3

### The total effect

```{r}
fit_totaleffect<-glm(as.factor(hosp_mortality)~ 
sodium_cat +                       
apache_iv+age+gender+ethnicity+unittype+ charlson_liver1 +charlson_chf1+mechvent_day01+apachedxgroup
# apache related variables
                   , sodium_cat_3_df
,family = 'binomial')
summary(fit_totaleffect)

OR_table<-as.data.frame(round(exp(cbind(OR=coef(fit_totaleffect), confint.default(fit_totaleffect))),2))

OR_table

```

### The effect of the baseline sodium onto the mediator

```{r}
fit_mediator<-lm( normal_gcs ~ sodium_cat +                       
apache_iv+age+gender+ethnicity+unittype+ charlson_liver1 +charlson_chf1+mechvent_day01+apachedxgroup
# apache related variables
 
                   ,data =  sodium_cat_3_df)


summary(fit_mediator)

```

### The effect of the mediator on the the dependent variable

```{r}
fit_dv<-glm(as.factor(hosp_mortality) ~ normal_gcs + sodium_cat +apache_iv+age+gender+ethnicity+unittype+ charlson_liver1 +charlson_chf1+mechvent_day01+apachedxgroup
                   + mechvent_day01 
                   , sodium_cat_3_df
           ,family = 'binomial')
summary(fit_dv)

OR_table<-as.data.frame(round(exp(cbind(OR=coef(fit_dv), confint.default(fit_dv))),2))

OR_table
```

### Causal Mediation Analysis

```{r}
#treat is the independend variable
results <- mediate(fit_mediator, fit_dv, treat="sodium_cat", mediator="normal_gcs",sims = 20)

summary(results)

```

## Mediation analysis: Baseline sodium_cat_4

### The total effect

```{r}
fit_totaleffect<-glm(as.factor(hosp_mortality)~ 
sodium_cat +                       
apache_iv+age+gender+ethnicity+unittype+ charlson_liver1 +charlson_chf1+mechvent_day01+apachedxgroup
# apache related variables
                   , sodium_cat_4_df
,family = 'binomial')
summary(fit_totaleffect)

OR_table<-as.data.frame(round(exp(cbind(OR=coef(fit_totaleffect), confint.default(fit_totaleffect))),2))

OR_table

```

### The effect of the baseline sodium onto the mediator

```{r}
fit_mediator<-lm( normal_gcs ~ sodium_cat +                       
apache_iv+age+gender+ethnicity+unittype+ charlson_liver1 +charlson_chf1+mechvent_day01+apachedxgroup
# apache related variables
 
                   ,data =  sodium_cat_4_df)


summary(fit_mediator)

```

### The effect of the mediator on the the dependent variable

```{r}
fit_dv<-glm(as.factor(hosp_mortality) ~ normal_gcs + sodium_cat +apache_iv+age+gender+ethnicity+unittype+ charlson_liver1 +charlson_chf1+mechvent_day01+apachedxgroup
                   + mechvent_day01 
                   , sodium_cat_4_df
           ,family = 'binomial')
summary(fit_dv)

OR_table<-as.data.frame(round(exp(cbind(OR=coef(fit_dv), confint.default(fit_dv))),2))

OR_table
```

### Causal Mediation Analysis

```{r}
#treat is the independend variable
results <- mediate(fit_mediator, fit_dv, treat="sodium_cat", mediator="normal_gcs",sims = 20)

summary(results)

```

## Mediation analysis: Baseline sodium_cat_6

### The total effect

```{r}
fit_totaleffect<-glm(as.factor(hosp_mortality)~ 
sodium_cat +                       
apache_iv+age+gender+ethnicity+unittype+ charlson_liver1 +charlson_chf1+mechvent_day01+apachedxgroup
# apache related variables
                   , sodium_cat_6_df
,family = 'binomial')
summary(fit_totaleffect)

OR_table<-as.data.frame(round(exp(cbind(OR=coef(fit_totaleffect), confint.default(fit_totaleffect))),2))

OR_table

```

### The effect of the baseline sodium onto the mediator

```{r}
fit_mediator<-lm( normal_gcs ~ sodium_cat +                       
apache_iv+age+gender+ethnicity+unittype+ charlson_liver1 +charlson_chf1+mechvent_day01+apachedxgroup
# apache related variables
 
                   ,data =  sodium_cat_6_df)


summary(fit_mediator)

```

### The effect of the mediator on the the dependent variable

```{r}
fit_dv<-glm(as.factor(hosp_mortality) ~ normal_gcs + sodium_cat +apache_iv+age+gender+ethnicity+unittype+ charlson_liver1 +charlson_chf1+mechvent_day01+apachedxgroup
                   + mechvent_day01 
                   , sodium_cat_6_df
           ,family = 'binomial')
summary(fit_dv)

OR_table<-as.data.frame(round(exp(cbind(OR=coef(fit_dv), confint.default(fit_dv))),2))

OR_table
```

### Causal Mediation Analysis

```{r}
#treat is the independend variable
results <- mediate(fit_mediator, fit_dv, treat="sodium_cat", mediator="normal_gcs",sims = 20)

summary(results)

```

## Mediation analysis: Baseline sodium_cat_7

### The total effect

```{r}
fit_totaleffect<-glm(as.factor(hosp_mortality)~ 
sodium_cat +                       
apache_iv+age+gender+ethnicity+unittype+ charlson_liver1 +charlson_chf1+mechvent_day01+apachedxgroup
# apache related variables
                   , sodium_cat_7_df
,family = 'binomial')
summary(fit_totaleffect)

OR_table<-as.data.frame(round(exp(cbind(OR=coef(fit_totaleffect), confint.default(fit_totaleffect))),2))

OR_table

```

### The effect of the baseline sodium onto the mediator

```{r}
fit_mediator<-lm( normal_gcs ~ sodium_cat +                       
apache_iv+age+gender+ethnicity+unittype+ charlson_liver1 +charlson_chf1+mechvent_day01+apachedxgroup
# apache related variables
 
                   ,data =  sodium_cat_7_df)


summary(fit_mediator)

```

### The effect of the mediator on the the dependent variable

```{r}
fit_dv<-glm(as.factor(hosp_mortality) ~ normal_gcs + sodium_cat +apache_iv+age+gender+ethnicity+unittype+ charlson_liver1 +charlson_chf1+mechvent_day01+apachedxgroup
                   + mechvent_day01 
                   , sodium_cat_7_df
           ,family = 'binomial')
summary(fit_dv)

OR_table<-as.data.frame(round(exp(cbind(OR=coef(fit_dv), confint.default(fit_dv))),2))

OR_table
```

### Causal Mediation Analysis

```{r}
#treat is the independend variable
results <- mediate(fit_mediator, fit_dv, treat="sodium_cat", mediator="normal_gcs",sims = 20)

summary(results)

```

## Mediation analysis: Baseline sodium_cat_8

### The total effect

```{r}
fit_totaleffect<-glm(as.factor(hosp_mortality)~ 
sodium_cat +                       
apache_iv+age+gender+ethnicity+unittype+ charlson_liver1 +charlson_chf1+mechvent_day01+apachedxgroup
# apache related variables
                   , sodium_cat_8_df
,family = 'binomial')
summary(fit_totaleffect)

OR_table<-as.data.frame(round(exp(cbind(OR=coef(fit_totaleffect), confint.default(fit_totaleffect))),2))

OR_table

```

### The effect of the baseline sodium onto the mediator

```{r}
fit_mediator<-lm( normal_gcs ~ sodium_cat +                       
apache_iv+age+gender+ethnicity+unittype+ charlson_liver1 +charlson_chf1+mechvent_day01+apachedxgroup
# apache related variables
 
                   ,data =  sodium_cat_8_df)


summary(fit_mediator)

```

### The effect of the mediator on the the dependent variable

```{r}
fit_dv<-glm(as.factor(hosp_mortality) ~ normal_gcs + sodium_cat +apache_iv+age+gender+ethnicity+unittype+ charlson_liver1 +charlson_chf1+mechvent_day01+apachedxgroup
                   + mechvent_day01 
                   , sodium_cat_8_df
           ,family = 'binomial')
summary(fit_dv)

OR_table<-as.data.frame(round(exp(cbind(OR=coef(fit_dv), confint.default(fit_dv))),2))

OR_table
```

### Causal Mediation Analysis

```{r}
#treat is the independend variable
results <- mediate(fit_mediator, fit_dv, treat="sodium_cat", mediator="normal_gcs",sims = 20)

summary(results)

```





