---
title: "Sodium GCS Mediation Mortality"
author: Miguel Ángel Armengol de la Hoz
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output:
  html_notebook:
    code_folding: hide
    number_sections: yes
    theme: flatly
    toc: yes
    toc_float: yes

knit: (function(inputFile, encoding) { rmarkdown::render(inputFile, encoding = encoding, output_file = paste0(substr(inputFile,1,nchar(inputFile)-4)," ",Sys.Date(),'.html')) })
---

# Environment

```{r message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)
library(mediation)
library(mgcv)
library(plotly)
library(voxel)
library(htmlwidgets)
library(readr)

'%!in%' <- function(x,y)!('%in%'(x,y))

```

# Continuous Baseline GCS Analysis

## The total effect

```{r}
fit_totaleffect<-gam(as.factor(hospmortality)~  s(sodium1) + apachescore+patientage+patientgender+ethnicity+unittype + apachedxgroup + charlson_liver1 +charlson_chf1 
                   + mechvent_day01 # must mech vent be included ?
                   , john_dataset,family = 'binomial')


summary(fit_totaleffect)

OR_table<-as.data.frame(round(exp(cbind(OR=coef(fit_totaleffect), confint.default(fit_totaleffect))),2))

OR_table

# the most optimal model introduces a spline term on sodium1 stratified by patiengage but the difference was so small we are not stratifying by age

```

### GAM Plot

```{r}
plotgam_model_GFR_Cys_0C<-plotGAM(gamFit = fit_totaleffect, smooth.cov = c("sodium1"),rawOrFitted = F, orderedAsFactor = FALSE) 

plotgam_model_GFR_Cys_0C_exp<-as.data.frame(cbind(
   plotgam_model_GFR_Cys_0C[["data"]][["sodium1"]]
  ,plotgam_model_GFR_Cys_0C[["data"]][["fit"]]
  ,plotgam_model_GFR_Cys_0C[["data"]][["se.fit"]])
)

colnames(plotgam_model_GFR_Cys_0C_exp)[1]<-'sodium1'
colnames(plotgam_model_GFR_Cys_0C_exp)[2]<-'fit'
colnames(plotgam_model_GFR_Cys_0C_exp)[3]<-'se.fit'

plotgam_model_GFR_Cys_0C_exp$lci<-(plotgam_model_GFR_Cys_0C_exp$fit) - 2 * (plotgam_model_GFR_Cys_0C_exp$se.fit)
plotgam_model_GFR_Cys_0C_exp$uci<-(plotgam_model_GFR_Cys_0C_exp$fit) + 2 * (plotgam_model_GFR_Cys_0C_exp$se.fit)

p<-ggplot(plotgam_model_GFR_Cys_0C_exp, aes(sodium1, fit))+
    geom_line(colour="#16a085")+
    geom_ribbon( john_dataset=plotgam_model_GFR_Cys_0C_exp,aes(ymin=lci,ymax=uci),alpha=0.3)+
    xlab('Baseline Sodium')+
    ylab('Mortality')+
    theme_minimal()+ labs(title="NDE: GAM with spline in sodium 1 (adjusted)")   

#p<-ggplotly(p)%>% layout(showlegend = FALSE)

p
```

## The effect of the baseline sodium onto the mediator

```{r}
fit_mediator<-lm(gcs_baseline ~ sodium1 + apachescore+patientage+patientgender+ethnicity+unittype + apachedxgroup + charlson_liver1 +charlson_chf1 + mechvent_day01
                   ,data= john_dataset)


summary(fit_mediator)

```

### GAM Plot

```{r}
fit_gam<-gam(gcs_baseline ~ s(sodium1) + apachescore+patientage+patientgender+ethnicity+unittype + apachedxgroup + charlson_liver1 +charlson_chf1 + mechvent_day01
                   ,data= john_dataset)


plotgam_model_GFR_Cys_0C<-plotGAM(gamFit = fit_gam, smooth.cov = c("sodium1") 
        ,rawOrFitted = F, orderedAsFactor = FALSE) 

plotgam_model_GFR_Cys_0C_exp<-as.data.frame(cbind(
   plotgam_model_GFR_Cys_0C[["data"]][["sodium1"]]
  ,plotgam_model_GFR_Cys_0C[["data"]][["fit"]]
  ,plotgam_model_GFR_Cys_0C[["data"]][["se.fit"]])
)

colnames(plotgam_model_GFR_Cys_0C_exp)[1]<-'sodium1'
colnames(plotgam_model_GFR_Cys_0C_exp)[2]<-'fit'
colnames(plotgam_model_GFR_Cys_0C_exp)[3]<-'se.fit'

plotgam_model_GFR_Cys_0C_exp$lci<-(plotgam_model_GFR_Cys_0C_exp$fit) - 2 * (plotgam_model_GFR_Cys_0C_exp$se.fit)
plotgam_model_GFR_Cys_0C_exp$uci<-(plotgam_model_GFR_Cys_0C_exp$fit) + 2 * (plotgam_model_GFR_Cys_0C_exp$se.fit)

p<-ggplot(plotgam_model_GFR_Cys_0C_exp, aes(sodium1, fit,colour="#16a085"))+
    geom_line()+
    geom_ribbon( john_dataset=plotgam_model_GFR_Cys_0C_exp,aes(ymin=lci,ymax=uci),alpha=0.3)+
    xlab('Baseline Sodium')+
    ylab('Baseline GCS')+
    theme_minimal()+ labs(title="Generalized Additive Model with spline in sodium 1 (adjusted)")   

p<-ggplotly(p)%>% layout(showlegend = FALSE)

p
```


## The effect of the mediator on the the dependent variable

```{r}
fit_dv<-gam(as.factor(hospmortality) ~ s(gcs_baseline) + s(sodium1) + apachescore+patientage+patientgender+ethnicity+unittype + apachedxgroup + charlson_liver1 +charlson_chf1 
                   + mechvent_day01 
                   , john_dataset
           ,family = 'binomial')
summary(fit_dv)

OR_table<-as.data.frame(round(exp(cbind(OR=coef(fit_dv), confint.default(fit_dv))),2))

OR_table
```

### GAM Plot

```{r}
plotgam_model_GFR_Cys_0C<-plotGAM(gamFit = fit_dv, smooth.cov = c("gcs_baseline"),rawOrFitted = F, orderedAsFactor = FALSE) 

plotgam_model_GFR_Cys_0C_exp<-as.data.frame(cbind(
   plotgam_model_GFR_Cys_0C[["data"]][["gcs_baseline"]]
  ,plotgam_model_GFR_Cys_0C[["data"]][["fit"]]
  ,plotgam_model_GFR_Cys_0C[["data"]][["se.fit"]])
)

colnames(plotgam_model_GFR_Cys_0C_exp)[1]<-'gcs_baseline'
colnames(plotgam_model_GFR_Cys_0C_exp)[2]<-'fit'
colnames(plotgam_model_GFR_Cys_0C_exp)[3]<-'se.fit'

plotgam_model_GFR_Cys_0C_exp$lci<-(plotgam_model_GFR_Cys_0C_exp$fit) - 2 * (plotgam_model_GFR_Cys_0C_exp$se.fit)
plotgam_model_GFR_Cys_0C_exp$uci<-(plotgam_model_GFR_Cys_0C_exp$fit) + 2 * (plotgam_model_GFR_Cys_0C_exp$se.fit)

p<-ggplot(plotgam_model_GFR_Cys_0C_exp, aes(gcs_baseline, fit))+
    geom_line(colour="#16a085")+
    geom_ribbon( john_dataset=plotgam_model_GFR_Cys_0C_exp,aes(ymin=lci,ymax=uci),alpha=0.3)+
    xlab('Baseline GCS')+
    ylab('Mortality')+
    theme_minimal()+ labs(title="Effect of the mediator on the the dependent variable")   

#p<-ggplotly(p)%>% layout(showlegend = FALSE)

p
```

## Causal Mediation Analysis

```{r}
#treat is the independend variable
results <- mediate(fit_mediator, fit_dv, treat="sodium1", mediator="gcs_baseline")

summary(results)

```

Figures

```{r}
ggplot( john_dataset, aes(x=`sodium categories`, y=gcs_baseline, group= john_dataset$`sodium categories`) ) + geom_boxplot()+
    scale_fill_brewer(palette="BuPu")+theme_minimal()
```

# Binary Baseline GCS analysis

## The total effect

```{r}
fit_totaleffect<-glm(as.factor(hospmortality)~  sodium1 + apachescore+patientage+patientgender+ethnicity+unittype + apachedxgroup + charlson_liver1 +charlson_chf1 
                   + mechvent_day01 # must mech vent be included ?
                   , john_dataset,family = 'binomial')
summary(fit_totaleffect)

OR_table<-as.data.frame(round(exp(cbind(OR=coef(fit_totaleffect), confint.default(fit_totaleffect))),2))

OR_table

```

## The effect of the baseline sodium onto the mediator

```{r}
fit_mediator<-lm( normal_gcs ~ sodium1 + apachescore+patientage+patientgender+ethnicity+unittype + apachedxgroup + charlson_liver1 +charlson_chf1 + mechvent_day01
                   ,data =  john_dataset)


summary(fit_mediator)

```

### GAM Plot

```{r}
fit_gam<-gam(normal_gcs ~ s(sodium1) + apachescore+patientage+patientgender+ethnicity+unittype + apachedxgroup + charlson_liver1 +charlson_chf1 + mechvent_day01
                   ,data= john_dataset)


plotgam_model_GFR_Cys_0C<-plotGAM(gamFit = fit_gam, smooth.cov = c("sodium1") 
        ,rawOrFitted = F, orderedAsFactor = FALSE) 

plotgam_model_GFR_Cys_0C_exp<-as.data.frame(cbind(
   plotgam_model_GFR_Cys_0C[["data"]][["sodium1"]]
  ,plotgam_model_GFR_Cys_0C[["data"]][["fit"]]
  ,plotgam_model_GFR_Cys_0C[["data"]][["se.fit"]])
)

colnames(plotgam_model_GFR_Cys_0C_exp)[1]<-'sodium1'
colnames(plotgam_model_GFR_Cys_0C_exp)[2]<-'fit'
colnames(plotgam_model_GFR_Cys_0C_exp)[3]<-'se.fit'

plotgam_model_GFR_Cys_0C_exp$lci<-(plotgam_model_GFR_Cys_0C_exp$fit) - 2 * (plotgam_model_GFR_Cys_0C_exp$se.fit)
plotgam_model_GFR_Cys_0C_exp$uci<-(plotgam_model_GFR_Cys_0C_exp$fit) + 2 * (plotgam_model_GFR_Cys_0C_exp$se.fit)

p<-ggplot(plotgam_model_GFR_Cys_0C_exp, aes(sodium1, fit,colour="#16a085"))+
    geom_line()+
    geom_ribbon(data=plotgam_model_GFR_Cys_0C_exp,aes(ymin=lci,ymax=uci),alpha=0.3)+
    xlab('Baseline Sodium')+
    ylab('Baseline GCS')+
    theme_minimal()+ labs(title="Generalized Additive Model with spline in sodium 1 (adjusted)")   

p<-ggplotly(p)%>% layout(showlegend = FALSE)

p

```


## The effect of the mediator on the the dependent variable

```{r}
fit_dv<-glm(as.factor(hospmortality) ~ normal_gcs + sodium1 + apachescore+patientage+patientgender+ethnicity+unittype + apachedxgroup + charlson_liver1 +charlson_chf1 
                   + mechvent_day01 
                   , john_dataset
           ,family = 'binomial')
summary(fit_dv)

OR_table<-as.data.frame(round(exp(cbind(OR=coef(fit_dv), confint.default(fit_dv))),2))

OR_table
```

## Causal Mediation Analysis

```{r}
#treat is the independend variable
results <- mediate(fit_mediator, fit_dv, treat="sodium1", mediator="normal_gcs")

summary(results)

```

# Binary Baseline GCS and separate Sodium bands analisys

## Creating the subsets

```{r}
sodium_split <- split( john_dataset,  john_dataset$sodium_cats)
```

## Storing every subset in a different dataset

```{r}
for (i in c(1,2,3,4,6,7)){ 
assign( paste('sodium_cat_',i+1,'_ john_dataset',sep=''), sodium_split[[i]])}

```



## The total effect

```{r}
fit_totaleffect<-glm(as.factor(hospmortality)~  sodium1 + apachescore+patientage+patientgender+ethnicity+unittype + apachedxgroup + charlson_liver1 +charlson_chf1 
                   + mechvent_day01 # must mech vent be included ?
                   , john_dataset,family = 'binomial')
summary(fit_totaleffect)

OR_table<-as.data.frame(round(exp(cbind(OR=coef(fit_totaleffect), confint.default(fit_totaleffect))),2))

OR_table

```

## The effect of the baseline sodium onto the mediator

```{r}
fit_mediator<-lm( normal_gcs ~ sodium1 + apachescore+patientage+patientgender+ethnicity+unittype + apachedxgroup + charlson_liver1 +charlson_chf1 + mechvent_day01
                   ,data =  john_dataset)


summary(fit_mediator)

```

### GAM Plot

```{r}
fit_gam<-gam(normal_gcs ~ s(sodium1) + apachescore+patientage+patientgender+ethnicity+unittype + apachedxgroup + charlson_liver1 +charlson_chf1 + mechvent_day01
                   ,data= john_dataset)


plotgam_model_GFR_Cys_0C<-plotGAM(gamFit = fit_gam, smooth.cov = c("sodium1") 
        ,rawOrFitted = F, orderedAsFactor = FALSE) 

plotgam_model_GFR_Cys_0C_exp<-as.data.frame(cbind(
   plotgam_model_GFR_Cys_0C[["data"]][["sodium1"]]
  ,plotgam_model_GFR_Cys_0C[["data"]][["fit"]]
  ,plotgam_model_GFR_Cys_0C[["data"]][["se.fit"]])
)

colnames(plotgam_model_GFR_Cys_0C_exp)[1]<-'sodium1'
colnames(plotgam_model_GFR_Cys_0C_exp)[2]<-'fit'
colnames(plotgam_model_GFR_Cys_0C_exp)[3]<-'se.fit'

plotgam_model_GFR_Cys_0C_exp$lci<-(plotgam_model_GFR_Cys_0C_exp$fit) - 2 * (plotgam_model_GFR_Cys_0C_exp$se.fit)
plotgam_model_GFR_Cys_0C_exp$uci<-(plotgam_model_GFR_Cys_0C_exp$fit) + 2 * (plotgam_model_GFR_Cys_0C_exp$se.fit)

p<-ggplot(plotgam_model_GFR_Cys_0C_exp, aes(sodium1, fit,colour="#16a085"))+
    geom_line()+
    geom_ribbon(data=plotgam_model_GFR_Cys_0C_exp,aes(ymin=lci,ymax=uci),alpha=0.3)+
    xlab('Baseline Sodium')+
    ylab('Baseline GCS')+
    theme_minimal()+ labs(title="Generalized Additive Model with spline in sodium 1 (adjusted)")   

p<-ggplotly(p)%>% layout(showlegend = FALSE)

p

```


## The effect of the mediator on the the dependent variable

```{r}
fit_dv<-glm(as.factor(hospmortality) ~ normal_gcs + sodium1 + apachescore+patientage+patientgender+ethnicity+unittype + apachedxgroup + charlson_liver1 +charlson_chf1 
                   + mechvent_day01 
                   , john_dataset
           ,family = 'binomial')
summary(fit_dv)

OR_table<-as.data.frame(round(exp(cbind(OR=coef(fit_dv), confint.default(fit_dv))),2))

OR_table
```

## Causal Mediation Analysis

```{r}
#treat is the independend variable
results <- mediate(fit_mediator, fit_dv, treat="sodium1", mediator="normal_gcs")

summary(results)
```


