---
title: "03_tables_creation"
# output: html_notebook
---

# Environment

```{r}
library(tableone)
library(magrittr)
library(dplyr)
library(summarytools)
library(kableExtra)
library(openxlsx)
library(stringi)
library(officer)
library(DT)

# Function to save results to a Word document
save_to_word <- function(data, file_name, title) {
  doc <- read_docx()
  doc <- doc %>%
    body_add_par(title, style = "heading 1") %>%
    body_add_table(value = data, style = "table_template")
  
  print(doc, target = file_name)
}
```


## Sup Table 2: Not Sedated

All variables.  Stratification (ie column groupings) should be by sodium categories.

```{r,echo = F}
dataframe_for_table1<-df_sub_notsedated


# we are re-difining the 'Others' category
dataframe_for_table1$unittype<-plyr::revalue(
  dataframe_for_table1$unittype,c( 'Neuro ICU'='Others' , 'SICU'= 'Others', 'Cardiac ICU'='Others' , 'CCU-CTICU'='Others'  ,'CSICU'='Others' , 'CTICU'='Others')
)

vars_in_table1<-c("age", "gender", "AP_grp", "ethnicity", "apachescore", "mechvent_day01", "unittype", "charlson_chf1", "charlson_liver1", "creatinine", "creatinine_great2", "gcs_baseline", "normal_gcs", "hosp_mortality","sodium_cat","adm_source","sedation_status")
# lists the variables we are going to be including in table1 (excepts for the stratification variable)

dataframe_for_table1<-dataframe_for_table1[,vars_in_table1]
cat_variables<-rep(NA, length(vars_in_table1))
# creates a dataset only with select variable using our original dataset and generated an empty list that is going to be used for storing categorical variables 

stratifyby<-"sodium_cat"
label(dataframe_for_table1$age) <- 'Age'
label(dataframe_for_table1$gender) <- 'Gender'
label(dataframe_for_table1$AP_grp) <- 'APACHE Diagnosis Group'
label(dataframe_for_table1$ethnicity) <- 'Ethnicity'
label(dataframe_for_table1$apachescore) <- 'APACHE Score'
label(dataframe_for_table1$mechvent_day01) <- 'Mechanical Ventilation on Day 1'
label(dataframe_for_table1$unittype) <- 'ICU Type'
label(dataframe_for_table1$charlson_chf1) <- 'Charlson Comorbidity - Congestive Heart Failure'
label(dataframe_for_table1$charlson_liver1) <- 'Charlson Comorbidity - Liver Disease'
label(dataframe_for_table1$creatinine) <- 'Creatinine'
label(dataframe_for_table1$creatinine_great2) <- 'Creatinine > 2'
label(dataframe_for_table1$gcs_baseline) <- 'Baseline Glasgow Coma Scale'
label(dataframe_for_table1$normal_gcs) <- 'Normal Glasgow Coma Scale'
label(dataframe_for_table1$hosp_mortality) <- 'Hospital Mortality'
label(dataframe_for_table1$sodium_cat) <- 'Sodium Category'
label(dataframe_for_table1$adm_source) <- 'Admission Source'
label(dataframe_for_table1$sedation_status) <- 'Sedation Status'
# selects variable to stratify by and label is useful for prettyfying variable names we want to appear on table 1

# detects whether a variable is categorical or not based on the number of distinct variables per category (10 is the threshold)
cont<-1
for (i in 1:length(vars_in_table1) ) {
  if ( n_distinct(dataframe_for_table1[vars_in_table1[i] ])<=10 ) {
    print(i)
    print(vars_in_table1[i])
    print(names(dataframe_for_table1[vars_in_table1[i]]))
    cat_variables[cont]<-names(dataframe_for_table1[vars_in_table1[i]])
    cont<-cont+1
  }
}  
cat_variables<-cat_variables[!is.na(cat_variables)]
table1_base<-print(CreateTableOne(vars = vars_in_table1
                                   ,strata = stratifyby
                                  , factorVars = cat_variables
    ,data = dataframe_for_table1, addOverall=T),varLabels = T ,nonnormal  = c('creatinine','gcs_baseline')
)

write.csv(table1_base, file = "sup_tables/table1_not_sedated.csv")

# run this in console for html output, the code below uses library kableExtra::
# the line add_header_above() will need to be modified depending on the number of columns of the table
starification_cats<-n_distinct(dataframe_for_table1[,stratifyby])
table1_base %>%
  kbl(caption = "Table 1 Not Sedated" , align = "c") %>%
  kable_classic_2(full_width = F, html_font = "Cambria")%>%
  add_header_above(c(" "," ", 'Sodium Band Category' = starification_cats," ", "" ))

```
## Sup Table 3: Sedated

All variables.  Stratification (ie column groupings) should be by sodium categories.

```{r,echo = F}
dataframe_for_table1<-df_sub_sedated


# we are re-difining the 'Others' category
dataframe_for_table1$unittype<-plyr::revalue(
  dataframe_for_table1$unittype,c( 'Neuro ICU'='Others' , 'SICU'= 'Others', 'Cardiac ICU'='Others' , 'CCU-CTICU'='Others'  ,'CSICU'='Others' , 'CTICU'='Others')
)

vars_in_table1<-c("age", "gender", "AP_grp", "ethnicity", "apachescore", "mechvent_day01", "unittype", "charlson_chf1", "charlson_liver1", "creatinine", "creatinine_great2", "gcs_baseline", "normal_gcs", "hosp_mortality","sodium_cat","adm_source","sedation_status")
# lists the variables we are going to be including in table1 (excepts for the stratification variable)

dataframe_for_table1<-dataframe_for_table1[,vars_in_table1]
cat_variables<-rep(NA, length(vars_in_table1))
# creates a dataset only with select variable using our original dataset and generated an empty list that is going to be used for storing categorical variables 

stratifyby<-"sodium_cat"
label(dataframe_for_table1$age) <- 'Age'
label(dataframe_for_table1$gender) <- 'Gender'
label(dataframe_for_table1$AP_grp) <- 'APACHE Diagnosis Group'
label(dataframe_for_table1$ethnicity) <- 'Ethnicity'
label(dataframe_for_table1$apachescore) <- 'APACHE Score'
label(dataframe_for_table1$mechvent_day01) <- 'Mechanical Ventilation on Day 1'
label(dataframe_for_table1$unittype) <- 'ICU Type'
label(dataframe_for_table1$charlson_chf1) <- 'Charlson Comorbidity - Congestive Heart Failure'
label(dataframe_for_table1$charlson_liver1) <- 'Charlson Comorbidity - Liver Disease'
label(dataframe_for_table1$creatinine) <- 'Creatinine'
label(dataframe_for_table1$creatinine_great2) <- 'Creatinine > 2'
label(dataframe_for_table1$gcs_baseline) <- 'Baseline Glasgow Coma Scale'
label(dataframe_for_table1$normal_gcs) <- 'Normal Glasgow Coma Scale'
label(dataframe_for_table1$hosp_mortality) <- 'Hospital Mortality'
label(dataframe_for_table1$sodium_cat) <- 'Sodium Category'
label(dataframe_for_table1$adm_source) <- 'Admission Source'
label(dataframe_for_table1$sedation_status) <- 'Sedation Status'
# selects variable to stratify by and label is useful for prettyfying variable names we want to appear on table 1

# detects whether a variable is categorical or not based on the number of distinct variables per category (10 is the threshold)
cont<-1
for (i in 1:length(vars_in_table1) ) {
  if ( n_distinct(dataframe_for_table1[vars_in_table1[i] ])<=10 ) {
    print(i)
    print(vars_in_table1[i])
    print(names(dataframe_for_table1[vars_in_table1[i]]))
    cat_variables[cont]<-names(dataframe_for_table1[vars_in_table1[i]])
    cont<-cont+1
  }
}  
cat_variables <- cat_variables[!is.na(cat_variables)]
table1_base <- print(CreateTableOne(vars = vars_in_table1,
                                    strata = stratifyby,
                                    factorVars = cat_variables,
                                    data = dataframe_for_table1, addOverall = TRUE), varLabels = TRUE, nonnormal = c('creatinine', 'gcs_baseline')
)

## Save to a CSV file
write.csv(table1_base, file = "sup_tables/table1_sedated.csv")

# Run this in console for HTML output, the code below uses library kableExtra::
# The line add_header_above() will need to be modified depending on the number of columns of the table
starification_cats <- n_distinct(dataframe_for_table1[, stratifyby])
table1_base %>%
  kbl(caption = "Table 1 Sedated", align = "c") %>%
  kable_classic_2(full_width = FALSE, html_font = "Cambria") %>%
  add_header_above(c(" ", " ", 'Sodium Band Category' = starification_cats, " ", ""))



```

## Exploratory Table Suspected error

All variables.  Stratification (ie column groupings) should be by sodium categories.

```{r,echo = F}
dataframe_for_table1<-df_sub_susp_error


# we are re-difining the 'Others' category
dataframe_for_table1$unittype<-plyr::revalue(
  dataframe_for_table1$unittype,c( 'Neuro ICU'='Others' , 'SICU'= 'Others', 'Cardiac ICU'='Others' , 'CCU-CTICU'='Others'  ,'CSICU'='Others' , 'CTICU'='Others')
)

vars_in_table1<-c("age", "gender", "AP_grp", "ethnicity", "apachescore", "mechvent_day01", "unittype", "charlson_chf1", "charlson_liver1", "creatinine", "creatinine_great2", "gcs_baseline", "normal_gcs", "hosp_mortality","sodium_cat","adm_source","sedation_status")
# lists the variables we are going to be including in table1 (excepts for the stratification variable)

dataframe_for_table1<-dataframe_for_table1[,vars_in_table1]
cat_variables<-rep(NA, length(vars_in_table1))
# creates a dataset only with select variable using our original dataset and generated an empty list that is going to be used for storing categorical variables 

stratifyby<-"sodium_cat"
label(dataframe_for_table1$age) <- 'Age'
label(dataframe_for_table1$gender) <- 'Gender'
label(dataframe_for_table1$AP_grp) <- 'APACHE Diagnosis Group'
label(dataframe_for_table1$ethnicity) <- 'Ethnicity'
label(dataframe_for_table1$apachescore) <- 'APACHE Score'
label(dataframe_for_table1$mechvent_day01) <- 'Mechanical Ventilation on Day 1'
label(dataframe_for_table1$unittype) <- 'ICU Type'
label(dataframe_for_table1$charlson_chf1) <- 'Charlson Comorbidity - Congestive Heart Failure'
label(dataframe_for_table1$charlson_liver1) <- 'Charlson Comorbidity - Liver Disease'
label(dataframe_for_table1$creatinine) <- 'Creatinine'
label(dataframe_for_table1$creatinine_great2) <- 'Creatinine > 2'
label(dataframe_for_table1$gcs_baseline) <- 'Baseline Glasgow Coma Scale'
label(dataframe_for_table1$normal_gcs) <- 'Normal Glasgow Coma Scale'
label(dataframe_for_table1$hosp_mortality) <- 'Hospital Mortality'
label(dataframe_for_table1$sodium_cat) <- 'Sodium Category'
label(dataframe_for_table1$adm_source) <- 'Admission Source'
label(dataframe_for_table1$sedation_status) <- 'Sedation Status'
# selects variable to stratify by and label is useful for prettyfying variable names we want to appear on table 1

# detects whether a variable is categorical or not based on the number of distinct variables per category (10 is the threshold)
cont<-1
for (i in 1:length(vars_in_table1) ) {
  if ( n_distinct(dataframe_for_table1[vars_in_table1[i] ])<=10 ) {
    print(i)
    print(vars_in_table1[i])
    print(names(dataframe_for_table1[vars_in_table1[i]]))
    cat_variables[cont]<-names(dataframe_for_table1[vars_in_table1[i]])
    cont<-cont+1
  }
}  
cat_variables<-cat_variables[!is.na(cat_variables)]
table1_base<-print(CreateTableOne(vars = vars_in_table1
                                   ,strata = stratifyby
                                  , factorVars = cat_variables
    ,data = dataframe_for_table1, addOverall=T),varLabels = T ,nonnormal  = c('creatinine','gcs_baseline')
)


# run this in console for html output, the code below uses library kableExtra::
# the line add_header_above() will need to be modified depending on the number of columns of the table
starification_cats<-n_distinct(dataframe_for_table1[,stratifyby])
table1_base %>%
  kbl(caption = "Table 1 Suspected Error" , align = "c") %>%
  kable_classic_2(full_width = F, html_font = "Cambria")%>%
  add_header_above(c(" "," ", 'Sodium Band Category' = starification_cats," ", "" ))

# Function to save results to a Word document
save_to_word <- function(data, file_name, title) {
  doc <- read_docx()
  doc <- doc %>%
    body_add_par(title, style = "heading 1") %>%
    body_add_table(value = data, style = "table_template")
  
  print(doc, target = file_name)
}

# Save overall elective and not elective results
save_to_word(table1_base, "sup_tables/Mediation Analysis Elective Surgery.docx","Mediation Analysis Elective Surgery")

```

If suspected error looks a lot like “sedated” then it may be reasonable to group them together.
After reviewing table one we conclude we wouldn’t integrate. GCS suggests they’re not sedated.
Nate 16/4/24 via whatsapp

### Results export

```{r}
# Save overall elective and not elective results

save_to_word(results_overall_not_elective, "sup_tables/Mediation Analysis Not Elective Surgery.docx","Mediation Analysis Not Elective Surgery")

```



# Sup. Table 4A: Mediation tables for ICU Types

```{r}
library(dplyr)
library(tibble)

# Define the function to process mediation results
process_mediation_results_uci <- function(mediation_result, category) {
  mediation_result$cat <- category
  mediation_result$results <- c("Total Effect", "ACME (average)", "ADE (average)", "Prop. Mediated (average)")
  
  # Round p_value to 3 decimal places if p_value column exists
  if ("p_value" %in% colnames(mediation_result)) {
    mediation_result <- mediation_result %>% mutate(p_value = round(p_value, 3))
  }
  
  return(mediation_result)
}

# Function to calculate total effect OR for each category
calculate_total_effect_or <- function(df) {
  total_effect_rows <- df %>% filter(results == "Total Effect")
  if (!all(c("Estimate", "95% CI Lower", "95% CI Upper") %in% colnames(total_effect_rows))) {
    stop("Required columns are missing in the dataframe")
  }
  
  total_effect_or <- total_effect_rows %>%
    transmute(
      Estimate = exp(Estimate),
      `95% CI Lower` = exp(`95% CI Lower`),
      `95% CI Upper` = exp(`95% CI Upper`),
      cat,
      results = "Total Effect (OR)"
    )
  
  # Include p_value if it exists and is not all NA
  if ("p_value" %in% colnames(total_effect_rows) && !all(is.na(total_effect_rows$p_value))) {
    total_effect_or <- total_effect_or %>%
      mutate(p_value = mean(total_effect_rows$p_value, na.rm = TRUE))
  }
  
  return(total_effect_or)
}

# Initialize a list to store combined results for each ICU type
icu_types <- unique(sub("cat_\\d+", "", names(mediation_summary_results)))
results_list <- setNames(vector("list", length(icu_types)), icu_types)

# Iterate through the list and process each dataframe
for (name in names(mediation_summary_results)) {
  # Extract the ICU type and category from the name
  parts <- unlist(strsplit(name, "_cat_"))
  uci_type <- parts[1]
  category <- as.numeric(parts[2])
  
  # Process the mediation results
  processed_results <- process_mediation_results_uci(mediation_summary_results[[name]], category)
  
  # Combine the results by ICU type, initialize with an empty data frame if NULL
  if (is.null(results_list[[uci_type]])) {
    results_list[[uci_type]] <- processed_results
  } else {
    results_list[[uci_type]] <- bind_rows(results_list[[uci_type]], processed_results)
  }
}

# Calculate and add total effect OR for each ICU type
for (icu_type in names(results_list)) {
  if (!is.null(results_list[[icu_type]])) {
    total_effect_or <- calculate_total_effect_or(results_list[[icu_type]])
    
    # Append the Total Effect (OR) row
    results_list[[icu_type]] <- bind_rows(results_list[[icu_type]], total_effect_or)
    
    # Remove row names and empty p_value columns
    results_list[[icu_type]] <- results_list[[icu_type]] %>%
      rownames_to_column(var = "Row") %>%
      select(-Row) %>%
      select(-one_of("p_value"))
  }
}

# Assign the processed results to individual variables
icu_a_results <- results_list$icu_a
icu_b_results <- results_list$icu_b
icu_c_results <- results_list$icu_c
icu_d_results <- results_list$icu_d
icu_e_results <- results_list$icu_e
icu_f_results <- results_list$icu_f
icu_g_results <- results_list$icu_g
```

## Results export

```{r}
# List of ICU types and corresponding names
icu_types <- c("Cardiac ICU", "CCU-CTICU", "CSICU+CTICU", "Med-Surg ICU", "MICU", "Neuro ICU", "SICU")
icu_names <- c("icu_a", "icu_b", "icu_c", "icu_d", "icu_e", "icu_f", "icu_g")

# Create a list of results for each ICU type
icu_results_list <- list(
  icu_a = icu_a_results,
  icu_b = icu_b_results,
  icu_c = icu_c_results,
  icu_d = icu_d_results,
  icu_e = icu_e_results,
  icu_f = icu_f_results,
  icu_g = icu_g_results
)

# Function to save results to a Word document
save_to_word <- function(data, file_name) {
  doc <- read_docx()
  doc <- doc %>%
    body_add_par("Mediation Analysis ICU Type", style = "heading 1") %>%
    body_add_table(value = data, style = "table_template")
  
  print(doc, target = file_name)
}

# Iterate over each ICU type and save results to separate Word documents
for (i in seq_along(icu_names)) {
  icu_name <- icu_names[i]
  icu_type <- icu_types[i]
  
  if (!is.null(icu_results_list[[icu_name]])) {
    file_name <- paste0("sup_tables/Mediation Analysis ", icu_type, ".docx")
    save_to_word(icu_results_list[[icu_name]], file_name)
  }
}
```

# Sup. Table 4B: Mediation tables for Surgery Type

```{r}
# Load necessary libraries


# Define a function to process mediation results
process_mediation_results <- function(category, elective_status) {
  df_name <- paste0("mediation_results_cat", category, "_", elective_status, "_sum")
  df <- get(df_name)
  df$cat <- category
  df$results <- stri_replace_all_fixed(rownames(df), '(treated)', "")
  return(df)
}

# List of categories
categories <- c(1, 2, 3, 4, 6, 7, 8)

# Process elective data frames
mediation_results_elective <- do.call(rbind, lapply(categories, process_mediation_results, elective_status = "elective"))

# Process not elective data frames
mediation_results_not_elective <- do.call(rbind, lapply(categories, process_mediation_results, elective_status = "not_elective"))

# Remove rownames
rownames(mediation_results_elective) <- NULL
rownames(mediation_results_not_elective) <- NULL

# Add total effect (Odds ratio) to the tables
total_effect_elective <- total_effect_elective %>%
  mutate(lb = estimated - 1.96 * se, ub = estimated + 1.96 * se) %>%
  mutate(or = exp(estimated), or.lb = exp(lb), or.ub = exp(ub)) %>%
  transmute(Estimate = or, '95% CI Upper' = or.ub, '95% CI Lower' = or.lb, cat = 5, 'p-value' = p, results = "Total Effect (OR)")

total_effect_not_elective <- total_effect_not_elective %>%
  mutate(lb = estimated - 1.96 * se, ub = estimated + 1.96 * se) %>%
  mutate(or = exp(estimated), or.lb = exp(lb), or.ub = exp(ub)) %>%
  transmute(Estimate = or, '95% CI Upper' = or.ub, '95% CI Lower' = or.lb, cat = 5, 'p-value' = p, results = "Total Effect (OR)")

# Combine results
results_overall_elective <- rbind(mediation_results_elective, total_effect_elective)
results_overall_not_elective <- rbind(mediation_results_not_elective, total_effect_not_elective)

# Remove rownames
rownames(results_overall_elective) <- NULL
rownames(results_overall_not_elective) <- NULL

```

## Results export

```{r}
# Save overall elective and not elective results
save_to_word(results_overall_elective, "sup_tables/Mediation Analysis Elective Surgery.docx","Mediation Analysis Elective Surgery")
save_to_word(results_overall_not_elective, "sup_tables/Mediation Analysis Not Elective Surgery.docx","Mediation Analysis Not Elective Surgery")

```
