---
title: "Sodium GCS Mediation Mortality (Analysis) - Figures"
author: "Sicheng, Miguel"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output:
  # html_notebook:
    code_folding: hide
    number_sections: yes
    theme: flatly
    toc: yes
    toc_float: yes

knit: (function(inputFile, encoding) { rmarkdown::render(inputFile, encoding = encoding, output_file = paste0(substr(inputFile,1,nchar(inputFile)-4)," ",Sys.Date(),'.html')) })
---

# Environment

```{r message=FALSE, warning=FALSE,echo = F}
library(dplyr)
library(ggplot2)
library(gridExtra)
library(stringi)
library(purrr)
library(cowplot)
'%!in%' <- function(x,y)!('%in%'(x,y))
```

# Figure 3

## Figure 3A: OVERALL population

### Appending results for plot

  - *ACME:* Average Causal Mediation Effects. This is the indirect effect of the IV on the outcome that goes through the mediator.
  - *ADE:* Average direct effects. It describes the direct effect of the IV on the Outcome.
  - *Total Effect* the Total Effect (direct + indirect) of the IV on the Outcome.
  - *Prop. Mediated* describes the proportion of the effect of the IV on the Outcome that goes through the mediator. It's calculated by dividing the ACME through the total effect.

```{r}
# we want to be able to know which dataset our results are coming from
mediation_results_cat1_sum$cat<-1
mediation_results_cat1_sum$results<-stri_replace_all_fixed(rownames(mediation_results_cat1_sum),'(treated)',"" )
mediation_results_cat2_sum$cat<-2
mediation_results_cat2_sum$results<-stri_replace_all_fixed(rownames(mediation_results_cat2_sum),'(treated)',"" )
mediation_results_cat3_sum$cat<-3
mediation_results_cat3_sum$results<-stri_replace_all_fixed(rownames(mediation_results_cat3_sum),'(treated)',"" )
mediation_results_cat4_sum$cat<-4
mediation_results_cat4_sum$results<-stri_replace_all_fixed(rownames(mediation_results_cat4_sum),'(treated)',"" )
mediation_results_cat6_sum$cat<-6
mediation_results_cat6_sum$results<-stri_replace_all_fixed(rownames(mediation_results_cat6_sum),'(treated)',"" )
mediation_results_cat7_sum$cat<-7
mediation_results_cat7_sum$results<-stri_replace_all_fixed(rownames(mediation_results_cat7_sum),'(treated)',"" )
mediation_results_cat8_sum$cat<-8
mediation_results_cat8_sum$results<-stri_replace_all_fixed(rownames(mediation_results_cat8_sum),'(treated)',"" )

mediation_results_all_overall<-rbind(
 mediation_results_cat1_sum
,mediation_results_cat2_sum
,mediation_results_cat3_sum
,mediation_results_cat4_sum
,mediation_results_cat6_sum
,mediation_results_cat7_sum
,mediation_results_cat8_sum
)
rownames(mediation_results_all_overall)<-NULL

#add total effect (Odds ratio) to the table

total_effect_forplot_overall<-rbind(total_effect, data.frame(cat = 5,estimated = 0,se = 0,p = NA))

total_effect_forplot_overall<-total_effect_forplot_overall%>%
  mutate(lb = estimated -1.96*se,
         ub = estimated + 1.96*se,
         or = exp(estimated),
         or.lb = exp(lb),
         or.ub = exp(ub))

total_effect_overall_copy<-total_effect_forplot_overall%>%
  transmute(Estimate = or, 
            '95% CI Upper' = or.ub,
            '95% CI Lower' = or.lb,
            cat = cat,'p-value' = p,
            results = "Total Effect (OR)")

results_overall<-rbind(mediation_results_all_overall,total_effect_overall_copy)
```

### Exporing results for suplementary table

```{r}
table_results_overall<-results_overall %>%
  filter(results == "Total Effect (OR)") %>%
  mutate(across(where(is.numeric), ~ round(., 2)))

table_results_overall$`p-value`<-NULL
```


### Final subfigure: Prop. Mediated + Total Effect

· 
caption = Green Bars: Proportion of mortality mediated by GCS score\nBlue Dots: Odds Ratio of Hospital Mortality\nBlue Ribbon: Odds ratio's 95% CI

```{r }
mediation_results_overall_propmediated<-mediation_results_all_overall%>%
  filter(results %in% c("Prop. Mediated (average)") ) %>%
  dplyr::select(Estimate,cat,results)

mediation_results_overall_propmediated<-rbind(mediation_results_overall_propmediated,
                                      data.frame(Estimate = 0, cat = 5, results = "Prop. Mediated (average)"))

spline_x <-spline(total_effect_forplot_overall$cat, (total_effect_forplot_overall$or - 1) * 100, n = 1000)$x
spline_hi <-spline(total_effect_forplot_overall$cat, (total_effect_forplot_overall$or.ub - 1) * 100, n = 1000)$y
spline_lo <-spline(total_effect_forplot_overall$cat, (total_effect_forplot_overall$or.lb - 1) * 100, n = 1000)$y

foo_ribbon <-data.frame(spline_x, spline_hi, spline_lo)

fig3a <- ggplot(data = mediation_results_overall_propmediated, aes(x = cat, y = Estimate * 100)) +
  geom_bar(stat = 'identity', fill = "#27ae60") +
  xlab('Baseline Sodium Category') +
  ylab('%') +
  geom_text(data = mediation_results_overall_propmediated, aes(label = paste(round(Estimate * 100, 2), '%')), colour = "#2c3e50", vjust = 1.5) +
  scale_y_continuous(breaks = c(0, 50, 100)) +
  theme(legend.position = "bottom", plot.title.position = "plot", plot.caption.position =  "plot") +
  theme_minimal() +
  scale_x_continuous(breaks = c(1, 2, 3, 4, 5, 6, 7, 8),
                     labels = c("<120", "120- <125", "125- <130", "130- <135", "135- <140\n(Reference)", "140- <145", "145- <150", ">150")) +
  geom_point(data = total_effect_forplot_overall, aes(x = cat, y = (or - 1) * 100), colour = "#34495e") +
  geom_smooth(data = total_effect_forplot_overall, aes(x = cat, y = (or - 1) * 100), method = "loess", span = 0.5, se = FALSE, colour = "#34495e") +
  geom_ribbon(data = foo_ribbon, aes(x = spline_x, y = NULL, ymin = spline_lo, ymax = spline_hi),
              alpha = 0.4, inherit.aes = FALSE, width = 0.2, fill = '#2980b9') +
  scale_y_continuous(position = "right",
                     name = "Proportion of mortality mediated by GCS score",
                     sec.axis = sec_axis(~(. / 100) + 1, name = "Odds ratio (95%CI) of hospital mortality"))
```

## Figure 3B: NOT MECHANICALLY VENTILATED Subanalysis

### Appending results for plot

```{r}
# Convertir la lista en un dataframe
mediation_results_all_not_mechvent <- map_df(seq_along(mediation_summaries_not_mechvent), function(i) {
  if (is.null(mediation_summaries_not_mechvent[[i]])) {
    return(NULL)
  }
  df <- as.data.frame(mediation_summaries_not_mechvent[[i]])
  df$cat <- i  # Asignar el número de la categoría a cada fila
  return(df)
})



# Agregar la columna de resultados basados en los nombres de las filas originales
mediation_results_all_not_mechvent <- mediation_results_all_not_mechvent %>%
  mutate(results = case_when(
    row_number() %% 4 == 1 ~ "Total Effect",
    row_number() %% 4 == 2 ~ "ACME (average)",
    row_number() %% 4 == 3 ~ "ADE (average)",
    row_number() %% 4 == 0 ~ "Prop. Mediated (average)"
  ))

#add total effect (Odds ratio) to the table

total_effect_forplot_not_mechvent<-rbind(total_effect_not_mechvent,data.frame(cat = 5,estimated = 0,se = 0,p = NA))

total_effect_forplot_not_mechvent <- total_effect_forplot_not_mechvent %>%
  mutate(lb = estimated - 1.96 * se, 
         ub = estimated + 1.96 * se,
         or = exp(estimated), 
         or.lb = exp(lb), 
         or.ub = exp(ub))

# Crear una copia con las transformaciones específicas
total_effect_forplot_not_mechvent_copy <- total_effect_forplot_not_mechvent %>%
  transmute(Estimate = or, 
            '95% CI Upper' = or.ub, 
            '95% CI Lower' = or.lb, 
            cat = cat, 
            'p-value' = p, 
            results = "Total Effect (OR)")

results_not_mechvent<-rbind(mediation_results_all_not_mechvent,total_effect_forplot_not_mechvent_copy)

rownames(results_not_mechvent)<-c()
```

### Final subfigure: Prop. Mediated + Total Effect

```{r}
# Filter and select relevant data for Prop. Mediated (average)
mediation_results_not_mechvent_propmediated <- results_not_mechvent %>%
  filter(results %in% c("Prop. Mediated (average)")) %>%
  dplyr::select(Estimate, cat, results)

# Add a dummy data point for category 5 with an Estimate of 0
mediation_results_not_mechvent_propmediated <- rbind(
  mediation_results_not_mechvent_propmediated,
  data.frame(Estimate = 0, cat = 5, results = "Prop. Mediated (average)"))

spline_x <- spline(total_effect_forplot_not_mechvent$cat, (total_effect_forplot_not_mechvent$or - 1) * 100, n = 1000)$x
spline_hi <- spline(total_effect_forplot_not_mechvent$cat, (total_effect_forplot_not_mechvent$or.ub - 1) * 100, n = 1000)$y
spline_lo <- spline(total_effect_forplot_not_mechvent$cat, (total_effect_forplot_not_mechvent$or.lb - 1) * 100, n = 1000)$y

foo_ribbon <- data.frame(spline_x, spline_hi, spline_lo)

# Plotting the results
fig3b <- ggplot(data = mediation_results_not_mechvent_propmediated, aes(x = cat, y = Estimate * 100)) +
  geom_bar(stat = 'identity', fill = "#27ae60") +
  xlab('Baseline Sodium Category') +
  ylab('%') +
  geom_text(data = mediation_results_not_mechvent_propmediated, aes(label = paste(round(Estimate * 100, 2), '%')), colour = "#2c3e50", vjust = 1.5) +
  scale_y_continuous(breaks = c(0, 50, 100)) +
  theme(legend.position = "bottom", plot.title.position = "plot", plot.caption.position = "plot") +
  theme_minimal() +
  scale_x_continuous(breaks = c(1, 2, 3, 4, 5, 6, 7, 8),
                     labels = c("<120", "120- <125", "125- <130", "130- <135", "135- <140\n(Reference)", "140- <145", "145- <150", ">150")) +
  geom_point(data = total_effect_forplot_not_mechvent, aes(x = cat, y = (or - 1) * 100), colour = "#34495e") +
  geom_smooth(data = total_effect_forplot_not_mechvent, aes(x = cat, y = (or - 1) * 100), method = "loess", span = 0.5, se = FALSE, colour = "#34495e") +
  geom_ribbon(data = foo_ribbon, aes(x = spline_x, y = NULL, ymin = spline_lo, ymax = spline_hi),
              alpha = 0.4, inherit.aes = FALSE, width = 0.2, fill = '#2980b9') +
  scale_y_continuous(position = "right", 
                     limits = c(-90, 245),
                     name = "Proportion of mortality mediated by GCS score",
                     sec.axis = sec_axis(~(. / 100) + 1, name = "Odds ratio (95%CI) of hospital mortality")) +
  labs(caption = "")
```

## Figure 3C: MECHANICALLY VENTILATED Subanalysis

### Appending results for plot

```{r}
# Convertir la lista en un dataframe
mediation_results_all_mechvent <- map_df(seq_along(mediation_summaries_mechvent), function(i) {
  if (is.null(mediation_summaries_mechvent[[i]])) {
    return(NULL)
  }
  df <- as.data.frame(mediation_summaries_mechvent[[i]])
  df$cat <- i  # Asignar el número de la categoría a cada fila
  return(df)
})

# Agregar la columna de resultados basados en los nombres de las filas originales
mediation_results_all_mechvent <- mediation_results_all_mechvent %>%
  mutate(results = case_when(
    row_number() %% 4 == 1 ~ "Total Effect",
    row_number() %% 4 == 2 ~ "ACME (average)",
    row_number() %% 4 == 3 ~ "ADE (average)",
    row_number() %% 4 == 0 ~ "Prop. Mediated (average)"
  ))

#add total effect (Odds ratio) to the table

total_effect_forplot_mechvent<-rbind(total_effect_mechvent,data.frame(cat = 5,estimated = 0,se = 0,p = NA))

total_effect_forplot_mechvent <- total_effect_forplot_mechvent %>%
  mutate(lb = estimated - 1.96 * se, 
         ub = estimated + 1.96 * se,
         or = exp(estimated), 
         or.lb = exp(lb), 
         or.ub = exp(ub))

# Crear una copia con las transformaciones específicas
total_effect_forplot_mechvent_copy <- total_effect_forplot_mechvent %>%
  transmute(Estimate = or, 
            '95% CI Upper' = or.ub, 
            '95% CI Lower' = or.lb, 
            cat = cat, 
            'p-value' = p, 
            results = "Total Effect (OR)")

results_mechvent<-rbind(mediation_results_all_mechvent,total_effect_forplot_mechvent_copy)

rownames(results_mechvent)<-c()
```

### Final subfigure: Prop. Mediated + Total Effect

```{r}
# Filter and select relevant data for Prop. Mediated (average)
mediation_results_mechvent_propmediated <- results_mechvent %>%
  filter(results %in% c("Prop. Mediated (average)")) %>%
  dplyr::select(Estimate, cat, results)

# Add a dummy data point for category 5 with an Estimate of 0
mediation_results_mechvent_propmediated <- rbind(
  mediation_results_mechvent_propmediated,
  data.frame(Estimate = 0, cat = 5, results = "Prop. Mediated (average)"))

spline_x <- spline(total_effect_forplot_mechvent$cat, (total_effect_forplot_mechvent$or - 1) * 100, n = 1000)$x
spline_hi <- spline(total_effect_forplot_mechvent$cat, (total_effect_forplot_mechvent$or.ub - 1) * 100, n = 1000)$y
spline_lo <- spline(total_effect_forplot_mechvent$cat, (total_effect_forplot_mechvent$or.lb - 1) * 100, n = 1000)$y

foo_ribbon <- data.frame(spline_x, spline_hi, spline_lo)

# Plotting the results
fig3c <- ggplot(data = mediation_results_mechvent_propmediated, aes(x = cat, y = Estimate * 100)) +
  geom_bar(stat = 'identity', fill = "#3498db") +
  xlab('Baseline Sodium Category') +
  ylab('%') +
  geom_text(data = mediation_results_mechvent_propmediated, aes(label = paste(round(Estimate * 100, 2), '%')), colour = "#2c3e50", vjust = 1.5) +
  scale_y_continuous(breaks = c(0, 50, 100)) +
  theme(legend.position = "bottom", plot.title.position = "plot", plot.caption.position = "plot") +
  theme_minimal() +
  scale_x_continuous(breaks = c(1, 2, 3, 4, 5, 6, 7, 8),
                     labels = c("<120", "120- <125", "125- <130", "130- <135", "135- <140\n(Reference)", "140- <145", "145- <150", ">150")) +
  geom_point(data = total_effect_forplot_mechvent, aes(x = cat, y = (or - 1) * 100), colour = "#c0392b") +
  geom_smooth(data = total_effect_forplot_mechvent, aes(x = cat, y = (or - 1) * 100), method = "loess", span = 0.5, se = FALSE, colour = "#c0392b") +
  geom_ribbon(data = foo_ribbon, aes(x = spline_x, y = NULL, ymin = spline_lo, ymax = spline_hi),
              alpha = 0.4, inherit.aes = FALSE, width = 0.2, fill = '#e74c3c') +
  scale_y_continuous(position = "right", 
                     limits = c(-90, 245),
                     name = "Proportion of mortality mediated by GCS score",
                     sec.axis = sec_axis(~(. / 100) + 1, name = "Odds ratio (95%CI) of hospital mortality")) +
  labs(caption = "")
```

## Combining all subfigures

```{r}
# Assuming plot_supA and plot_supB are already created
# Arrange the plots side by side
combined_plot <- combined_plot <- plot_grid(
  fig3a, 
  plot_grid(fig3b, fig3c, ncol = 2), 
  ncol = 1,
  rel_heights = c(1, 2)
)

# Display the combined plot
print(combined_plot)
```


## Export the ggplot figure as a TIFF image

```{r}
tiff("figures/figure3.tiff", width = 16*250, height = 2*9*200, res = 300)
print(combined_plot)
dev.off()
```

# Supplemental Figure 1: SEDATED BY MED Subanalysis

Mediating effect of decreased consciousness on the association between sodium concentrations and mortality among individuals “sedated by medications”

## Appending results for plot

-   *ACME:* Average Causal Mediation Effects. This is the indirect effect of the IV on the outcome that goes through the mediator.
-   *ADE:* Average direct effects. It describes the direct effect of the IV on the Outcome.
-   *Total Effect* the Total Effect (direct + indirect) of the IV on the Outcome.
-   *Prop. Mediated* describes the proportion of the effect of the IV on the Outcome that goes through the mediator. It's calculated by dividing the ACME through the total effect.

```{r}
# we want to be able to know which dataset our results are coming from
mediation_results_cat1_sedated_sum$cat<-1
mediation_results_cat1_sedated_sum$results<-stri_replace_all_fixed(rownames(mediation_results_cat1_sedated_sum),'(treated)',"" )
mediation_results_cat2_sedated_sum$cat<-2
mediation_results_cat2_sedated_sum$results<-stri_replace_all_fixed(rownames(mediation_results_cat2_sedated_sum),'(treated)',"" )
mediation_results_cat3_sedated_sum$cat<-3
mediation_results_cat3_sedated_sum$results<-stri_replace_all_fixed(rownames(mediation_results_cat3_sedated_sum),'(treated)',"" )
mediation_results_cat4_sedated_sum$cat<-4
mediation_results_cat4_sedated_sum$results<-stri_replace_all_fixed(rownames(mediation_results_cat4_sedated_sum),'(treated)',"" )
mediation_results_cat6_sedated_sum$cat<-6
mediation_results_cat6_sedated_sum$results<-stri_replace_all_fixed(rownames(mediation_results_cat6_sedated_sum),'(treated)',"" )
mediation_results_cat7_sedated_sum$cat<-7
mediation_results_cat7_sedated_sum$results<-stri_replace_all_fixed(rownames(mediation_results_cat7_sedated_sum),'(treated)',"" )
mediation_results_cat8_sedated_sum$cat<-8
mediation_results_cat8_sedated_sum$results<-stri_replace_all_fixed(rownames(mediation_results_cat8_sedated_sum),'(treated)',"" )
mediation_results_all_sedated<-rbind(
 mediation_results_cat1_sedated_sum
,mediation_results_cat2_sedated_sum
,mediation_results_cat3_sedated_sum
,mediation_results_cat4_sedated_sum
,mediation_results_cat6_sedated_sum
,mediation_results_cat7_sedated_sum
,mediation_results_cat8_sedated_sum
)

rownames(mediation_results_all_sedated)<-NULL

#add total effect (Odds ratio) to the table

total_effect_forplot_sedated<-rbind(total_effect_sedated,data.frame(cat = 5,estimated = 0,se = 0,p = NA))

total_effect_forplot_sedated<-total_effect_forplot_sedated%>%
  mutate(lb = estimated -1.96*se, 
         ub = estimated + 1.96*se,
         or = exp(estimated),
         or.lb = exp(lb),
         or.ub = exp(ub))

total_effect_forplot_sedated_copy<-total_effect_forplot_sedated%>%
  transmute(Estimate = or, 
            '95% CI Upper' = or.ub,
            '95% CI Lower' = or.lb,
            cat = cat,
            'p-value' = p,
            results = "Total Effect (OR)")

results_sedated<-rbind(mediation_results_all_sedated,total_effect_forplot_sedated_copy)
```

## Final subfigure: Prop. Mediated + Total Effect

```{r }
mediation_results_sedated_propmediated<-mediation_results_all_sedated%>%
  filter(results %in% c("Prop. Mediated (average)") ) %>%
  dplyr::select(Estimate,cat,results)

mediation_results_sedated_propmediated<-rbind(mediation_results_sedated_propmediated,
                                      data.frame(Estimate = 0, cat = 5, results = "Prop. Mediated (average)"))

spline_x <-spline(total_effect_forplot_sedated$cat, (total_effect_forplot_sedated$or - 1) * 100, n = 1000)$x
spline_hi <-spline(total_effect_forplot_sedated$cat, (total_effect_forplot_sedated$or.ub - 1) * 100, n = 1000)$y
spline_lo <-spline(total_effect_forplot_sedated$cat, (total_effect_forplot_sedated$or.lb - 1) * 100, n = 1000)$y

foo_ribbon <-data.frame(spline_x, spline_hi, spline_lo)

sup_fig1 <- ggplot(data = mediation_results_sedated_propmediated, aes(x = cat, y = Estimate * 100)) +
  geom_bar(stat = 'identity', fill = "#3498db") +
  xlab('Baseline Sodium Category') +
  ylab('%') +
  geom_text(data = mediation_results_sedated_propmediated, aes(label = paste(round(Estimate * 100, 2), '%')), colour = "#2c3e50", vjust = 1.5) +
  scale_y_continuous(breaks = c(0, 50, 100)) +
  theme(legend.position = "bottom", plot.title.position = "plot", plot.caption.position =  "plot") +
  theme_minimal() +
  scale_x_continuous(breaks = c(1, 2, 3, 4, 5, 6, 7, 8),
                     labels = c("<120", "120- <125", "125- <130", "130- <135", "135- <140\n(Reference)", "140- <145", "145- <150", ">150")) +
  geom_point(data = total_effect_forplot_sedated, aes(x = cat, y = (or - 1) * 100), colour = "#c0392b") +
  geom_smooth(data = total_effect_forplot_sedated, aes(x = cat, y = (or - 1) * 100), method = "loess", span = 0.5, se = FALSE, colour = "#c0392b") +
  geom_ribbon(data = foo_ribbon, aes(x = spline_x, y = NULL, ymin = spline_lo, ymax = spline_hi),
              alpha = 0.4, inherit.aes = FALSE, width = 0.2, fill = '#e74c3c') +
  scale_y_continuous(position = "right", 
                     limits = c(-90, 245),
                     name = "Proportion of mortality mediated by GCS score",
                     sec.axis = sec_axis(~(. / 100) + 1, name = "Odds ratio (95%CI) of hospital mortality")) +  
labs(caption = "")
```

## Export the ggplot figure as a TIFF image

```{r}
tiff("figures/sup_fig1.tiff", width = 16*250, height = 9*250, res = 300)
print(sup_fig1)
dev.off()
```