---
title: "Sodium GCS Mediation Mortality (Analysis) - Figures"
author: "Sicheng, Miguel"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output:
  # html_notebook:
    code_folding: hide
    number_sections: yes
    theme: flatly
    toc: yes
    toc_float: yes

knit: (function(inputFile, encoding) { rmarkdown::render(inputFile, encoding = encoding, output_file = paste0(substr(inputFile,1,nchar(inputFile)-4)," ",Sys.Date(),'.html')) })
---

# Environment

```{r message=FALSE, warning=FALSE,echo = F}
library(dplyr)
library(ggplot2)
library(gridExtra)
library(stringi)
'%!in%' <- function(x,y)!('%in%'(x,y))
```

# Appending results for plot

-   *ACME:* Average Causal Mediation Effects. This is the indirect effect of the IV on the outcome that goes through the mediator.
-   *ADE:* Average direct effects. It describes the direct effect of the IV on the Outcome.
-   *Total Effect* the Total Effect (direct + indirect) of the IV on the Outcome.
-   *Prop. Mediated* describes the proportion of the effect of the IV on the Outcome that goes through the mediator. It's calculated by dividing the ACME through the total effect.

```{r}
# we want to be able to know which dataset our results are coming from
mediation_results_cat1_sum$cat<-1
mediation_results_cat1_sum$results<-stri_replace_all_fixed(rownames(mediation_results_cat1_sum),'(treated)',"" )
mediation_results_cat2_sum$cat<-2
mediation_results_cat2_sum$results<-stri_replace_all_fixed(rownames(mediation_results_cat2_sum),'(treated)',"" )
mediation_results_cat3_sum$cat<-3
mediation_results_cat3_sum$results<-stri_replace_all_fixed(rownames(mediation_results_cat3_sum),'(treated)',"" )
mediation_results_cat4_sum$cat<-4
mediation_results_cat4_sum$results<-stri_replace_all_fixed(rownames(mediation_results_cat4_sum),'(treated)',"" )
mediation_results_cat6_sum$cat<-6
mediation_results_cat6_sum$results<-stri_replace_all_fixed(rownames(mediation_results_cat6_sum),'(treated)',"" )
mediation_results_cat7_sum$cat<-7
mediation_results_cat7_sum$results<-stri_replace_all_fixed(rownames(mediation_results_cat7_sum),'(treated)',"" )
mediation_results_cat8_sum$cat<-8
mediation_results_cat8_sum$results<-stri_replace_all_fixed(rownames(mediation_results_cat8_sum),'(treated)',"" )
mediation_results_all<-rbind(
 mediation_results_cat1_sum
,mediation_results_cat2_sum
,mediation_results_cat3_sum
,mediation_results_cat4_sum
,mediation_results_cat6_sum
,mediation_results_cat7_sum
,mediation_results_cat8_sum
)
rownames(mediation_results_all)<-NULL

#add total effect (Odds ratio) to the table

total_effect<-rbind(total_effect,data.frame(cat = 5,estimated = 0,se = 0,p = NA))

total_effect<-total_effect%>%
  mutate(lb = estimated -1.96*se, ub = estimated + 1.96*se)%>%
  mutate(or = exp(estimated),or.lb = exp(lb),or.ub = exp(ub))

total_effect_copy<-total_effect%>%transmute(Estimate = or, '95% CI Upper' = or.ub,'95% CI Lower' = or.lb,cat = cat,'p-value' = p,results = "Total Effect (OR)")

results_overall<-rbind(mediation_results_all,total_effect_copy)
```

# Figure 3: Prop. Mediated + Total Effect

```{r }
mediation_results_propmediated<-mediation_results_all%>%
  filter(results %in% c("Prop. Mediated (average)") ) %>%
  dplyr::select(Estimate,cat,results)

mediation_results_propmediated<-rbind(mediation_results_propmediated,
                                      data.frame(Estimate = 0, cat = 5, results = "Prop. Mediated (average)"))

spline_x <-spline(total_effect$cat, (total_effect$or - 1) * 100, n = 1000)$x
spline_hi <-spline(total_effect$cat, (total_effect$or.ub - 1) * 100, n = 1000)$y
spline_lo <-spline(total_effect$cat, (total_effect$or.lb - 1) * 100, n = 1000)$y

foo_ribbon <-data.frame(spline_x, spline_hi, spline_lo)

p <- ggplot(data = mediation_results_propmediated, aes(x = cat, y = Estimate * 100)) +
  geom_bar(stat = 'identity', fill = "#27ae60") +
  xlab('Baseline Sodium Category') +
  ylab('%') +
  geom_text(data = mediation_results_propmediated, aes(label = paste(round(Estimate * 100, 2), '%')), colour = "#2c3e50", vjust = 1.5) +
  scale_y_continuous(breaks = c(0, 50, 100)) +
  theme(legend.position = "bottom", plot.title.position = "plot", plot.caption.position =  "plot") +
  theme_minimal() +
  scale_x_continuous(breaks = c(1, 2, 3, 4, 5, 6, 7, 8),
                     labels = c("<120", "120- <125", "125- <130", "130- <135", "135- <140\n(Reference)", "140- <145", "145- <150", ">150")) +
  geom_point(data = total_effect, aes(x = cat, y = (or - 1) * 100), colour = "#34495e") +
  geom_smooth(data = total_effect, aes(x = cat, y = (or - 1) * 100), method = "loess", span = 0.5, se = FALSE, colour = "#34495e") +
  geom_ribbon(data = foo_ribbon, aes(x = spline_x, y = NULL, ymin = spline_lo, ymax = spline_hi),
              alpha = 0.4, inherit.aes = FALSE, width = 0.2, fill = '#2980b9') +
  scale_y_continuous(position = "right",
                     name = "Proportion of mortality mediated by GCS score",
                     sec.axis = sec_axis(~(. / 100) + 1, name = "Odds ratio (95%CI) of hospital mortality")) +
  labs(caption = "Green Bars: Proportion of mortality mediated by GCS score\nBlue Dots: Odds Ratio of Hospital Mortality\nBlue Ribbon: Odds ratio's 95% CI") 

p


# Export the ggplot figure as a TIFF image
tiff("figures/figure3.tiff", width = 2300, height = 1500, res = 300)
print(p)
dev.off()

```

# Deprecated figures

## Prop. Mediated

```{r}
mediation_results_filtered<-mediation_results_all%>% filter(results %in% c('Prop. Mediated (average)'))
ggplot(data=mediation_results_filtered,aes(x=cat,y=Estimate,colour=results,fill=results))+
    geom_point()+
    geom_line()+
    geom_errorbar(data=mediation_results_filtered
                  ,aes(x=cat,ymin=`95% CI Lower`,ymax=`95% CI Upper`,fill=results)
                  ,alpha=0.3
                  ,inherit.aes=FALSE
                  ,width=0.2)+
    xlab('Baseline Sodium Category')+
    ylab('Estimate')+
    #scale_color_manual(labels = c("Male", "Female"), values = c("#1abc9c","#f1c40f"))+
    #scale_fill_manual(labels = c("Male", "Female"), values = c("#1abc9c","#f1c40f"))+
    #labs(colour='sex', fill='sex') +
    theme_minimal()+
  ggtitle('Proportion Mediated')+
  scale_x_continuous(breaks = c(1,2,3,4,6,7,8))
```

## Indirect Effect

```{r}
mediation_results_filtered<-mediation_results_all%>% filter(results %in% c('ACME (average)'))
ggplot(data=mediation_results_filtered,aes(x=cat,y=Estimate,colour=results,fill=results))+
    geom_point()+
    geom_line()+
    geom_errorbar(data=mediation_results_filtered
                  ,aes(x=cat,ymin=`95% CI Lower`,ymax=`95% CI Upper`,fill=results)
                  ,alpha=0.3
                  ,inherit.aes=FALSE
                  ,width=0.2)+
    xlab('Baseline Sodium Category')+
    ylab('Estimate')+
    #scale_color_manual(labels = c("Male", "Female"), values = c("#1abc9c","#f1c40f"))+
    #scale_fill_manual(labels = c("Male", "Female"), values = c("#1abc9c","#f1c40f"))+
    #labs(colour='sex', fill='sex') +
    theme_minimal()+
  ggtitle('ACME\nAverage Causal Mediation Effects\nIndirect effect of Sodium on mortality that goes through the mediator (GCS)')+
  scale_x_continuous(breaks = c(1,2,3,4,6,7,8))
```

## Direct Effect

```{r}
mediation_results_filtered<-mediation_results_all%>% filter(results %in% c('ADE (average)'))
ggplot(data=mediation_results_filtered,aes(x=cat,y=Estimate,colour=results,fill=results))+
    geom_point()+
    geom_line()+
    geom_errorbar(data=mediation_results_filtered
                  ,aes(x=cat,ymin=`95% CI Lower`,ymax=`95% CI Upper`,fill=results)
                  ,alpha=0.3
                  ,inherit.aes=FALSE
                  ,width=0.2)+
    xlab('Baseline Sodium Category')+
    ylab('Estimate')+
    #scale_color_manual(labels = c("Male", "Female"), values = c("#1abc9c","#f1c40f"))+
    #scale_fill_manual(labels = c("Male", "Female"), values = c("#1abc9c","#f1c40f"))+
    #labs(colour='sex', fill='sex') +
    theme_minimal()+
  ggtitle("Average Direct Effect")+
  scale_x_continuous(breaks = c(1,2,3,4,6,7,8))
```

## Total Effect from Mediation Package

```{r}
mediation_results_filtered<-mediation_results_all%>% filter(results %in% c('Total Effect'))
ggplot(data=mediation_results_filtered,aes(x=cat,y=Estimate,colour=results,fill=results))+
    geom_point()+
    geom_line()+
    geom_errorbar(data=mediation_results_filtered
                  ,aes(x=cat,ymin=`95% CI Lower`,ymax=`95% CI Upper`,fill=results)
                  ,alpha=0.3
                  ,inherit.aes=FALSE
                  ,width=0.2)+
    xlab('Baseline Sodium Category')+
    ylab('Estimate')+
    #scale_color_manual(labels = c("Male", "Female"), values = c("#1abc9c","#f1c40f"))+
    #scale_fill_manual(labels = c("Male", "Female"), values = c("#1abc9c","#f1c40f"))+
    #labs(colour='sex', fill='sex') +
    theme_minimal()+
  ggtitle('Total Effect')+
  scale_x_continuous(breaks = c(1,2,3,4,6,7,8))
```

## Total Effect in Odds Ratio

```{r}

ggplot(data=total_effect,aes(x=cat,y=or))+
    geom_point()+
    geom_line(aes(color = "red"))+
    geom_errorbar(data=total_effect
                  ,aes(x=cat,ymin=or.lb,ymax=or.ub)
                  ,alpha=0.3
                  ,inherit.aes=FALSE
                  ,width=0.2)+
    xlab('Baseline Sodium Category')+
    ylab('Estimate')+
    #scale_color_manual(labels = c("Male", "Female"), values = c("#1abc9c","#f1c40f"))+
    #scale_fill_manual(labels = c("Male", "Female"), values = c("#1abc9c","#f1c40f"))+
    #labs(colour='sex', fill='sex') +
    theme_minimal()+
  ggtitle('Total Effect')+
  scale_x_continuous(breaks = c(1,2,3,4,5,6,7,8))
```

